### 2019-08-14 content/_index.Rmd
Customize course visuals

---
title: "Econometrics"
author: "Ryan Safner"
---

^[{-} `r blogdown::shortcode("courseinfo")`] <span class="newthought">This site contains</span> the syllabus, schedule, and assignments for ECON 480: Econometrics, held during Fall 2019 at Hood College.

By the end of this course, you will:

1. understand how to evaluate statistical and empirical claims;
2. use the fundamental models of causal inference and research design;
2. gather, analyze, and communicate with real data in R.

<figure>
[![](/images/metrics_flowchart_blankbg.png)](/images/metrics_flowchart_blankbg.png)
</figure>