### 2019-08-14 content/_index.Rmd
Updated learning goals

---
title: "Econometrics"
author: "Ryan Safner"
---

^[{-} `r blogdown::shortcode("courseinfo")`] <span class="newthought">This site contains</span> the syllabus, schedule, and assignments for ECON 480: Econometrics, held during Fall 2019 at Hood College.

By the end of this course, you will:

1. understand how to evaluate statistical and empirical claims;
2. use the fundamental models of causal inference and research design;
3. gather, analyze, and communicate with real data in R.

Given these objectives, this course fulfills all three of the learning outcomes for [the George B. Delaplaine, Jr. School of Business](https://www.hood.edu/academics/departments/george-b-delaplaine-jr-school-business/student-learning-outcomes) Economics B.A. program:

- Use quantitative tools and techniques in the preparation, interpretation, analysis and presentation of data and information for problem solving and decision making [...]
- Apply economic reasoning and models to understand and analyze problems of public policy [...]
- Demonstrate effective oral and written communications skills for personal and professional success[...]

<figure>
[![](/images/metrics_flowchart_blankbg.png)](/images/metrics_flowchart_blankbg.png)
</figure>