### 2019-08-22 content/schedule.Rmd

---
title: "Schedule"
slug: schedule
---

This page contains all of the following resources for each class meeting:

- <i class="fas fa-book-reader"></i> **Readings** include textbook chapters and occasional journal articles
- <i class="fas fa-laptop-code"></i> **Assignments** are due by the beginning of class unless otherwise stated
- <i class="fas fa-info-circle"></i> **Class** materials contain extra explanations, math/statistical properties & proofs, and other helpful resources^[These "online appendices" keep the slides nice and de-cluttered!]
- <i class="fas fa-chalkboard-teacher"></i> **Slides** are "[Xaringan](https://github.com/yihui/xaringan)" presentations in html that can be opened in any browser^[`File > Print > Save as PDF` if you wish to download them]
- <i class="fas fa-registered"></i> **R** materials contain extra tutorials, videos, practice exercises for using `R`

Relevant materials (if applicable, icons will become links) will be posted before class meets.

**Last Update**: [`r Sys.time()`](https://github.com/ryansafner/metricsF19/commits/master).

---

`r blogdown::shortcode("schedule")`