### 2019-08-27 content/class/01-class.Rmd
Update for first day!

---
title: "Class 1: Introduction to Econometrics"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

# Overview

An introduction to the course.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/01-slides.html)

# Slack

- [Click here to join our class Slack channel](https://join.slack.com/t/hoodeconometrics/shared_invite/enQtNzQxMzI1MTU0ODM4LTVhOWI5Zjc5YzdiOGU5YzY0MmZiM2QxOGQwMWQwMjc4Y2FmYTk5Yjg5ZDZhZTY4ZDZiYWU2YzAzMTU1MDIwYTE)^[This link will expire in 7 days]