### 2019-08-29 static/slides/02-slides.Rmd

---
title: "1.2: The Quest for Causality"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: ocean #rainbow #monokai rainbow solarized-light #tomorrow-night -blue -eighties zenburn
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
               warning=F,
               message=F)
set.seed(256)
```

class: inverse, center, middle

# The 2 Big Problems with Data

---

# Two Big Problems with Data

.pull-left[

- We want to use econometrics to **identify** causal relationships and make **inferences** about them

1. Problem for **identification**: .shout[endogeneity]

2. Problem for **inference**: .shout[randomness]

]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/phhet82tcnuegxp/randomimage.jpg?raw=1)]
]

---

# Identification Problem: Endogeneity I

.pull-left[

- An independent variable $(X)$ is .shout[exogenous] if its variation is *unrelated* to other factors that affect the dependent variable $(Y)$

- An independent variable $(X)$ is .shout[endogenous] if its variation is *related* to other factors that affect the dependent variable $(Y)$
]

.pull-right[

.center[
![](https://www.dropbox.com/s/v5vwsadw5vs448t/causality.jpg?raw=1)
]
]

---

# Identification Problem: Endogeneity II

.pull-left[

- An independent variable $(X)$ is .shout[exogenous] if its variation is *unrelated* to other factors that affect the dependent variable $(Y)$

]

.pull-right[

```{r, fig.retina=3, outwidth="90%", fig.align="center"}
library(ggdag)
#theme_set(theme_dag_blank())
library(gridExtra)

h1 <- dagify(Y ~ X,
             coords=list(
               x=c(X=1,Y=2),
               y=c(X=1,Y=1)
               )) %>% ggdag_parents("Y", stylized = FALSE) + 
  labs(title = "X causes Y") + theme_dag_blank()+theme(legend.position = "none")

h2 <- dagify(Y ~ X+Z,
             coords=list(
               x=c(Z=1,X=1,Y=2),
               y=c(Z=2,X=1,Y=1.5)
             )) %>% ggdag_parents("Y", stylized = FALSE) +
  labs(title = "X and Z (independently) cause Y") + theme_void()+theme(legend.position = "none")

grid.arrange(h1, h2, ncol=1)

```

]

---

# Identification Problem: Endogeneity III

.pull-left[

- An independent variable $(X)$ is .shout[endogenous] if its variation is *related* to other factors that affect the dependent variable $(Y)$

]

.pull-right[

```{r, fig.retina=3, outwidth="90%", fig.align="center"}
h3 <- dagify(Y ~ X+Z, X ~Z,
             coords=list(
               x=c(X=1, Z=1.5, Y=2),
               y=c(X=1, Z=2, Y=1)
             )) %>% ggdag_parents("Y", stylized = FALSE) +
  labs(title = "Z causes X and Y") + theme_void()+theme(legend.position = "none")

h4 <- dagify(Y ~ Z, Z ~ X,
             coords=list(
               x=c(X=1, Z=1.5, Y=2),
               y=c(X=1, Z=1, Y=1)
               )) %>% ggdag_parents("Y", stylized = FALSE) + 
  labs(title = "X Causes Y Indirectly Through Z") + theme_void()+theme(legend.position = "none")

h5 <- dagify(Y ~ X+Z, Z ~X,
             coords=list(
               x=c(X=1, Z=1.5, Y=2),
               y=c(X=1, Z=0.5, Y=1)
             )) %>% ggdag_parents("Y", stylized = FALSE) +
  labs(title = "X Causes Y (Directly and Via Z)") + theme_void()+theme(legend.position = "none")

grid.arrange(h3, h4, h5, nrow=3)

```

]

---

# Identification Problem: Endogeneity IV

.pull-left[

- Confusingly, these terms mean something different in econometrics than in the theoretical economics models where you have heard them before

- In *Theoretical* models:
    - "Exogenous": a parameter determined outside of the model and taken as given
    - "Endogenous": a variable whose value is determined by the model

]

.pull-right[

.center[
![:scale 90%](https://www.dropbox.com/s/qvr240t5j6t3arm/optimize.jpeg?raw=1)
]
]

---

# Identification Problem: Endogeneity V

.pull-left[

.content-box-green[
.green[**Example 1**:] In a classic supply and demand model:

- Exogenous parameters: income, prices of other goods, cost, technology
- Endogenous variables: equilibrium price, equilibrium quantity
]
]

.pull-right[

.center[
![:scale 90%](https://www.dropbox.com/s/qvr240t5j6t3arm/optimize.jpeg?raw=1)
]
]

---

# Identification Problem: Endogeneity VI

.pull-left[

.content-box-green[

.green[**Example 2**:] In a consumer optimization model:

- Exogenous parameters: market prices, income, utility function 
- Endogenous variables: utility-maximizing bundle of goods
]
]

.pull-right[

.center[
![:scale 90%](https://www.dropbox.com/s/qvr240t5j6t3arm/optimize.jpeg?raw=1)
]
]

---

# Inference Problem: Randomness

.pull-left[
- Data is random due to **natural sampling variation**

    - Taking one sample of a population will yield slightly different information than another sample of the same population

- Common in statistics, *easy to fix*

- **Inferential Statistics**: making claims about a wider population using sample data

    - We use common tools and techniques to deal with randomness
]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/bsdtuddzjouwzr1/sampling.jpg?raw=1)
]
]

---

class: inverse, center, middle

# *Identifying* Causal Effects: Random Controlled Trials

---

# Random Controlled Trials (RCTs) I

.pull-left[

- The *ideal* way to demonstrate causation is through a .hl[randomized controlled trial (RCT)] or "random experiment"
    - *Randomly* assign experimental units (e.g. people, firms, etc.) into groups
    - **Treatment group(s)** get a (kind of) treatment
    - **Control group** gets no treatment 
    - Compare results of treatment and control groups to observe the **average treatment effect (ATE)**
- .hl[We will understand "causality" to mean the ATE from an ideal RCT]
]

.pull-right[

.center[
![:scale 90%](https://www.dropbox.com/s/ninee93he1zgcon/groupsplit.jpeg?raw=1)
]
]

---

# Random Controlled Trials (RCTs) II

.center[
![:scale 75%](https://www.dropbox.com/s/wqjisnpxaoaa2yy/rct.png?raw=1)

Classic (simplified) procedure of a randomized control trial (RCT) from medicine
]

---

# Random Controlled Trials (RCTs) III

.pull-left[

- .hl[Random assignment] to groups ensures that the *only* differences between members of the treatment(s) and control groups is *receiving treatment or not*

]

.pull-right[

.pull-left[
.center[
![:scale 80%](https://www.dropbox.com/s/vcqojr6k058e8hl/3apples.jpg?raw=1)
Treatment Group
]
]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/xvjphzlwechgip6/3oranges.jpg?raw=1)
Control Group
]
]
]

---

# Random Controlled Trials (RCTs) IV

.pull-left[

- .hl[Random assignment] to groups ensures that the *only* differences between members of the treatment(s) and control groups is *receiving treatment or not*

- .hl[Selection bias:] (pre-existing) differences between members of treatment and control groups *other* than treatment, that affect the outcome

]

.pull-right[

.pull-left[
.center[
![:scale 80%](https://www.dropbox.com/s/vcqojr6k058e8hl/3apples.jpg?raw=1)
Treatment Group
]
]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/xvjphzlwechgip6/3oranges.jpg?raw=1)
Control Group

(Selection Bias)
]
]
]

---

class: inverse, center, middle

# Some Theory By Example

---

background-image: url(https://www.dropbox.com/s/zsx3pa4m51p82dj/twopaths.jpg?raw=1)
background-size: cover

---

# Example: The Effect of Having Health Insurance I

.pull-left[

.content-box-green[
.green[**Example:**] What is the effect of having health insurance on health outcomes?

- National Health Interview Survey (NHIS) asks "Would you say your health in general is excellent, very good, good, fair, or poor?"

- **Outcome variable $(Y)$**: Index of health (1-poor to 5-excellent) in a sample of married NHIS respondents in 2009 who may or may not have health insurance

- **Treatment $(X)$**: Having health insurance  (vs. not)

]
]

.pull-right[

![](https://www.dropbox.com/s/h9c24dls3cisip1/healthcare.jpg?raw=1)
]

---

# Example: The Effect of Having Health Insurance II

.center[
![:scale 42%](https://www.dropbox.com/s/5lpltvdfim6elr6/econometricshealthnocaption.png?raw=1)]

.source[Angrist, Joshua & Jorn-Steffen Pischke, 2015, *Mostly Harmless Econometrics*]
---

# Example: The Effect of Having Health Insurance III

- $Y$: outcome variable (health index score, 1-5)

- $Y_i$: health score of an individual $i$

- Individual $i$ has a choice, leading to one of two outcomes:

    - $Y^0_i$: individual $i$ has *not* purchased health insurance

    - $Y^1_i$: individual $i$ has purchased health insurance

- $Y^1_i-Y^0_i$: causal effect for individual $i$ of purchasing health insurance

---

# Example: A Hypothetical Comparison

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J^0=3$ | $Y_M^0=5$ | 
| $Y_J^1=4$ | $Y_M^1=5$ |
]

--

.right-column[
- John will choose to buy health insurance

- Maria will choose to not buy health insurance
]

---

# Example: A Hypothetical Comparison

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J^0=3$ | $Y_M^0=5$ | 
| $Y_J^1=4$ | $Y_M^1=5$ |
| $Y_J^1-Y_J^0=1$ | $Y_M^1-Y_M^0=0$ |
]

.right-column[
- John will choose to buy health insurance

- Maria will choose to not buy health insurance

- Health insurance improves John's score by 1, has no effect on Maria's score
]

---

# Example: A Hypothetical Comparison

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J^0=3$ | $Y_M^0=5$ | 
| $Y_J^1=4$ | $Y_M^1=5$ |
| $Y_J^1-Y_J^0=1$ | $Y_M^1-Y_M^0=0$ |
| $Y_J=(Y_J^1)=4$ | $Y_M=(Y_M^0)=5$ |

]

.right-column[
- John will choose to buy health insurance

- Maria will choose to not buy health insurance

- Health insurance improves John's score by 1, has no effect on Maria's score

- Note, all we can observe in the data are their health outcomes *after* they have chosen (not) to buy health insurance: $Y_J=4$ and $Y_M=5$
]

---

# Example: A Hypothetical Comparison

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J^0=3$ | $Y_M^0=5$ | 
| $Y_J^1=4$ | $Y_M^1=5$ |
| $Y_J^1-Y_J^0=1$ | $Y_M^1-Y_M^0=0$ |
| $Y_J=(Y_J^1)=4$ | $Y_M=(Y_M^0)=5$ |

]

.right-column[
- John will choose to buy health insurance

- Maria will choose to not buy health insurance

- Health insurance improves John's score by 1, has no effect on Maria's score

- Note, all we can observe in the data are their health outcomes *after* they have chosen (not) to buy health insurance: $Y_J=4$ and $Y_M=5$

- .whisper[*Observed* difference] between John and Maria: $$Y_J-Y_M=-1$$
]

---

# Counterfactuals

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J=4$ | $Y_M=5$ |

This is all the data we *actually* observe
]

.right-column[
- Observed difference between John and Maria: 
$$Y_J-Y_M=\underbrace{Y^1_J-Y^0_M}_{=-1}$$

- Recall:
    - John has bought health insurance $(Y^1_J)$
    - Maria has not bought insurance $(Y^0_M)$

- We don't see the .shout[counterfactuals]: 
    - John's score *without* insurance
    - Maria score *with* insurance

]

---

# Counterfactuals

.left-column[

| John      | Maria     |
|-----------|-----------|
| ![](https://www.dropbox.com/s/5vh181yeg0p91t7/coldman2.jpg?raw=1) | ![](https://www.dropbox.com/s/7qa0gtuy29us4p2/femalejogger2.jpg?raw=1) |
| $Y_J=4$ | $Y_M=5$ |

This is all the data we *actually* observe

]

.right-column[

- Observed difference between John and Maria: 
$$Y_J-Y_M=\underbrace{Y^1_J-Y^0_M}_{=-1}$$

- Algebra trick: add and subtract $Y^0_J$ to equation

$$\begin{align*}
Y_j-Y_M=\underbrace{Y^1_J-Y^0_J}_{=1}+\underbrace{Y^0_J-Y^0_M}_{=-2}
\end{align*}$$

- $Y^1_J-Y^0_J=1$: Causal effect for John of buying insurance
- $Y^0_J-Y^0_M=-2$: Difference between John & Maria pre-treatment, .hl["selection bias"]

]

---

# Example II

.pull-left[

$$Y^0_J-Y^0_M \neq 0$$

- .hl[Selection bias:] (pre-existing) differences between members of treatment and control groups *other* than treatment, that affect the outcome
    - i.e. John and Maria *start out* with very *different* health scores before either decides to buy insurance or not ("recieve treatment" or not)

]

.pull-right[

.pull-left[
.center[
![:scale 80%](https://www.dropbox.com/s/vcqojr6k058e8hl/3apples.jpg?raw=1)

John (Treated)
]
]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/xvjphzlwechgip6/3oranges.jpg?raw=1)

Maria (Control)
]
]
]

---

# Example: Thinking about the Data

.pull-left[
```{r, eval = F, results="asis"}
library("tibble")
tribble(
  ~i, ~Insured, ~Not_Insured, ~Diff,
  "John", 4, 3, 1,
  "Maria", 5, 5, 0,
  "AVERAGE", 4.5, 4.0, 0.5
) %>%
  knitr::kable(., format="html")
```

.center[**Ideal** (but impossible) Data]

| Individual | Insured | Not Insured | Diff |
|---|---------|-------------|------|
| John | 4.0 | 3.0 | 1.0 |
| Maria | 5.0 | 5.0 | 0.0 | 
| **Average** | 4.5 | 4.0 | **0.5** |

- .hl[Individual treatment effect] (for individual $i$):
$$TE_i=Y_i^1-Y_i^0$$
- .hl[*Average* treatment effect]:
$$ATE=\frac{1}{n}\sum^n_{i=1}Y_i^1-Y_i^0$$

]

--

.pull-right[

.center[**Actual** (observed) Data]

| Individual | Insured | Not Insured | Diff |
|---|---------|-------------|------|
| John | 4.0 | ? | ? |
| Maria | ? | 5.0 | ? | 
| **Average** | ? | ? | ? |

- We never get to see each person's **counterfactual** state to compare and calculate ITEs or ATE

]

---

# Example: Thinking about the Data

.pull-left[

- Basic comparisons tell us *something* about outcomes, but not **ATE** (the causal effect we're looking for)

$$\text{Difference in Group Outcomes} = ATE + \text{Selection Bias}$$

- Selection bias: difference in average $Y^0_i$ between groups pre-treatment

- $Y^0_i$ includes *everything* about person $i$ relevant to health *except* treatment (insurance) status
    - Age, sex, height, weight, climate, smoker, exercise, diet, etc. 
    - Imagine a world where *nobody* has insurance (treatment), who would have highest health scores?
]

.pull-right[

.center[**Actual** (observed) Data]

| Individual | Insured | Not Insured | Diff |
|---|---------|-------------|------|
| John | 4.0 | ? | ? |
| Maria | ? | 5.0 | ? | 
| **Average** | ? | ? | ? |

]

---

# Random Assignment: The Silver Bullet

.pull-left[

- If treatment is .hl[randomly assigned] for a large sample, it eliminates selection bias!

- Treatment and control groups differ *on average* by nothing *except* treatment status

- Creates *ceterus paribus* conditions in economics: groups are identical *on average* (holding constant age, sex, height, etc.)

]

.pull-right[

.pull-left[
.center[
![](https://www.dropbox.com/s/vcqojr6k058e8hl/3apples.jpg?raw=1)
Treatment Group
]
]

.pull-right[
.center[
![](https://www.dropbox.com/s/vcqojr6k058e8hl/3apples.jpg?raw=1)
Control Group
]
]
]

---

class: inverse, center, middle
# Quasi-Experiments

---

# The Quest for Causal Effects I

.pull-left[
- RCTs are considered the "gold standard" for causal claims

- But society is not our laboratory (probably a good thing!)

- We can rarely conduct experiments to get data
]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/j0kwzqb2ptusckt/madlaboratory.jpg?raw=1)
]
]

---

# The Quest for Causal Effects II

.pull-left[
- Instead, we often rely on .whisper[observational data]

- This data is *not random*!

- Must take extra care in forming an .shout[identification strategy]

- To make good claims about causation in society, we must get clever! 
]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/3j2yflalbv2ylo1/fieldobserver.jpg?raw=1)
]
]

---

# Natural Experiments

.pull-left[

- Economists often resort to searching for .shout[natural experiments]

- Some events beyond our control occur that separate *otherwise similar* entities into a "treatment" group and a "control" group that we can compare

- e.g. natural disasters, U.S. State laws, military draft 
]

.pull-right[
.center[
![](https://www.dropbox.com/s/2yiww124mnvhj8l/nskorea2.png?raw=1)
]
]

---

# The First Natural Experiment

.left-column[
.center[
![:scale 100%](https://www.dropbox.com/s/kc1yx6q6iao26mo/jonsnow.jpg?raw=1)

Jon Snow
]
]

.right-column[

]

---

# The First Natural Experiment

.left-column[
.center[
![:scale 80%](https://www.dropbox.com/s/w5jy5an586l7ep7/johnsnow.jpg?raw=1)

1813-1858
]
]

.right-column[

- John Snow utilized the [first famous natural experiment](http://blogs.plos.org/publichealth/2013/03/11/john-snow-the-first-epidemiologist/) to establish the foundations of epidemiology and the germ theory of disease

- Water pumps with sources *downstream* of a sewage dump in the Thames river spread cholera while water pumps with sources *upstream* did not

.center[
![:scale 60%](https://www.dropbox.com/s/s8ar35t7sjklf4k/choleramap.jpg?raw=1)
]
]

---

# Famous Natural Experiments

- **Oregon Health Insurance Experiment**: Oregon used lottery to grant Medicare access to 10,000 people, showing access to Medicaid increased use of health services, lowered debt, etc. relative to those not on Medicaid
- **Angrist (1990)** finds that lifetime earnings of (random) drafted Vietnam veterans is 15% lower than non-veterans
- **Card & Kreuger (1994)** find that minimum wage hike in fast-food restaurants on NJ side of border had no disemployment effects relative to restaurants on PA side of border during the same period
- **Acemoglu, Johnson, and Robinson (2001)** find that inclusive institutions lead to higher economic development than extractive institutions, determined by a colony's disease environment in 1500
- We will look at some of these in greater detail throughout the course
- [A great list, with explanations is here](http://economicspsychologypolicy.blogspot.com/2015/06/list-of-19-natural-experiments.html)

---

# For the Next Few Classes

.center[
![:scale 45%](https://www.dropbox.com/s/k1xtcebidkonen9/r.png?raw=1)
]