### 2019-08-30 content/reading/02-reading.Rmd
Fixed reading pages for classes 1 & 2, added numbers to schedule; added R4DS to bib

---
title: "1.2 The Quest for Causality - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-book"></i> @Bailey[Ch. 1]

Bailey dives right in with the basic linear regression model

$$Y=\beta_0+\beta_1 X$$

and also begins with some demonstrations of the model. Don't worry about this just yet. What is more important right now is the **identification problem** (*endogeneity*) and the **inference problem** (*randomness*) and how these challenges motivate a lot of what we are going to do.