### 2019-09-03 content/class/03-class.Rmd

---
title: "1.3 Meet R - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 3, 2019*

# Overview

Today we begin the long slog to your mastery of `R`.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/02-slides.html)

# Appendix

## Installing R and R Studio

1. [Install R](http://www.r-project.org/) from CRAN^[The Comprehensive R Archive Network] by clicking "Download R" (or the CRAN link under Downloads on the left menu). This will take you to a mirrors page, where you can select a location in the U.S. and download a copy of R
2. [Install R Studio (Desktop Version)](https://www.rstudio.com/products/rstudio/download/), choose the "Free" option

(This will also be posted on the [Reference](/reference) page.)

## R Packages

Packages come from multiple sources.

The polished, publicly released versions are found on [CRAN](https://cran.r-project.org/). When installing a package available on CRAN, it is sufficient simply to tell R the following:^[Note the plural `s` on `packages`, and the quotes around the "package name"]

```{r, eval=F}
install.packages("packagename") 
```

Other packages, which may be in various developmental states (including perfectly functional!) are often hosted on [GitHub](http://github.com) before they make their way onto CRAN. Simply telling R `install.packages("packagename")` will fail to find it (as R only looks in CRAN for packages), so you must use another package called `devtools`^[Which you will need to install first if you (probably) don't already have it!] to install packages directly from Github:^[Note the `::` allows you to use the function `install_github()` from the `devtools` package without having to first load the devtools package with `library(devtools)`.]

```{r, eval=F}
devtools::install_github("username/packagename") 
```

For example, to install Hadley Wickham's package `r4ds` from its Github page [https://github.com/hadley/r4ds](https://github.com/hadley/r4ds), we would type:

```{r, eval=F}
devtools::install_github("hadley/r4ds")
```

To **use** a package, you need to ensure it is loaded to your workspace (you only need to do this once)^[When we learn how to write `R Markdown` documents, . with `library("package_name")`.^[Quotes are not necessary this time.]

## Getting Help for R

For specific functions or commands, you can simply type:

```{r, eval=F}
?functionname()

# example
?mean()
```

This will display a help page specific to that function in the Viewer pane. R functions and packages are extremely well-documented; help pages [normally ](http://socviz.co/assets/ch-09-read-a-help-page.png)^[This useful guide comes from Kieran Healy's excellent (free online!) [book on Data Visualization](http://socviz.co/).] a short description of the function, arguments and options (as well as their default values), and several examples or vignettes to demonstrate usage of the function.

Additionally, you can turn to the community.

## Other Useful Commands to Know

One of the best/worst things about `R` is that it is a language, meaning there are multiple ways that you can accomplish the same task.

### Creating Vectors

We know vectors can be created with the `c()` command (and stored with `<-`), but there are other shortcuts to combine objects into a vector, particularly numeric data:

- `:` creates a series of integers

```{r}
1:5 # create a vector of 1 through 5

12:17 # create a vector of 12 through 17
```

- `seq(from = , to = , by = )` creates a numeric sequence, and is not restricted to integers

```{r}
seq(from = 1, to = 10, by = 2) # sequence from 1 to 10, by 2s

# note you do not need to fully write out the name of each argument, just the input!

seq(32.5,40,1.5) # sequence from 32.5 to 40, by 1.5
```

- `rep(., times =)`^[`.` is a placeholder here.] repeats an element a specified number of times

```{r}
rep(2, times = 4) # repeat "2" four times

rep(2, 4) # does the same thing

# the thing repeated could itself be a vector
rep(c(1,4,7), 3) # repeat the vector "1, 4, 7" three times
```

- We can combine these:

```{r}

# combine (the sequence of 4 to 8 by 2's repeated three times) and 1 and 5

c(rep(seq(4,8,2),3),1,5)
```

## Style Guide for Coding

We want to maximize *human*-readability of code, not just machine-readability. I try to follow [Hadley Wickham's style guide](http://adv-r.had.co.nz/Style.html) for all of my code, including that in this class.

*You will not be graded on the style of your code.* But now's the best time to learn best practices (while you don't know any alternatives!) to save yourself and your potential colleagues (including your future self) from unnecessary frustration.