### 2019-09-03 content/reading/03-reading.Rmd

---
title: "1.3 Meet R - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-external-link-square-alt"></i>@R4DS[Ch. 1] 1^[[https://r4ds.had.co.nz/introduction.html](https://r4ds.had.co.nz/introduction.html)]

Now that we start working with R, we will