### 2019-09-03 static/r/03-r-practice.Rmd

---
title: "1.3 Meet R - R Practice"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Set Up

Before we begin, start a new file with \texttt{File} $\rightarrow$ \texttt{New File} $\rightarrow$ \texttt{R Script}. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

## Creating Objects

### 1. Create a vector called "me" with two objects, your first name, and your last name. Then call the vector to inspect it. Confirm it is a character class vector.

### 2. Create a vector called "x" with all the even integers from 2 to 10.

### 3. Find the mean of x with `mean()`

### 4. Now take the following pdf of random variable $Y$:

\begin{table}[h!]
\centering
  \begin{tabular}{ll}
  $y_i$ & $p_i$\\ \hline
  2 & 0.50\\
  4 & 0.25\\
  6 & 0.25\\ \hline
  \end{tabular}
\end{table}

Calculate the standard deviation "manually" using our table method. You can look at the source code of Lecture 4 for my example.

### a. Creating two vectors, one called $y.i$ and one called $p.i$, with the data above.

### b. Merge them into a data frame called $rv$ with `data.frame(y.i,p.i)`. Call `rv` to inspect it.

### c. Find the expected value of $Y$ by taking the sum of each value of `y.i` multiplied by `p.i` with the `sum()` command.

### d. Creating a new column in `rv` called `deviations`, where you subtract the mean from each `y.i` value. Call `rv` again to make sure it's now there.

### e. Create another column in `rv` called `devsq`, where you square the deviations from part d. Call `rv` again to make sure it's now there.

### f. Now add another column in `rv` called `weighteddevsq`, where you multiply the squared deviations in part e. by the associated probability `p.i`. Call `rv` again to make sure it's now there.

### g. Finally, take the sum of `weighteddevsq` to get variance. Square root this to get standard deviation.

### 5. The mean height of adults is 65 inches, with a standard deviation of 4 inches. Use the normal distribution to find the probabilities of the following scenarios:

### a. Find the probability of someone being *at least* 60 inches tall using `pnorm()`.

### b. Find the probability of someone being *at most* 60 inches tall.

### c. Find the probability of someone being between 61 and 69 inches tall. Why is this number familiar?

### d. Find the probability of someone being between 57 and 73 inches tall. Why is this number familiar?

## Playing with a Data Set

For the following questions, use the \texttt{diamonds} dataset, included as part of `ggplot2`.

### 6. Install `ggplot2`

### 7. Load `ggplot2` with the `library()` command

### 8. Get the structure of the diamonds data.frame. What are the different variables and what kind of data does each contain?

### 9. Get summary statistics for `carat`, `depth`, `table`, and `price`

### 10. `color`, `cut`, and `clarity` are categorical variables (factors). Use the `table()` command to generate frequency tables for each.

### 11. Now rerun the `summary()` command on the entire data frame

### 12. Plot a histogram of price.

### 13. Plot a boxplot of price by diamond color.

## Execute your R Script

Save the R Script you created at the beginning and (hopefully) have been pasting all of your valid commands to. This creates a *.R* file wherever you choose to save it to. Now looking at the file in the upper left pane of *R Studio* look for the button in the upper right corner that says **Run**. Sit back and watch `R` redo everything you've carefully worked on, all at once.