### 2019-09-03 static/r/03-r-practice.Rmd
Update class 3 materials

---
title: "1.3 Meet R - R Practice"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

# Creating Objects

## 1.

a. Create a vector called `me` with two objects, your first name, and your last name.
b. Call the vector to inspect it.
c. Confirm it is a character class vector.

## 2. Use R's help functions to determine what the `paste()` function does. Then paste together your name from `me`.

## 3. Create a vector called `my_vector` with all the even integers from 2 to 10.

## 4. Find the mean of x with `mean()`.

## 5. Add up all the numbers from 18 to 763, then get the mean

\clearpage

# Playing with Data

For the following questions, we will the `diamonds` dataset, included as part of `ggplot2`.

### 6. Install `ggplot2`.

### 7. Load `ggplot2` with the `library()` command.

### 8. Get the `str`ucture of the diamonds data frame. What are the different variables and what kind of data does each contain?

### 9. Get summary statistics separately for `carat`, `depth`, `table`, and `price`.

### 10. `color`, `cut`, and `clarity` are categorical variables (factors). Use the `table()` command to generate frequency tables for each.

### 11. Now rerun the `summary()` command on the entire data frame.

### 12. Now look only at (subset) the first 4 diamonds in the dataset.

### 13. Now look only at (subset) the third and seventh diamond in the dataset.

### 14. Now look only at (subset) the second column of the dataset.

### 15. Do this again, but look using the `$` to pull up the second column by name.

### 16. Now look only at diamonds that have a `carat` greater than or equal to 1.

### 17. Now look only at diamonds that have a VVS1 `clarity`.

### 17. Now look only at dimaonds that have a color of E, F, I, and J.

### 18. Now look only at diamonds that have a `carat` greater than or equal to 1 and a VVS1 `clarity`.

### 19. Get the average price of diamonds in question 18.^[Hints: use your subset command as an argument to the mean function. You will not need a comma here because you are looking for a single row.]

### 20. What is the highest price for a diamond with a 1.0 carat, D color, and VVS1 clarity?

# Execute your R Script

Save the R Script you created at the beginning and (hopefully) have been pasting all of your valid commands to. This creates a *.R* file wherever you choose to save it to. Now looking at the file in the upper left pane of *R Studio* look for the button in the upper right corner that says **Run**. Sit back and watch `R` redo everything you've carefully worked on, all at once.