### 2019-09-03 content/reading/03-reading.Rmd
Fix reading and tweak slides for class 3

---
title: "1.3 Meet R - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-external-link-square-alt"></i>Ch.1^[http://r4ds.had.co.nz/introduction.html] in @R4DS[Ch. 1]

Now that we start working with R, you should consider this book to be your primary reference for R-related questions. We will broadly cover the first few chapters in order over the next 2-3 class periods.