### 2019-09-05 content/class/04-class.Rmd
Add all content for class 1.4

---
title: "1.3 Meet R - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 3, 2019*

# Overview

We may spend at maximum 10 minutes at the beginning of class continuing to work on last class' R practice questions. I will put up the answer key after class. Today we start the fun stuff - data visualization. We will cover how to build plots with the package `ggplot2` (part of the `tidyverse`): I will lecture for the first half (again, it will be a lot of information that you can refer back to as needed over the semester) and you will practice making plots.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/04-slides.html)

# Practice Problems

Today you will be working on [R practice problems](/r/04-r-practice.html).

# Statistics and Software Survey

I will post "answers" after both sections have turned in their surveys.

# Additional Useful Information