### 2019-09-05 content/r/04-r-practice.Rmd

---
title: "1.4 Data Visualization with ggplot2 - R Practice"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```
- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/r/04-r-practice.pdf)

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

# Exploring the Data

## 1. We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

## 2. Let's get a quick look at `gapminder` to see what we're dealing with.

a. Get the `str`ucture of the `gapminder` data. 
b. What variables are there?
c. Look at the `head` of the dataset to get an idea of what the data looks like. 
d. Get `summary` statistics of all variables.

# Simple Plots in Base R

## 3. Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

## 4. Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

## 5. Now make it a *boxplot* by `continent`.^[Hint: use formula notation with`~`.]

## 6. Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

# Plots with `ggplot2`

## 7. Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

## 8. Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

## 9. For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

## 10. Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `color.`

## 11. Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

## 12. Redo your plot from 10 for `lifeExp` instead of `gdpPercap`.

## 13. Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

## 14. Add some color by mapping `continent` to `color` in your `aes`thetics.

## 15. Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

## 16. Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

## 17. Now add an `aes`thetic to your `point`s to map `pop` to `size`.

## 18. Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

## 19. Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

## 20. Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

## 21. Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

## 22. Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 16. (You might want to take a look at your new dataframe to make sure it worked first!)

## 23. Try this again for the *whole* world, but just for observations in the year 2000.