### 2019-09-05 static/r/03-r-practice-answers.Rmd

---
title: "1.3 Meet R - R Practice (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

# Creating Objects

## 1. Work on the following parts:

a. Create a vector called `me` with two objects, your first name, and your last name.

```{r}
me <- c("Ryan", "Safner")
```

b. Call the vector to inspect it.

```{r}
me
```

c. Confirm it is a character class vector.

```{r}
class(me)
```

## 2. Use R's help functions to determine what the `paste()` function does. Then paste together your first name and last name.

```{r}
?paste() # or help(paste)
 # paste is a function that combines (concatenates) multiple string objects into a single object
paste("Ryan", "Safner")

# note you can choose how to separate string objects with the "sep" argument
# for example
paste("Ryan", "Safner", sep="") # no separation
paste("Ryan", "Safner", sep=" ") # separate with a space " " (the default)
paste("Ryan", "Safner", sep="_") # separate with underscore
```
## 3. Create a vector called `my_vector` with all the even integers from 2 to 10.

```{r}
my_vector <- c(2,4,6,8,10)

# verify it worked
my_vector
# alternatively, you can use the sequence function, seq()
# see the Class page for more about this function
my_vector <- seq(from = 2, # starting integer
                 to = 10, # ending integer
                 by = 2) # by 2's

# you can shorten it by not including the names of the arguments:
my_vector <- seq(2,10,2)

# verify it worked
my_vector
```

## 4. Find the mean of `my_vector` with `mean()`.

```{r}
mean(my_vector)
```

## 5. Take all the integers from 18 to 763,^[Hint: use the `:` operator to create a sequence from a starting number to an ending number] then get the mean.

```{r}

# create a sequence of integers by 1 with starting_number:ending_number
# see Class 3 page for more

# you can do this all at once without making an object
mean(18:763)

# alternatively you can save this as a vector and run the mean on it
vec1 <- 18:763

mean(vec1)
```

\clearpage

# Playing with Data

For the following questions, we will use the `diamonds` dataset, included as part of `ggplot2`.

### 6. Install `ggplot2`.

```{r, eval=F}
install.packages("ggplot2") # note the s and the quotes
```

### 7. Load `ggplot2` with the `library()` command.

```{r}
library("ggplot2") # quotes not necessary, but can be used
```

### 8. Get the `str`ucture of the diamonds data frame. What are the different variables and what kind of data does each contain?

```{r}
str(diamonds)

# We have
# - carat: a number
# - cut: an ordered factor
# - color: an ordered factor
# - clarity: an ordered factor
# - depth: a number
# - table: a number
# - price: an integer
# - x: a number
# - y: a number
# - z: a number
```

### 9. Get summary statistics separately for `carat`, `depth`, `table`, and `price`.

```{r}
summary(diamonds$carat)
summary(diamonds$depth)
summary(diamonds$table)
summary(diamonds$price)
```

### 10. `color`, `cut`, and `clarity` are categorical variables (factors). Use the `table()` command to generate frequency tables for each.

```{r}
table(diamonds$color)
table(diamonds$cut)
table(diamonds$clarity)
```

### 11. Now rerun the `summary()` command on the entire data frame.

```{r}
summary(diamonds)
```

### 12. Now look only at (subset) the first 4 diamonds in the dataset.

```{r}
# remember, dataframes are indexed by: df[row#s, column#s]
diamonds[1:4,] # select first through fourth rows, all columns

# alternatively
diamonds[c(1,2,3,4),] # using a vector-approach
```

### 13. Now look only at (subset) the third and seventh diamond in the dataset.

```{r}
diamonds[c(3,7),] # select 3rd and 7th row, all columns
```

### 14. Now look only at (subset) the second column of the dataset.

```{r}
diamonds[,2] # select all rows, 2nd column
```
### 15. Do this again, but look using the `$` to pull up the second column by name.

```{r}
# second column is called "cut"
diamonds$cut
```

### 16. Now look only at diamonds that have a `carat` greater than or equal to 1.

```{r}
# use the [square brackets] to subset, 
# first argument (rows) are chosen by conditional: 
# - choose diamonds based on their carat, and only carats >= 1
diamonds[diamonds$carat >= 1,] # select rows on condition, and all columns
```

### 17. Now look only at diamonds that have a VVS1 `clarity`.

```{r}
# we are testing for equality, so we need two ==
# we are selecting based on clarity, a character/factor, so we need quotes
diamonds[diamonds$clarity=="VVS1",] # select rows on condition, and all columns
```

### 18. Now look only at dimaonds that have a color of E, F, I, and J.

```{r}
# same idea as last problem, except now we want one of any of these 4 colors

# first (tedious) way, a series of checking equality and using "OR"s (|) 
diamonds[diamonds$color=="E" | diamonds$color=="F" | diamonds$color=="I" | diamonds$color=="J",] # select rows on condition, and all columns

# second (better) way, using group membership operator (%in%) and list the elements as a vector
diamonds[diamonds$color %in% c("E","F","I","J"),] # select rows on condition, and all columns 
```

### 19. Now look only at diamonds that have a `carat` greater than or equal to 1 and a VVS1 `clarity`.

```{r}
# testing for two conditions (AND)
diamonds[diamonds$carat>=1 & diamonds$clarity=="VVS1",] # select rows on condition, and all columns 
```

### 20. Get the average price of diamonds in question 18.^[Hints: use your subset command as an argument to the mean function. You will not need a comma here because you are looking for a single row.]

```{r}
# use command from last question as the argument to the mean function, 
## but be sure that you look at the price, specifically

mean(diamonds$price[diamonds$carat>=1 & diamonds$color=="D" & diamonds$clarity=="VVS1"])
```

### 21. What is the highest price for a diamond with a 1.0 carat, D color, and VVS1 clarity?

```{r}
max(diamonds$price[diamonds$carat>=1 & diamonds$color=="D" & diamonds$clarity=="VVS1"])
```

# Execute your R Script

Save the R Script you created at the beginning and (hopefully) have been pasting all of your valid commands to. This creates a *.R* file wherever you choose to save it to. Now looking at the file in the upper left pane of *R Studio* look for the button in the upper right corner that says **Run**. Sit back and watch `R` redo everything you've carefully worked on, all at once.

Your `.R` file should look something like this:

```{r, eval=F}

# 1 -------------------

## a 
me <- c("Ryan", "Safner")

## b
me

## c 
class(me)

# 2 ----------------

?paste() # or help(paste)
 # paste is a function that combines (concatenates) multiple string objects into a single object
paste("Ryan", "Safner")

# note you can choose how to separate string objects with the "sep" argument
# for example
paste("Ryan", "Safner", sep="") # no separation
paste("Ryan", "Safner", sep=" ") # separate with a space " " (the default)
paste("Ryan", "Safner", sep="_") # separate with underscore

# 3 ---------------

my_vector <- c(2,4,6,8,10)

# verify it worked
my_vector
# alternatively, you can use the sequence function, seq()
# see the Class page for more about this function
my_vector <- seq(from = 2, # starting integer
                 to = 10, # ending integer
                 by = 2) # by 2's

# you can shorten it by not including the names of the arguments:
my_vector <- seq(2,10,2)

# verify it worked
my_vector

# 4 -------------------

mean(my_vector)

# 5 -------------------

# create a sequence of integers by 1 with starting_number:ending_number
# see Class 3 page for more

# you can do this all at once without making an object
mean(18:763)

# alternatively you can save this as a vector and run the mean on it
vec1 <- 18:763

mean(vec1)

# 6 ------------------

install.packages("ggplot2") # note the s and the quotes

# 7 ------------------

library("ggplot2") # quotes not necessary, but can be used

# 8 ------------------

str(diamonds)

# We have
# - carat: a number
# - cut: an ordered factor
# - color: an ordered factor
# - clarity: an ordered factor
# - depth: a number
# - table: a number
# - price: an integer
# - x: a number
# - y: a number
# - z: a number

# 9 ------------------

summary(diamonds$carat)
summary(diamonds$depth)
summary(diamonds$table)
summary(diamonds$price)

# 10 ------------------

table(diamonds$color)
table(diamonds$cut)
table(diamonds$clarity)

# 11 ------------------

summary(diamonds)

# 12 ------------------

# remember, dataframes are indexed by: df[row#s, column#s]
diamonds[1:4,] # select first through fourth rows, all columns

# alternatively
diamonds[c(1,2,3,4),] # using a vector-approach

# 13 ------------------

diamonds[c(3,7),] # select 3rd and 7th row, all columns

# 14 ------------------
diamonds[,2] # select all rows, 2nd column

# 15 ------------------

# second column is called "cut"
diamonds$cut

# 16 -------------------

# use the [square brackets] to subset, 
# first argument (rows) are chosen by conditional: 
# - choose diamonds based on their carat, and only carats >= 1
diamonds[diamonds$carat >= 1,] # select rows on condition, and all columns

# 17 -------------------

# we are testing for equality, so we need two ==
# we are selecting based on clarity, a character/factor, so we need quotes
diamonds[diamonds$clarity=="VVS1",] # select rows on condition, and all columns

# 18 -------------------

# same idea as last problem, except now we want one of any of these 4 colors

# first (tedious) way, a series of checking equality and using "OR"s (|) 
diamonds[diamonds$color=="E" | diamonds$color=="F" | diamonds$color=="I" | diamonds$color=="J",] # select rows on condition, and all columns

# second (better) way, using group membership operator (%in%) and list the elements as a vector
diamonds[diamonds$color %in% c("E","F","I","J"),] # select rows on condition, and all columns

# 19 -------------------

# testing for two conditions (AND)
diamonds[diamonds$carat>=1 & diamonds$clarity=="VVS1",] # select rows on condition, and all columns

# 20 -------------------

# use command from last question as the argument to the mean function
## but be sure that you look at the price, specifically
mean(diamonds$price[diamonds$carat>=1 & diamonds$color=="D" & diamonds$clarity=="VVS1"])

# 21 -------------------

max(diamonds$price[diamonds$carat>=1 & diamonds$color=="D" & diamonds$clarity=="VVS1"])
```