### 2019-09-05 static/slides/04-slides.Rmd

---
title: "1.4: Data Visualization with ggplot2"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: solarized-dark #ocean rainbow #monokai rainbow solarized-light #tomorrow-night -blue -eighties zenburn
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=F,
                      warning=F)
library(tidyverse)

```

# Graphics and Statistics

- Admittedly, we still need to cover basic descriptive statistics and data fundamentals
    - continuous, discrete, cross-sectional, time series, panel data
    - mean, median, variance, standard deviation
    - random variables, distributions, PDFs, Z-scores
    - bargraphs, boxplots, histograms, scatterplots

- All of this is coming in 2 weeks as we return to statistics and econometric theory

- But let's start with the fun stuff right away, even if you don't fully know the *reasons*: .shout[data visualiation]

---

class: inverse, center, middle

# Plotting in Base R

---

# Plotting in Base R

- Base `R` is very powerful and intuitive to plot, but not very sexy

- Basic syntax for most types of plots:

```{r, eval=F}
plot_type(my_df$variable)
```

- If using multiple variables, you can avoid typing `$` by just typing the variable names and then in another argument to the plotting function, specify `data = my_df`

```{r, eval=F}
plot_type(my_df$variable1, my_df$variable2, data = my_df)
```

---

# Plotting in Base R: Histogram

.left-code[
- Using the built-in `mtcars` data, plotting a histogram of `mpg`

```{r base-hist, eval=F}
hist(mtcars$mpg)
```

]

.right-plot[
.center[
```{r, ref.label="base-hist", echo=F, fig.retina=3}
```
]
]

---

# Plotting in Base R: Boxplot

.left-code[
- Using the built-in `mtcars` data, plotting a boxplot of `mpg`

```{r base-boxplot, eval=F}
boxplot(mtcars$mpg)
```

]

.right-plot[
.center[
```{r, ref.label="base-boxplot", echo=F, fig.retina=3}
```
]
]

---

# Plotting in Base R: Boxplot by Category

.left-code[
- Using the built-in `mtcars` data, plotting a boxplot of `mpg` by `cyl`

```{r base-boxplot2, eval=F}
boxplot(mtcars$mpg~mtcars$cyl)
```

```{r, eval=F}
# second method
boxplot(mpg~cyl, data = mtcars)
```

- The `~` is part of `R`'s **"formula" notation**: 
    - Dependent variable goes to left
    - Independent variable(s) to right, separated with `+`'s
    - Think `y~x+z` means "`y` is explained by `x` and `z`"

]

.right-plot[
.center[
```{r, ref.label="base-boxplot2", echo=F, fig.retina=3}
```
]
]

---

# Plotting in Base R: Scatterplot

.left-code[
- Using the built-in `mtcars` data, plotting a scatterplot of `mpg` against `disp`

```{r base-scatter, eval=F}
plot(mtcars$mpg~mtcars$disp)
```

```{r, eval=F}
# second method
plot(mpg~disp, data = mtcars)
```

]

.right-plot[
.center[
```{r, ref.label="base-scatter", echo=F, fig.retina=3}
```
]
]

---

class: inverse, center, middle

# ggplot2 and the tidyverse

---

class: blank

background-image: url(https://www.dropbox.com/s/np8v4ihjotkckc1/tidyverse1.png?raw=1)
background-size: cover

---

# The tidyverse

> "The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

- Largely (but not only) created by Hadley Wickham

- We will look at this much more extensively next week!

- This "flavor" of `R` will make your coding life *so much easier!*

![](https://www.dropbox.com/s/81ijxswbwqor0up/tidyverse.png?raw=1)

---

# ggplot

.pull-left[

- `ggplot2` is perhaps the most popular package in `R` and a core element of the `tidyverse`

- `gg` stands for a .whisper[grammar of graphics]

- Very powerful and beautiful graphics, very customizable and reproducible, but requires a bit of a learning curve

- All those "cool graphics" you've seen in the New York Times, fivethirtyeight, the Economist, Vox, etc use the grammar of graphics
]

.pull-right[
.center[
![:scale 65%](https://www.dropbox.com/s/4bxo0ac3gwwgd9y/hadleyplot2.png?raw=1)
]
]

---

# ggplot: All Your Figure are Belong to Us

.pull-left[
.center[
![:scale 80%](https://www.dropbox.com/s/4ij96m7h66i9o4p/538gg1.png?raw=1)

Source: [fivethirtyeight](https://fivethirtyeight.com/features/fastest-airlines-fastest-airports/)
]
]
.pull-left[
.center[
![:scale 80%](https://www.dropbox.com/s/2bkklsrsl0ep902/538gg.png?raw=1)

Source: [fivethirtyeight](https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/)
]
]

---

# Why Go gg?

.left-column[

.center[
![:scale 80%](https://www.dropbox.com/s/u34dakjyq4hkyov/hadley.jpg?raw=1)

Hadley Wickham

Chief Scientist, R Studio
]
]

.right-column[

> "The transferrable skills from ggplot2 are not the idiosyncracies of plotting syntax, but a powerful way of thinking about visualisation, as a way of **mapping between variables and the visual properties of geometric objects** that you can perceive."

]

.source[[http://disq.us/p/sv640d](http://disq.us/p/sv640d)]

---

# The Grammar of Graphics (gg) I

.pull-left[

- Any graphic can be built from the same components:
    1. **Data** to be drawn from
    2. **Aesthetic mappings** from data to some visual marking
    3. **Geometric objects** that appear on the plot
    4. **Coordinates** to organize location of objects
    5. **Scales** define the range of values
    6. **Facets** group into subplots

- Not every plot needs *every* component, but all plots *must* have the first 3!
]

.pull-right[
.center[
![](https://www.dropbox.com/s/flw1dbvhssn1q23/ggplotlayers.png?raw=1)
]
]

---

# The Grammar of Graphics (gg) II

.pull-left[

- Any graphic can be built from the same components:
    1. `data` to be drawn from
    2. `aes`**thetic mappings** from data to some visual marking
    3. `geom`**etric objects** that appear on the plot
    4. `coord`**inates** to scale axes
    5. `scale`**s** define the range of values
    6. `facet`**s** group into subplots

- Not every plot needs *every* component, but all plots *must* have the first 3!
]

.pull-right[
.center[
![](https://www.dropbox.com/s/flw1dbvhssn1q23/ggplotlayers.png?raw=1)
]
]

---

# Example Data

- We'll use the `mpg` example dataset that is built-in with `ggplot2`

```{r}
head(mpg, n=10)
```

---

# The Grammar of Graphics (gg): Data

.left-code[
### Data
```{r, eval=F}
ggplot(data = mpg)
```

]

.right-plot[

**Data** is the source of our data. As part of the `tidyverse`, `ggplot2` requires data to be .whisper["tidy"]<sup>.red[1]</sup>:

1. Each variable forms a .hl[column]

2. Each observation forms a .hl[row]

3. Each observational unit forms a table

]

.footnote[<sup>.red[1]</sup> Data "tidyness" is the core element of all `tidyverse` packages. Much more on all of this next class.]

---

# The Grammar of Graphics (gg): Adding Layers

.left-code[
### Data
]

.right-plot[
- Add a layer with `+` at the end of a line (never at the beginning!)

- Style recommendation: start a new line after each `+` to improve legibility!

- We will build a plot layer-by-layer
]

---

# The Grammar of Graphics (gg): Aesthetics I

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

.center[
![](https://www.dropbox.com/s/d226uyjrru1sn36/ggplotaes.png?raw=1)
]
]

---

# The Grammar of Graphics (gg): Aesthetics II

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

- `displ`

- `hwy`

- `class`

]

---

# The Grammar of Graphics (gg): Aesthetics II

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

- `displ` → **x**

- `hwy` → **y**

- `class` → *shape*, *size*, **color**, etc.

]

---

# The Grammar of Graphics (gg): Aesthetics II

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

.center[
![](https://www.dropbox.com/s/e1nvb7x330c7ruk/aes2.png?raw=1)
]
]
---

# The Grammar of Graphics (gg): Aesthetics III

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

.center[
![](https://www.dropbox.com/s/f5msffqgvbs4nkb/aes.png?raw=1)
]

---

# The Grammar of Graphics (gg): Aesthetics IV

.left-code[
### Data
### Aesthetics
```{r, eval=F}
+ aes()
```
]

.right-plot[
**Aesthetics** map data to visual elements or parameters

```{r, eval=F}
aes(x = displ,
    y = hwy,
    color = class)
```
]

---

# The Grammar of Graphics (gg): Geoms I

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
**Geometric objects** displayed on the plot

.center[
```{r geom_demo, echo=FALSE, fig.retina=3, fig.height=6}
minimal_theme <- theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    text = element_text(family = "Fira Mono"),
    plot.background = element_rect(fill = "#FAFAFA", color = NA),
    panel.background = element_rect(fill = "#FAFAFA", color = NA)
  )

set.seed(20)
df_geom <- data_frame(y = rnorm(10), x = 1:10)

g_geom <- list()
g_geom$point <- ggplot(df_geom, aes(x, y)) + geom_point() + ggtitle("geom_point()")
g_geom$line <- ggplot(df_geom, aes(x, y)) + geom_line() + ggtitle("geom_line()")
g_geom$bar <- ggplot(df_geom, aes(x, y)) + geom_col() + ggtitle("geom_col()")
g_geom$boxplot <- ggplot(df_geom, aes(y = y)) + geom_boxplot() + ggtitle("geom_boxplot()")
g_geom$histogram <- ggplot(df_geom, aes(y)) + geom_histogram(binwidth = 1) + ggtitle("geom_histogram()")
g_geom$density <- ggplot(df_geom, aes(y)) + geom_density(fill = "grey40", alpha = 0.25) + ggtitle("geom_density()") + xlim(-4, 4)

g_geom <- map(g_geom, ~ . + minimal_theme)

cowplot::plot_grid(plotlist = g_geom)
```
]
]

---

# The Grammar of Graphics (gg): Geoms II

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
**Geometric objects** displayed on the plot

- What `geom`s you should use depends on what you want to show:

| Type | `geom` |
|------|--------|
| Point | `geom_point()` |
| Line  | `geom_line()`, `geom_path()`  |
| Bar   | `geom_bar()`, `geom_col()` |
| Histogram | `geom_histogram()` |
| Regression | `geom_smooth()` |
| Boxplot | `geom_boxplot()` |
| Text | `geom_text()` |
| Density | `geom_density()` |
]

---

# The Grammar of Graphics (gg): Geoms III

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
**Geometric objects** displayed on the plot

```{r, eval=F}
##  [1] "geom_abline"     "geom_area"       "geom_bar"        "geom_bin2d"     
##  [5] "geom_blank"      "geom_boxplot"    "geom_col"        "geom_contour"   
##  [9] "geom_count"      "geom_crossbar"   "geom_curve"      "geom_density"   
## [13] "geom_density_2d" "geom_density2d"  "geom_dotplot"    "geom_errorbar"  
## [17] "geom_errorbarh"  "geom_freqpoly"   "geom_hex"        "geom_histogram" 
## [21] "geom_hline"      "geom_jitter"     "geom_label"      "geom_line"      
## [25] "geom_linerange"  "geom_map"        "geom_path"       "geom_point"     
## [29] "geom_pointrange" "geom_polygon"    "geom_qq"         "geom_qq_line"   
## [33] "geom_quantile"   "geom_raster"     "geom_rect"       "geom_ribbon"    
## [37] "geom_rug"        "geom_segment"    "geom_sf"         "geom_sf_label"  
## [41] "geom_sf_text"    "geom_smooth"     "geom_spoke"      "geom_step"      
## [45] "geom_text"       "geom_tile"       "geom_violin"     "geom_vline"
```

See [http://ggplot2.tidyverse.org/reference](http://ggplot2.tidyverse.org/reference) for many more options

]

---

# The Grammar of Graphics (gg): Geoms IV

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
**Geometric objects** displayed on the plot

Or just start typing `geom_` in R Studio!

.center[
![](https://www.dropbox.com/s/xui9bxccztlqmyz/geoms.gif?raw=1)
]
]

---

# Let's Make a Plot!

.left-code[
```{r plot-1, eval=F, echo=T}
ggplot(data = mpg)
```
]

.right-plot[
```{r, ref.label="plot-1", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Make a Plot!

.left-code[
```{r plot-2, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ, #<<
        y = hwy) #<<
```
]

.right-plot[
```{r, ref.label="plot-2", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Make a Plot!

.left-code[
```{r plot-3, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point() #<<
```
]

.right-plot[
```{r, ref.label="plot-3", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Make a Plot!

.left-code[
```{r plot-4, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class)) #<<
```
]

.right-plot[
```{r, ref.label="plot-4", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Make a Plot!

.left-code[
```{r plot-5, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point()+
  geom_smooth() #<<
```
]

.right-plot[
```{r, ref.label="plot-5", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# More Geoms

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
`geom_*(aes, data, stat, position)`

- `data`: geoms can have their own data
    - has to map onto global coordinates

- `aes`: geoms can have their own aesthetics
    - inherits global aesthetics by default
    - different geoms have different available aesthetics
]

---

# Change Our Plot

.left-code[
```{r plot-5a, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+ #<<
  geom_smooth()
```
]

.right-plot[
```{r, ref.label="plot-5a", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# More Geoms II

.left-code[
### Data
### Aesthetics
### Geoms
```{r, eval=F}
+ geom_*()
```
]

.right-plot[
`geom_*(aes, data, stat, position)`

- `stat`: some geoms statistically transform data
    - `geom_histogram()` uses `stat_bin()` to group observations into bins

- `position`: some adjust location of objects
    - `dodge`, `stack`, `jitter`
]

---

# Let's Change Our Plot

.left-code[
```{r plot-box, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = class, #<<
      y = hwy)+ #<<
  geom_boxplot() #<<
```
]

.right-plot[
```{r, ref.label="plot-box", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Change Our Plot

.left-code[
```{r plot-6, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = class)+ #<<
  geom_bar() #<<
```
]

.right-plot[
```{r, ref.label="plot-6", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Change Our Plot

.left-code[
```{r plot-7, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = class,
      fill = drv)+ #<<
  geom_bar() 
```
]

.right-plot[
```{r, ref.label="plot-7", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Let's Change Our Plot

.left-code[
```{r plot-8, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = class,
      fill = drv)+ 
  geom_bar(position = "dodge") #<<
```
]

.right-plot[
```{r, ref.label="plot-8", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# Back to the Original

.left-code[
```{r plot-8a, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth() 
```
]

.right-plot[
```{r, ref.label="plot-8a", echo=F, fig.retina=3, fig.align="center"}
```

]

---

# The Grammar of Graphics (gg): Facets I

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
```{r, eval=F}
+ facet_wrap()

+ facet_grid()
```
]

.right-plot[

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year) #<<
```
]

---

# The Grammar of Graphics (gg): Facets II

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
```{r, eval=F}
+ facet_wrap()

+ facet_grid()
```
]

.right-plot[

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_grid(cyl~year) #<<
```
]

---

# The Grammar of Graphics (gg): Labels

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
### Labels
```{r, eval=F}
+ labs()
```
]

.right-plot[

```{r, fig.retina=3, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")#<<
```
]

---

# The Grammar of Graphics (gg): Scales

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
### Labels
### Scales
```{r, eval=F}
+ scale_*_*()
```
]

.right-plot[

`scale`+`_`+`<aes>`+`_`+`<type>`+`()`

- `<aes>`: parameter you want to adjust
- `<type`: type of parameter

- I want to change my discrete x-axis: `scale_x_discrete()`
- I want to change my continuous y-axis: `scale_y_continuous()`
- I want to rescale x-axis to log: `scale_x_log10()`
- I want to use a different color palette: `scale_fill_discrete()`, `scale_color_manual()`
]
---

# The Grammar of Graphics (gg): Scales

.left-code[

```{r scales, eval=F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()#<<
```
]

.right-plot[

```{r, echo=F, ref.label="scales", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
### Labels
### Scales
### Theme
```{r, eval=F}
+ theme_*()
```
]

.right-plot[

**Theme** changes appearance of plot decorations (things not mapped to data)

- Some themes that come with `ggplot2`:

- `+ theme_bw()`
- `+ theme_dark()`
- `+ theme_gray()`
- `+ theme_minimal()`
- `+ theme_light()`
- `+ theme_classic()`
]
---

# The Grammar of Graphics (gg): Themes

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
### Labels
### Scales
### Theme
```{r, eval=F}
+ theme_*()
```
]

.right-plot[

**Theme** changes appearance of plot decorations (things not mapped to data)

- Many parameters we could change

- Global options: `line`, `rect`, `text`, `title`
- `axis`: x-, y-, or other axis title, ticks, lines
- `legend`: plot legends for fill or color
- `panel`: actual plot area
- `plot`: whole image
- `strip`: facet labels
]

---

# The Grammar of Graphics (gg): Themes

.left-code[

```{r theme1, eval=F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_bw()#<<
```
]

.right-plot[

```{r, echo=F, ref.label="theme1", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes II

.left-code[

```{r theme2, eval=F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_minimal()#<<
```
]

.right-plot[

```{r, echo=F, ref.label="theme2", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes III

.left-code[

```{r theme3, eval=F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_minimal()+
  theme(text = element_text(family = "Fira Sans"))#<<
```
]

.right-plot[

```{r, echo=F, ref.label="theme3", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes III

.left-code[

```{r theme4, eval=F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_minimal()+
  theme(text = element_text(family = "Fira Sans"),
        legend.position="bottom") #<<
```
]

.right-plot[

```{r, echo=F, ref.label="theme4", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes (ggthemes)

.left-code[
### Data
### Aesthetics
### Geoms
### Facets
### Labels
### Scales
### Theme
```{r, eval=F}
+ theme_*()
```
]

.right-plot[

- `ggthemes` package adds some other nice themes

```{r}
# install if you don't have it
# install.packages("ggthemes")
library("ggthemes") # load package
```

]

---

# The Grammar of Graphics (gg): Themes IV

.left-code[

```{r theme5, eval=F}
library("ggthemes") #<<
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_economist()+ #<<
  theme(text = element_text(family = "Fira Sans"),
        legend.position="bottom") 
```
]

.right-plot[

```{r, echo=F, ref.label="theme5", fig.retina=3, fig.align="center"}

```
]

---

# The Grammar of Graphics (gg): Themes V

.left-code[

```{r theme6, eval=F}
library("ggthemes") #<<
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")+
  scale_color_viridis_d()+
  theme_fivethirtyeight()+ #<<
  theme(text = element_text(family = "Fira Sans"),
        legend.position="bottom") 
```
]

.right-plot[

```{r, echo=F, ref.label="theme6", fig.retina=3, fig.align="center"}

```
]

---

class: inverse, center, middle

# Some Troubleshooting

---

# Global vs. Local Aesthetics

- `aes()` can go in base (`data`) layer and/or in individual `geom()` layers
- All `geoms` will inherit global `aes` from `data` layer unless overridden

.pull-left[

```{r fig.retina=3, fig.height=3.5, fig.align="center"}
# ALL GEOMS will map data to colors
ggplot(data = mpg, aes(x = displ,
                       y = hwy,
                       color = class))+
  geom_point()+
  geom_smooth()
```

]

.pull-right[
```{r fig.retina=3, fig.height=3.5, fig.align="center"}
# ONLY points will map data to colors
ggplot(data = mpg, aes(x = displ,
                       y = hwy))+
  geom_point(aes(color = class))+
  geom_smooth()
```
]

---

# Mapped vs. Set Aesthetics

- `aes`thetics such as `size` and `color` can be mapped from data or set to a single value
- Map *inside* of `aes()`, set *outside* of `aes()`

.pull-left[

```{r fig.retina=3, fig.height=3.5, fig.align="center"}
# Point colors are mapped from class data
ggplot(data = mpg, aes(x = displ,
                       y = hwy))+
  geom_point(aes(color = class))+
  geom_smooth()
```
]

.pull-right[
```{r fig.retina=3, fig.height=3.5, fig.align="center"}
# Point colors are all set to blue
ggplot(data = mpg, aes(x = displ,
                       y = hwy))+
  geom_point(aes(), color = "red")+
  geom_smooth(aes(), color = "blue")
```
]

---

class: inverse, center, middle
# Saving Plots

---

# Saving Plots

---

# Go Crazy

.left-code[
```{r gap-anim, eval=F}
library("gapminder")
library("gganimate")
ggplot(gapminder) +
  aes(x = gdpPercap, y = lifeExp, size = pop, color = country) +
  geom_point() +
  guides(color = FALSE, size = FALSE) +
  scale_x_log10(
    breaks = c(10^3, 10^4, 10^5), 
    labels = c("$1k", "$10k", "$100k")) +
  scale_color_manual(values = gapminder::country_colors) +
  scale_size(range = c(0.5, 12)) +
  labs(
    x = "GDP per capita",
    y = "Life Expectancy",
    caption = "Source: Hans Rosling's gapminder.org") +
  theme_minimal(14, base_family = "Fira Sans") +
  theme(
    strip.text = element_text(size = 16, face = "bold"),
    panel.border = element_rect(fill = NA, color = "grey40"),
    panel.grid.minor = element_blank())+
  transition_states(year, 1, 0)+
  ggtitle("Income and Life Expectancy - {closest_state}")
```
]

.right-plot[
```{r ref.label="gap-anim", fig.retina=3, fig.height=5.5, fig.align="center", echo=F}
```
]

---

# Data Visualization and Graphic Design Principles

.pull-left[

- We will return to various graphics as we cover descriptive statistics and regression

- I hope to cover some basic principles of good graphic design for figures and plots
    - If not in class, I will make a page on the website, and/or a video
]

--

.pull-right[

Remember:

.center[
![](https://www.dropbox.com/s/noj9n3ohqpgq2pc/powerresponsibility.jpg?raw=1)
]
]

---

# Less is More

.pull-left[

.center[

"Shoot me"
![](https://www.dropbox.com/s/azh5zxzqkmyuxoh/badfigure.png?raw=1)
]

]

--

.pull-right[

.center[
Less is More:

![](http://stat545.com/img/less-is-more-darkhorse-analytics.gif)
]
]
---

# Try to Show One Trend Really Clearly

.center[
![](https://www.dropbox.com/s/gd9eisleu21lgga/goodfigure.png?raw=1)
]

.source[New York Times: ["How Stable Are Democracies? ‘Warning Signs Are Flashing Red’"](https://www.nytimes.com/2016/11/29/world/americas/western-liberal-democracy.html), Nov 29, 2016]

---

# Reference: R Studio Makes Great "Cheat Sheet"s!

.center[
![](https://www.dropbox.com/s/k7fe3fpe6sc5oq7/ggplot2cheat.png?raw=1)
]

.source[[RStudio: ggplot2 Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)]
---

# Reference

On `ggplot2`

- **R Studio's [ggplot2 Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)**
- **`ggplot2`'s website [reference section](https://ggplot2.tidyverse.org/reference/index.html)**
- STHDA's [be awesome in ggplot2](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization)
- r-statistic's [top 50 ggplot2 visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
- Hadley Wickham's \href{http://r4ds.had.co.nz/data-visualisation.html}{R for Data Science book chapter on ggplot2}

On data visualization

- **Kieran Healy's [Data Visualization: A Practical Guide](http://socviz.co/)**
- **Claus Wilke's [Fundamentals of Data Visualization](https://serialmentor.com/dataviz)**
- PolicyViz [Better Presentations](https://policyviz.com/) 
- Karl Broman's [How to Display Data Badly](https://www.biostat.wisc.edu/~kbroman/presentations/IowaState2013/graphs_combined.pdf)
- [I Want Hue](http://tools.medialab.sciences-po.fr/iwanthue/)