### 2019-09-06 static/slides/05-slides.Rmd

---
title: "1.5: Data Wrangling in the tidyverse"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: solarized-dark #ocean rainbow #monokai rainbow solarized-light #tomorrow-night -blue -eighties zenburn
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=F,
                      warning=F)
library(tidyverse)

```

# Data Wrangling

- Most real world data analysis transforming chaos into order
    - Data strewn from multiple sources
    - Missing data
    - Data not in readable form

---

# Workflow of a Data Scientist

.pull-left[
.center[.polaroid[![](https://www.dropbox.com/s/zke6m65wp3upouj/workflow.png?raw=1)]]
]

.pull-right[
- Consider Hadley Wickham's famous diagram from [R for Data Science](http://r4ds.had.co.nz)
1. .shoutcaps[Import] raw data from out there in the world
2. .shoutcaps[Tidy] it into a form that you can use to 
3. .shoutcaps[Explore] the data
4. .shoutcaps[Communicate] results to target audience
]

---

# Workflow of a Data Scientist

.pull-left[
.center[.polaroid[![](https://www.dropbox.com/s/zke6m65wp3upouj/workflow.png?raw=1)]]
]

.pull-right[

3. .shoutcaps[Explore] the data:
- .whisper[Transform] existing variables into more useful ones
- Make plots to .whisper[visualize] .onfire[relationships]
- Create .whisper[models] to analyze, predict, and interpret causal relationships

- 95% of the day-to-day work of data scientists, researchers, econometricians: ".hl[data wrangling]"
]

---

# Workflow of a Data Scientist

.pull-left[
.center[
![](https://www.dropbox.com/s/jd68ky17gb7pmu0/datawranglingnyt.png?raw=1)
]
]

.pull-right[
.center[
![](https://www.dropbox.com/s/8kzy823yubdv8gh/datawranglingnyt2.png?raw=1)
]
]

.center[
[New York Times](https://www.nytimes.com/2014/08/18/technology/for-big-data-scientists-hurdle-to-insights-is-janitor-work.html)
]

---

class: blank

background-image: url(https://www.dropbox.com/s/np8v4ihjotkckc1/tidyverse1.png?raw=1)
background-size: cover

---

# The tidyverse

> "The tidyverse is an opinionated  collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

- Core packages include ones we've discussed before: `ggplot2`, `dplyr`, `magrittr` among several others (`tidy` `readr`, `purr`, `forcats`, `stringr`)
- Loading any `tidyverse` package loads `magrittr` (so you can use `%>%`)
- Learn more at [tidyverse.org](tidyverse.org)

![](https://www.dropbox.com/s/81ijxswbwqor0up/tidyverse.png?raw=1)

---

# The tidyverse II

- Easiest to just load the core tidyverse all at once
    - Note loading the tidyverse is "noisy", it will spew a lot of messages
    - Hide them with `suppressPackageStartupMessages()` and insert the `library()` command inside the `()`

```{r, echo=FALSE}
detach("package:tidyverse", TRUE) # just to show start up message of tidyverse, let's unload it first
```

```{r, message=TRUE, warning=TRUE}
library("tidyverse")
```

---

# The tidyverse III

![](https://www.dropbox.com/s/9h1trto0iapnklx/tidyverse2.png?raw=1)

---

# Tidyverse Packages

- We will make **extensive** use of:

1. `dplyr` for data wrangling
2. `tidyr` for tidying data
3. `ggplot2` for plotting data

- Loading `tidyverse` makes **automatically** uses

4. `tibble` for better data frames
5. `magrittr` for "pipeable" code

- We might later look at

6. `readr` for importing data
7. `forcats` for working with factors
8. `stringr` for working with strings

---

# Tibble

.left-column[

.center[![](https://www.dropbox.com/s/u6ubqkng1ja5jwf/tibble.png?raw=1)]
]

.right-column[

- `tibble` converts all `data.frames` into a .whisper[friendlier] version called `tibbles`

]

---

# Tibble

.pull-left[

```{r, eval=F}
diamonds
```

<!--Can't fit in columns so artificially selecting to print only first 7 columns -->

```{r, echo=F} 
diamonds %>% 
  select(1:7)
```

]

.pull-right[

- Prints much nicer output in R
- Shows a bit of the `str`ucture of a dataframe automatically:
    - `nrow() x ncol()`
    - `<dbl>` is numeric ("double")
    - `<ord>` is an ordered factor
    - `<int>` is an integer
- Fundamental grammar of tidyverse: 
    1. start with a tibble 
    2. run a function on it
    3. output a new tibble
]

---

# Magrittr I

.right-column[

- The .hl[magrittr] package allows us to use the **"pipe" operator** (`%>%`)<sup>.red[1]</sup>
- `%>%` "pipes" the *output* of the *left* of the pipe *into* the *argument* of the *right* 
    - Running a function `f` on object `x` as `f(x)` becomes `x %>% f` in pipeable form
    - i.e. "take `x` and then run function `f` on it"
]

.left-column[

.center[![](https://www.dropbox.com/s/74srdjrw13mgx4x/magrittr.png?raw=1)]
]

.footnote[<sup>.red[1]</sup> Keyboard shortcuts in R Studio: `CTRL+Shift+M` (Windows) or `Cmd+Shift+M` (Mac)]

---

# Magrittr II

.right-column[

- With ordinary math functions, read from outside $\leftarrow$ (inside): 
$$g(f(x))$$

i.e. take `x` and perform function `f()` on `x` and then take that result and perform function `g()` on it
- With pipes, read operations from left $\rightarrow$ right:

```{r, eval=FALSE}
x %>% f %>% g
```

take `x` and then perform function `f` on it, then perform function `g` on that result

- So read `%>%` mentally as "and then"

]

.left-column[

.center[![](https://www.dropbox.com/s/74srdjrw13mgx4x/magrittr.png?raw=1)]
]

---

# Magrittr III

.right-column[

.content-box-green[
.green[**Example**]

$$ln(exp(x))$$

- First, exponentiate $x$, then take the natural log of that (resulting in just x)
- In pipes:

```{r, eval=F}
x %>% exp() %>% ln()
```
]

]

.left-column[

.center[![](https://www.dropbox.com/s/74srdjrw13mgx4x/magrittr.png?raw=1)]
]

---

# Magrittr III

.right-column[
.content-box-green[
.green[**Example**]

- Sequence: find keys, unlock car, drive to school, park 
- Using nested functions in pseudo-"code":

```{r, eval=F}
park(drive(start_car(find("keys")), to = "campus"))
```

- Using pipes:

```{r, eval=F}
find("keys") %>%
  start_car() %>%
  drive(to = "campus") %>%
  park()
```

]

]

.left-column[

.center[![](https://www.dropbox.com/s/74srdjrw13mgx4x/magrittr.png?raw=1)]
]

---

# Magrittr: Simple Example

```{r, echo=FALSE}
library("gapminder")
```

.pull-left[

```{r}
# look at top 6 rows
head(gapminder) 
```
]

.pull-right[
```{r}
# use the pipe instead 
gapminder %>% head()
```
]

---

# Magrittr: More Involved Example

- These two methods produce the same output (avg hwy mpg of Audi cars)

- Without the pipe

```{r, eval=F}
summarise(group_by(filter(mpg, manufacturer=="audi"), model), hwy_mean = mean(hwy))
```

- Using the pipe

```{r}
mpg %>%
  filter(manufacturer=="audi") %>%
  group_by(model) %>%
  summarise(hwy_mean = mean(hwy))

```

---

# Readr

.left-column[

.center[![](https://www.dropbox.com/s/cqv60z3cznsmr9a/readr.png?raw=1)]
]

.right-column[

- `Readr` lets you load common spreadsheet files (`.csv`, `.tsv`) with simple commands:
- `read_csv(path/to/my_data.csv)`<sup>.red[1]</sup>
    - `read_tsv(path/to/my_data.tsv)`
- Often this is enough, but there are many more customizations as necessary
- You can also export your data from R into a common spreadsheet file with:
- `write_csv(my_dataframe, path=path/to/file_name.csv)`<sup>.red[2]</sup> 
- Read more on the [tidyverse website](https://readr.tidyverse.org/) and the [Readr Cheatsheet](https://rawgit.com/rstudio/cheatsheets/master/data-import.pdf)
]

.footnote[<sup>.red[1]</sup> If you are using an R Project, this file path is in your main project folder, or any subfolders you decided to create, such as `data/my_data.csv`

<sup>.red[2]</sup> where `my_dataframe` is the name of the dataframe you want to export, and `file_name` is the name of the file you want to create.]

---

# Readxl and Haven: When Readr isn't Enough

.left-column[

.center[![](https://www.dropbox.com/s/rl1sj67y2m8fqy5/readxl.png?raw=1)]
.center[![](https://www.dropbox.com/s/ua6t5wz1se34vmo/haven.png?raw=1)]
]

.right-column[

- Some data types, particularly for other software programs like Excel, STATA, SAS, and SPSS are not `.csv` or `.txt` spreadsheets
- `readxl` has commands equivalent to `readr`'s for Excel data types:
    - `read_xls("path/to/my/data.xls")` for older Excel `.xls` files
    - `read_xlsx("path/to/my/data.xlsx")` for modern Excel `.xls` files
    - `write_xls(my_dataframe, path=path/to/file_name.xls)` or ``write_xlsx(my_dataframe, path=path/to/file_name.xlsx)` to export as `xls` or `xlsx` files
- `haven` has commands equivalent to `readr`'s for these data types:
    - `read_dta("path/to/my_data.dta")` for STATA `.dta` files
    - `read_sav("path/to/my_data.sav")` for SPSS `.sav` files
    - `read_sas("path/to/my_data.sas7bdat")` for SAS files 
]

---

# dplyr I

.left-column[

.center[![](https://www.dropbox.com/s/6ao0u4xc3w492ak/dplyr.png?raw=1)]
]

.right-column[

- `dplyr` uses much more efficient and intuitive commands<sup>.magenta[1]</sup> to manipulate dataframes
- Grammatically:
    - `Base R` is all about passively running functions on nouns: `function(object)`
    - `dplyr` is all about actively using verbs: `dplyr_verb(df, conditions)`<sup>.red[2]</sup>
- Three great features:

1. Allows use of `%>%` pipe operator
2. Input and output is always a dataframe<sup>.magenta[3]</sup>
3. Shows the output from a manipulation, but does not save/overwrite as an object unless explicitly assigned to an object

]

.footnote[<sup>.magenta[1]</sup> Particularly when combined with `magrittr`'s pipe (`%>%`) command!

<sup>.red[2]</sup> With the pipe, even simpler: `df %>% dplyr_verb(conditions)`

<sup>.red[3]</sup> Technically a `tibble` with `tidyverse`!]

---

# dplyr II

.left-column[

.center[![](https://www.dropbox.com/s/6ao0u4xc3w492ak/dplyr.png?raw=1)]
]

.right-column[

- Common `dplyr` functions:

| Verb | Does | 
|----------|------|
| `filter` | Keep only selected *observations* | 
| `select` | Keep only selected *variables* |
| `arrange` | Reorder rows (e.g. in numerical order) |
| `mutate` | Create new variables | 
| `recode` | Change a variable's values or categories/`factor` levels |
| `summarize` | Collapse data into summary statistics| 
| `group_by` | Perform any of the above functions by groups/categories |

]

---

# dplyr::filter

- `filter` keeps only selected **observations** (rows)

--

.left-code[

```{r, eval=F}
# look only at African observations
# syntax without the pipe
filter(gapminder, continent=="Africa")

# using the pipe

gapminder %>%
  filter(continent == "Africa")
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  filter(continent=="Africa")
```

]

---

# dplyr: saving and storing outputs I

- `dplyr` functions never modify their inputs (i.e. never overwrite the original `tibble`)
- If you want to save a result, use `<-` to assign it to a new `tibble`
- If assigned, you will not see the output until you call up the new `tibble` by name

--

.left-code[

```{r, eval = F}
# base syntax
africa <- filter(gapminder, continent=="Africa")

# using the pipe
africa <- gapminder %>%
  filter(continent == "Africa")

# look at new tibble
africa
```
]

.right-plot[
```{r, echo = F}
africa <- gapminder %>%
  filter(continent == "Africa")

# look at new tibble
africa
```
]

---

# dplyr: saving and storing outputs II

- If you want to *both* store and view the output at the same time, wrap the command in parentheses!

```{r}
(africa <- gapminder %>%
  filter(continent == "Africa"))
```

---

# dplyr: saving and storing outputs III

- If you were to assign the output to the original `tibble`, it would *overwrite* the original!

```{r, eval = F}
# base syntax
gapminder <- filter(gapminder, continent=="Africa")

# using the pipe
gapminder <- gapminder %>%
  filter(continent == "Africa")

# this overwrites gapminder!
```

---

# dplyr conditionals

- In many data wrangling contexts, you will want to select data .whisper[conditionally]
    - In computer code, you are selecting only observations for which a set of logical conditions are `TRUE`
- For example:
    - `>`, `<`: greater than, less than
    - `>=`, `<=`: greater than or equal to, less than or equal to
    - `==`, `!=`: is equal to, is not equal to
    - `%in%`: is a member of some defined set $(\in)$
    - `&`: AND<sup>1</sup>
    - `|`: OR

.footnote[<sup>1</sup> Commas also work in `dplyr`.]

---

# dplyr::filter with conditionals

--

.left-code[

```{r, eval=F}
# look only at African observations in 1997
gapminder %>%
  filter(continent == "Africa", year == 1997)
```
]

--

.right-plot[

```{r, echo=F}
gapminder %>%
filter(continent == "Africa", year == 1997)
```
]

---

# dplyr::filter with conditionals II

--

.left-code[

```{r, eval=F}
# look only at African observations or observations in 1997
gapminder %>%
  filter(continent == "Africa" | year == 1997)
```
]

--

.right-plot[

```{r, echo=F}
gapminder %>%
filter(continent == "Africa" | year == 1997)
```
]

---

# dplyr::filter with conditionals III

--

.left-code[

```{r, eval=F}
# look only at U.S. and U.K. observations in 2002
gapminder %>%
  filter(country %in% c("United States", "United Kingdom"), year == 2002)
```
]

--

.right-plot[

```{r, echo=F}
gapminder %>%
  filter(country %in% c("United States", "United Kingdom"), year == 2002)
```
]

---

# dplyr::arrange I

- `arrange` reorders **observations** (rows) in a logical order
    - e.g. alphabetical, numeric, small to large

--

.left-code[

```{r, eval=F}
# order by smallest to largest population
# syntax without the pipe
arrange(gapminder, pop)

# using the pipe

gapminder %>%
  arrange(pop)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  arrange(pop)
```

]

---

# dplyr::arrange II

- Break ties in the value of one variable with the values of additional variables

--

.left-code[

```{r, eval=F}
# order by year, with the smallest to largest population in each year
# syntax without the pipe
arrange(gapminder, year, pop)

# using the pipe

gapminder %>%
  arrange(year, pop)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  arrange(year, pop)
```

]

---

# dplyr::arrange III

- Use `desc()` to re-order in the opposite direction

--

.left-code[

```{r, eval=F}
# order by largest to smallest population
# syntax without the pipe
arrange(gapminder, desc(pop))

# using the pipe

gapminder %>%
  arrange(desc(pop))
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  arrange(desc(pop))
```

]

---

# dplyr::select I

- `select` keeps only selected **variables** (columns)
    - Don't need quotes around column names
--

.left-code[

```{r, eval=F}
# keep only country, year, and population variables
# syntax without the pipe
select(gapminder, country, year, pop)

# using the pipe

gapminder %>%
  select(country, year, pop)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  select(country, year, pop)
```

]

---

# dplyr::select II

- `select` "all except" by negating a variable with `-`

--

.left-code[

```{r, eval=F}
# keep all except gdpPercap
# syntax without the pipe
select(gapminder, -gdpPercap)

# using the pipe

gapminder %>%
  select(-gdpPercap)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  select(-gdpPercap)
```

]

---

# dplyr::select III

- `select` reorders the columns in the order you provide
    - sometimes useful to keep all variables, and drag one or a few  to the front, add `everything()` at the end

--

.left-code[

```{r, eval=F}
# keep all and move pop first
# syntax without the pipe
select(gapminder, pop, everything())

# using the pipe

gapminder %>%
  select(pop, everything())
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  select(pop, everything())
```

]

---

# dplyr::select IV

- `select` has a lot of helper functions, useful for when you have hundreds of variables
    - see `?select()` for a list

--

.pull-left[

```{r}
# keep all variables starting with "co"

gapminder %>%
  select(starts_with("co"))
```

]

--

.pull-right[

```{r}
# keep country and all variables containing "per"

gapminder %>%
  select(country, contains("per"))
```

]

---

# dplyr::rename

- `rename` changes the name of a variable (column)
    - Format: `new_name=old_name`
--

.left-code[

```{r, eval=F}
# rename gdpPercap to GDP
# syntax without the pipe
rename(gapminder, GDP = gdpPercap)

# using the pipe

gapminder %>%
  rename(GDP = gdpPercap)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  rename(GDP = gdpPercap)
```

]

---

# dplyr::mutate

- `mutate` creates a new variable (column)
    - always adds a new column at the end
    - general formula: `new_variable_name = operation`

---

# dplyr::mutate II

- Three major types of mutates:

1. Create a variable that is a specific value (often categorical)

--

.left-code[

```{r, eval=F}
# create variable "europe" if country is in Europe
# syntax without the pipe
mutate(gapminder, europe = ifelse(continent == "Europe", 
                                  yes = "In Europe", 
                                  no = "Not in Europe"))

# using the pipe

gapminder %>%
  mutate(europe = ifelse(continent == "Europe", 
                                  yes = "In Europe", 
                                  no = "Not in Europe"))
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  mutate(europe = ifelse(continent == "Europe", 
                                  yes = "In Europe", 
                                  no = "Not in Europe"))
```
]

---

# dplyr::mutate III

- Three major types of mutates:

1. Create a variable that is a specific value (often categorical)
2. Change an existing variable (often rescaling)

--

.left-code[

```{r, eval=F}
# create population in millions 
# syntax without the pipe
mutate(gapminder, pop_millions = pop / 1000000)

# using the pipe

gapminder %>%
  rename(pop_millions = pop / 1000000)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  mutate(pop_millions = pop / 1000000)
```
]

---

# dplyr::mutate IV

- Three major types of mutates:

1. Create a variable that is a specific value (often categorical)
2. Change an existing variable (often rescaling)
3. Create a variable based on other variables

--

.left-code[

```{r, eval=F}
# create GDP variable from gdpPercap and pop
# syntax without the pipe
mutate(gapminder, GDP = gdpPercap * pop)

# using the pipe

gapminder %>%
  rename(GDP = gdpPercap * pop)
```

]

--

.right-plot[

```{r, echo=F}
# syntax with the pipe 
gapminder %>%
  mutate(GDP = gdpPercap * pop)
```
]

---

# dplyr::mutate V

- Can create multiple new variables with commas:

--

```{r}
gapminder %>%
  mutate(GDP = gdpPercap * pop,
         pop_millions = pop / 1000000)
```

--

- `transmute` keeps *only* newly created variables (`select`s only the new `mutate`d variables)

```{r}
gapminder %>%
  transmute(GDP = gdpPercap * pop,
         pop_millions = pop / 1000000)
```

--

- Don't forget to assign the output to  new `tibble` if you want to "save" the new variables!

---

# Exporting Tables with knitr::kable

---

# dplyr::tally with group_by

- `tally` provides counts, best used with `group_by` for categorical data

```{r}
gapminder %>%
  tally
```

```{r}
gapminder %>%
  group_by(continent) %>%
  tally
```

---

# Advanced: tidyr and joining data

---

# tidyr: reshaping and tidying data

.left-column[

.center[![](https://www.dropbox.com/s/p6gmb4tds18ueyp/tidyr.png?raw=1)]
]

.right-column[

- `tidyr` helps clean data into more usable format for other `tidyverse` packages
- .hl["tidy" data].footnote[This is the core and eponym of the `tidyverse`: all associated packages and functions use or require this data format!] are (an opinionated view of) data where

1. Each .shout[variable] is in a .shout[column]
2. Each .whisper[observation] is a .whisper[row]
3. Each .onfire[value] is a .onfire[cell]

- Un-tidy data needs to be tidied into this format
    - You will spend less time fighting tools and more time on analysis!

]

---

# Tidying Data: `gather()` wide to long I

.pull-left[

```{r, echo=F}
ex_wide<-tribble(
  ~"Country", ~"2000", ~"2010",
  "United States", 140, 180,
  "Canada", 102, 98,
  "China", 111, 123
)
```

```{r}
ex_wide
```

]

.pull-right[

- Common source of "un-tidy" data: .l[Column headers are values, not variable names]
    - Column names are *values* of a `year` variable!
    - Each row represents *two* observations (one in 2000 and one in 2010)!
- We need to `gather()` these columns into a new pair of variables
    - set of columns that represent values, not variables (`2000` and `2010`)
    - `key`: name of variable whose values form the column names (we'll call it the `year`)
    - `value`: name of the variable whose values are spread over the cells (we'll call it number of `cases`)
]

---

# Tidying Data: `gather()` wide to long II

- `gather()` a wide data frame into a long data frame

.pull-left[

```{r}
ex_wide
```
]

.pull-right[

```{r}
ex_wide %>% gather("2000","2010",
              key = "year",
              value = "cases")
```

]

---

# Tidying Data: `spread()` long to wide I

.pull-left[

```{r, echo=F}
ex_long<-tribble(
  ~"Country", ~"Year", ~"Type", ~"Count",
  "United States", 2000, "Cases", 140,
  "United States", 2000, "Population", 300,
  "United States", 2010, "Cases", 180,
  "United States", 2010, "Population", 310,
  "Canada", 2000, "Cases", 102,
  "Canada", 2000, "Population", 110,
  "Canada", 2010, "Cases", 98,
  "Canada", 2010, "Population", 121,
  "China", 2000, "Cases", 111,
  "China", 2000, "Population", 1201,
  "China", 2010, "Cases", 123,
  "China", 2010, "Population", 1241
)

```

```{r}
ex_long
```

]

.pull-right[

- Common source of "un-tidy" data: .l[observations are scattered across multiple rows]
    - Each country has two rows per observation, one for `Cases` and one for `Population` (categorized by `type` of variable)
- We need to `spread()` these columns into a new pair of variables
    - `key`: column that contains variable names (here, the `type`)
    - `value`: column that contains values from multiple variables (here, the `count`)
]

---

# Tidying Data: `spread()` long to wide II

- `spread()` a long data frame into a wide data frame

.pull-left[

```{r}
ex_long
```
]

.pull-right[

```{r}
ex_long %>% spread(key = "Type",
                   value = "Count")
```

]

---

# Combining Datasets

- Often, data doesn't come from just one source, but several sources
- We can combine datasets into a single dataframe (tibble) using `dplyr` commands in several ways:
    1. `bind` dataframes together by row or by column
        - `bind_rows()` adds observations (rows) to existing dataset<sup>.magenta[1]</sup>
        - `bind_cols()` adds variables (columns) to existing dataset<sup>.magenta[2]</sup>
    2. `join` two existing dataframes by designating a variable(s) as a `key` to match rows by identical values of that `key`

.footnote[<sup>.magenta[1]</sup> Note the columns must be identical between the original dataset and the new observations

<sup>.magenta[1]</sup> Note the rows must be identical between original dataset and new variable]

---

# Two *Similar* Datasets

- Sometimes you want to add rows (observations) or columns (variables) that happen to match up perfectly
    - New observations contain all the same variables as existing data
    - OR
    - New variables contain all the same observations as existing data
- In this case, simply using `bind_*(old, new)` will work
    - `bind_columns(old, new)` adds columns from `new` to `old` dataframe
    - `bind_rows(old, new)` adds rows from `new` to `old` dataframe

---

# Two *Different* Datasets

.pull-left[

- For the following examples, consider the following two dataframes, `x` and `y`<sup>*</sup>
    - each has one unique variable, `x$x` and `y$y`
    - both have values for observations `1` and `2`
    - `x` has observation `3` which `y` does not have
    - `y` has observation `4` which `x` does not have
- We next consider the ways we can merge dataframes `x` and `y` into a single dataframe

]

.pull-right[

![](https://raw.githubusercontent.com/gadenbuie/tidyexplain/master/images/static/png/original-dfs.png)

]

.footnote[<sup>*</sup> Images on all following slides come from Gaden Buie's excellent [tidyexplain](https://github.com/gadenbuie/tidyexplain)]

---

# Inner-Join

.pull-left[

- Merge columns from `x` and `y` for which there are matching rows
    - Rows in `x` with no match in `y` (3) will be dropped
    - Rows in `y` with no match in `x` (4) will be dropped

]

.pull-right[

![](https://raw.githubusercontent.com/gadenbuie/tidyexplain/master/images/inner-join.gif)

]
---

# Left-Join

.pull-left[

- Start with all rows from `x` and add all columns from `y`
    - Rows in `x` with no match in `y` (3) will have `NA`s
    - Rows in `y` with no match in `x` (4) will be dropped

]

.pull-right[

![](https://raw.githubusercontent.com/gadenbuie/tidyexplain/master/images/left-join.gif)

]

---

# Right-Join

.pull-left[

- Start with all rows from `y` and add all columns from `x`
    - Rows in `y` with no match in `x` (4) will have `NA`s
    - Rows in `x` with no match in `y` (3) will be dropped

]

.pull-right[

![](https://raw.githubusercontent.com/gadenbuie/tidyexplain/master/images/right-join.gif)

]

---

# Full-Join

.pull-left[

- All rows and all columns from `x` and `y`
    - Rows that do not match (3 and 4) will have `NA`s

]

.pull-right[

![](https://raw.githubusercontent.com/gadenbuie/tidyexplain/master/images/full-join.gif)

]