### 2019-09-06 static/files/syllabus-pdf.tex
Add syllabus PDF

\documentclass{article}
\usepackage{booktabs, graphicx, hyperref, fontspec}
\usepackage{sectsty}
\allsectionsfont{\sffamily}
\usepackage[margin=1in]{geometry}
\hypersetup{
  colorlinks = true,
  urlcolor = cyan,
 }
 \providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{Fira Sans}

\begin{document}

\sffamily

\centerline{\Huge Econometrics}

\vspace{3 mm}

\centerline{\large Dr.~Ryan Safner}
\vspace{2 mm}
\centerline{\large \href{http://metricsf19.classes.ryansafner.com}{metricsF19.classes.ryansafner.com}}

\vspace{5 mm}

\begin{tabular}{@{}p{3.5in}p{3.5in}}           
\textbf{Course}: ECON 480  & \textbf{Email:}  \href{mailto:safner@hood.edu}{\nolinkurl{safner@hood.edu}}\\
\textbf{Room}: Rosenstock Trading Room & \textbf{Office:}  Rosenstock Hall 118\\
\textbf{Meets}: TuTh 2:00 P.M.--3:15 P.M. & \textbf{Hours:} TuTh 10:30--11:30 A.M.\\ 
\end{tabular}

\vspace{5 mm}

\hrule

\begin{quote}
``There are three kinds of lies: lies, damned lies, and statistics.'' --
Benjamin Disraeli
\end{quote}

\textbf{Econometrics} is the application of statistical tools to
quantify and measure economic relationships in the real world. It uses
real data to test economic hypotheses, quantitatively estimate causal
relationships between economic variables, and to make forecasts of
future events. The primary tool that economists use for empirical
analysis is ordinary least squares (OLS) linear regression, so the
majority of this course will focus on understanding, applying, and
extending OLS regressions.

I assume you have \emph{some} working knowledge of economics at the
intermediate level and some basic statistical tools.\footnote{The formal
  prerequisites for this course are \textbf{ECON 205} and \textbf{ECON
  206}; \textbf{ECMG 212} or \textbf{MATH 112}; and \textbf{ECON 305} or
  \textbf{ECON 306}} We will do some basic review of some necessary
statistics and probability at the beginning until everyone is
comfortable, before jumping right into regressions.

\hypertarget{course-objectives}{%
\section{Course objectives}\label{course-objectives}}

\textbf{By the end of this course,} you will:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  apply the models of microeconomics (constrained optimization and
  equilibrium) towards explaining real world behavior of individuals,
  firms, and governments
\item
  explore the effects of economic and political processes on market
  performance (competition, market prices, profits and losses, property
  rights, entrepreneurship, market power, market failures, public
  policy, government failures)
\item
  apply the economic way of thinking to real world issues in writing
\item
  understand how to evaluate statistical and empirical claims;
\item
  use the fundamental models of causal inference and research design;
\item
  gather, analyze, and communicate with real data in R.
\end{enumerate}

I am less concerned with forcing you to memorize and recite proofs of
statistical estimator properties, and more concerned with the
development of your intuitions and the ability to think critically as an
empirical social scientist---although this will require you to
demonstrate proficiency with some intermediate statistical and
mathematical tools.

Given these objectives, this course fulfills all three of the learning
outcomes for
\href{https://www.hood.edu/academics/departments/george-b-delaplaine-jr-school-business/student-learning-outcomes}{the
George B. Delaplaine, Jr.~School of Business} Economics B.A. program:

\begin{itemize}
\tightlist
\item
  Use quantitative tools and techniques in the preparation,
  interpretation, analysis and presentation of data and information for
  problem solving and decision making {[}\ldots{]}
\item
  Apply economic reasoning and models to understand and analyze problems
  of public policy {[}\ldots{]}
\item
  Demonstrate effective oral and written communications skills for
  personal and professional success{[}\ldots{]}
\end{itemize}

{Fair warning:} \textbf{Econometrics is hard.} \emph{It will be one of
the hardest economics courses that you will take, primarily due to the
mathematical content.} I will do my best to make this class intuitive
and helpful, if not interesting. If at any point you find yourself
struggling in this course for any reason, please come see me. Do not
suffer in silence! Coming to see me for help does not diminish my view
of you, in fact I will hold you in \emph{higher} regard for
understanding your own needs and taking charge of your own learning.
There are also a some fantastic resources on campus, such as the
\href{http://www.hood.edu/campus-services/academic-services/index.html}{Center
for Academic Achievement and Retention (CAAR)} and the
\href{http://www.hood.edu/library/\%7D\%7BBeneficial-Hodson\%20Library}{Beneficial-Hodson
Library}.

See my
\href{http://metricsf19.classes.ryansafner.com/reference\#tips}{tips for
success in this course}.

\hypertarget{course-materials}{%
\section{Course materials}\label{course-materials}}

You can find all course materials at my academic website dedicated for
(all sections of) this course
\href{http://microf19.classes.ryansafner.com}{microf19.classes.ryansafner.com}.

My lecture slides will be shared with you, and serve as your primary
resource, but our main ``textbook'' below is \textbf{recommended} as the
next best resource and will be available from the campus bookstore.

\hypertarget{books}{%
\subsection{Books}\label{books}}

The following book is \textbf{required}\footnote{You are not
  \emph{obligated} to buy it, I just \textbf{strongly recommend} it in
  the sense that you will still have access to all data and assignments
  without possessing the book. But this is a course where you really
  will want to understand the derivations or get additional context
  beyond just my slides\ldots{}} and will be available from the campus
bookstore.

\begin{itemize}
\tightlist
\item
  Bailey, Michael A, 2019, \emph{Real Econometrics}, New York: Oxford
  University Press, 2\textsuperscript{nd} ed.
\end{itemize}

You are welcome to purchase the book by other means (e.g.~Amazon,
half.com, etc). I have no financial stake in requiring you to purchase
this book. The (cheaper) 1st edition is sufficient, but makes
significantly less use of \texttt{R} (in favor of \texttt{STATA}).

The following two books are \textbf{recommended}, and are free
online\footnote{You can purchase a hard copy of the first one if you
  really want.}:

\begin{itemize}
\tightlist
\item
  Grolemund, Garrett and Hadley Wickham,
  \href{https://r4ds.had.co.nz/}{\emph{R For Data Science}}
\item
  Ismay, Chester and Albert Y Kim,
  \href{https://moderndive.com/}{\emph{Modern Dive: Statistical
  Inference Via Data Science}}
\end{itemize}

The first book is the number one resource for using \texttt{R} and
\texttt{tidyverse}, and is written for beginners. I still look at it
\emph{frequently.} The second is another great reference for using
\texttt{tidyverse} in the context of basic statistics.

\hypertarget{software}{%
\subsection{Software}\label{software}}

You are \textbf{strongly recommended} to download copies of
\href{https://www.r-project.org/}{\texttt{R}} and
\href{http://www.rstudio.com}{\texttt{R\ Studio}} on your own computers.
These software packages are available on all computers in the trading
room, and you will have access to them during the week to work on
assignments.

We will also have a shared class workspace in
\href{http://rstudio.cloud/}{RStudio.cloud} that runs a full instance of
R Studio in your web browser (so no need to install anything!) will let
you access files and assignments.

\hypertarget{articles}{%
\subsection{Articles}\label{articles}}

Throughout the course, I will post both required and supplemental
(non-required) readings that enrich your understanding for each topic.
Check Blackboard \emph{frequently} for announcements and updates to
assignments, readings, and grades.

\hypertarget{assignments-and-grades}{%
\section{Assignments and grades}\label{assignments-and-grades}}

\textbf{You can find descriptions} for all the assignments on the
\href{http://metricsf19.classes.ryansafner.com/assignments/}{assignments
page}.

\begin{center}

\begin{tabular}{lll}
\toprule
 & Assignment & Percent\\
\midrule
1 & Research Project & 30\%\\
n & Problem sets (Average) & 25\%\\
1 & Midterm & 20\%\\
1 & Final & 25\%\\
\bottomrule
\end{tabular}
\end{center}

All grades are based on the following traditional scale:

\begin{center}

\begin{tabular}{llll}
\toprule
Grade & Range & Grade1 & Range1\\
\midrule
A & 93–100\% & C & 73–76\%\\
A− & 90–92\% & C− & 70–72\%\\
B+ & 87–89\% & D+ & 67–69\%\\
B & 83–86\% & D & 63–66\%\\
B− & 80–82\% & D− & 60–62\%\\
\addlinespace
C+ & 77–79\% & F & < 60\%\\
\bottomrule
\end{tabular}
\end{center}

These grades are firm cutoffs, but I do of course round upwards
(\(\geq 0.5\)) for final grades. A necessary reminder, as an academic, I
am not in the business of \emph{giving} out grades, I merely report the
grade that you \emph{earn}. I will not alter your grade unless you
provide a reasonable argument that I am in error (which does happen from
time to time).

\hypertarget{policies-and-expectations}{%
\section{Policies and Expectations}\label{policies-and-expectations}}

This syllabus is a contract between you, the student, and me, your
instructor. It has been carefully and deliberately thought out\footnote{A
  syllabus can and will be used as a legal document for disputes tried
  at a court of law. Ask me how I know.}, and I will uphold my end of
the agreement and expect you to uphold yours.

In the language of game theory, this syllabus is my commitment device. I
am a very understanding person, and I know that exceptions to rules
often need to be made for students. However, to be \emph{fair} to
\emph{all} students the syllabus artificially constrains my ability to
make exceptions at a whim for anyone. This prevents clever students from
exploiting my congenial personality at everyone else's expense. Please
read and familiarize yourself with the course policies and expectations
of you. Chances are, if you have a question, it is answered herein.

\hypertarget{attendance-and-participation}{%
\subsubsection{Attendance and
Participation}\label{attendance-and-participation}}

I expect you to attend class and to come having already done the reading
assigned for that day. I will remind you in class and possibly through
Blackboard or email which readings I want you to read for the next
class. You are all adults and I will treat you as such. I do not take
attendance, nor do I grade formally for participation but I strongly
recommended you attend class and participate for your sake and the sake
of your classmates. If you are too distracted or are not prepared to
learn, I suggest you stay home, where you can check Facebook more
efficiently. I reserve the right to boost the final grades of students
that I believe have made consistent, quality contributions above and
beyond their peers in class conversations by up to 2.5 points.

\hypertarget{devices}{%
\subsubsection{Devices}\label{devices}}

You are allowed to have and use laptops and tablets in the classroom. I
will not stop you, but I strongly discourage you from using these to
take notes (see Tips for Success). As a courtesy to myself and to
others, do not use your phone in class. I reserve the right to embarrass
you in front of everyone if you do so.

\hypertarget{absences-and-make-ups}{%
\subsubsection{Absences and Make-Ups}\label{absences-and-make-ups}}

You generally do \emph{not} need to let me know if you are unable to
make class, \emph{unless} it is on the day of an exam. It will however,
be your responsibility to acquire the notes from a classmate for any
missed classes. If you are unable to attend an exam for a legitimate
reason (e.g.~sports/club events, traveling, illness, family issues),
please notify me at least \emph{one week} in advance, and we will
schedule a make-up exam date. If you are ill or otherwise unable to
attend on the day of the exam, contact me ASAP to make arrangements.
Failure to do so, including desperate attempts to make arrangements only
\emph{after} the absence will result in a grade of 0 and little
sympathy. I reserve the right to re-weight other assignments for
students who I believe are legitimately unable to complete a particular
assignment.

\hypertarget{late-assignments}{%
\subsubsection{Late Assignments}\label{late-assignments}}

I will accept late assignments, but will subtract a specified amount of
points as a penalty. See individual assignment descriptions for the
amount of points taken off (as it varies by assignment). If an answer
key is posted before you turn in your assignment, the maximum grade you
can earn is an 80. Even if it is the last week of the semester, I
encourage you to turn in late work: some points are better than no
points!

\hypertarget{grading}{%
\subsubsection{Grading}\label{grading}}

I will try my best to post grades on Blackboard's Grading Center and
return graded assignments to you within about one week of you turning
them in. There will be exceptions. Where applicable, I will post answer
keys once I know most homeworks are turned in (see Late Assignments
above for penalties). Blackboard's Grading Center is the place to look
for your most up-to-date grades. You will also be given an Excel
spreadsheet template where you can calculate your overall grade and
forecast ``what if'' scenarios.

\hypertarget{email-accounts}{%
\subsubsection{Email Accounts}\label{email-accounts}}

Students must regularly monitor their Hood email accounts to receive
important college information, including messages related to this class.
Email through the Blackboard system is my main method of communicating
announcements and deadlines regarding your assignments. \textbf{Please
do not reply to any automated Blackboard emails - I may not recieve
it!}. My Hood email
(\href{mailto:safner@hood.edu}{\nolinkurl{safner@hood.edu}}) is the best
means of contacting me. I will do my best to respond within 24 hours. If
I do not reply within 48 hours, do not take it personally, and
\emph{feel free to send a follow up email} in the very likely event that
I genuinely did not see your original message.

\hypertarget{office-hours}{%
\subsubsection{Office Hours}\label{office-hours}}

I am generally in my office Monday-Thursday during ``normal business
hours.'' You are always welcome to walk-in and chat about class,
college, careers, or anything at all. Please do try to use the official
office hours stated at the head of the syllabus if possible. If you need
to meet at a different time, I request that you send me an email or let
me know after class so I know when to expect you. If you want to go over
material from class, please have \emph{specific} questions you want help
with. I am not in the business of giving private lectures (particularly
if you missed class without a valid excuse).

Watch the excellent and accurate video explaining office hours (on
website syllabus page):

\hypertarget{enrollment}{%
\subsubsection{Enrollment}\label{enrollment}}

Students are responsible for verifying their enrollment in this class.
The last day to add or drop this class with no penalty is
\textbf{Tuesday, September 4}. Be aware of
\href{https://www.hood.edu/offices-services/registrars-office/academic-calendar}{important
dates}.

\hypertarget{honor-code}{%
\subsubsection{Honor Code}\label{honor-code}}

Hood College has an Academic Honor Code which requires all members of
this community to maintain the highest standards of academic honesty and
integrity. Cheating, plagiarism, lying, and stealing are all prohibited.
All violations of the Honor Code are taken seriously, will be reported
to appropriate authority, and may result in severe penalties, including
expulsion from the college. See
\href{http://hood.smartcatalogiq.com/en/2016-2017/Catalog/The-Spirit-of-Hood/The-Academic-Honor-Code-and-Code-of-Conduct}{here}
for more detailed information.

\hypertarget{van-halen-and-mms}{%
\subsubsection{Van Halen and M\&Ms}\label{van-halen-and-mms}}

When you have completed reading the syllabus, email me a picture of the
band Van Halen and a picture of a bowl of M\&Ms.~If you do this
\emph{before} the date of the first exam, you will get 3 bonus points on
the exam. Yes, this is real.

\hypertarget{accessibility-equity-and-accommodations}{%
\subsubsection{Accessibility, Equity, and
Accommodations}\label{accessibility-equity-and-accommodations}}

College courses can, and should, be challenging and bring you out of
your comfort zone in a safe and equitable environment. If, however, you
feel at any point in the semester that certain assignments or aspects of
the course will be disproportionately uncomfortable or burdensome for
you due to any factor beyond your control, please come see me or email
me. I am a very understanding person and am happy to work out a solution
together. I reserve the right to modify and reweight assignments at my
sole discretion for students that I belive would legitimately be at a
disadvantage, through no fault of their own, to complete them as
described.

If you are unable to afford required textbooks or other resources for
any reason, come see me and we can find a solution that works for you.

This course is intended to be accessible for all students, including
those with mental, physical, or cognitive disabilities, illness,
injuries, impairments, or any other condition that tends to negatively
affect one's equal access to education. If at any point in the term, you
find yourself not able to fully access the space, content, and
experience of this course, you are welcome to contact me to discuss your
specific needs. I also encourage you to contact the
\href{https://www.hood.edu/academics/josephine-steiner-center-academic-achievement-retention/accessibility-services}{Office
of Accessibility Services} (301-696-3421). If you have a diagnosis or
history of accommodations in high school or previous postsecondary
institutions, Accessibility Services can help you document your needs
and create an accommodation plan. By making a plan through Accessibility
Services, you can ensure appropriate accommodations without disclosing
your condition or diagnosis to course instructors.

\hypertarget{schedule}{%
\section{Schedule}\label{schedule}}

\textbf{You can find a full schedule} with resources for each class
meeting on the
\href{http://metricsF19.classes.ryansafner.com/schedule/}{course website
schedule page}.

\end{document}