### 2019-09-10 content/class/05-class.Rmd
Link to all lesson 5 content!

---
title: "1.3 Meet R - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 10, 2019*

# Overview

Today we will cover the heart of `tidyverse` and use it for ["data wrangling"](https://en.wikipedia.org/wiki/Data_wrangling). Today will again be a lot of content thrown at you, so you can look back at this as a reference all semester. Then we will do more practice problems.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/05-slides.html)

# Practice Problems

Today you will be working on [R practice problems](/r/05-r-practice.html). Check back later for solutions.

# Statistics and Software Survey

I have posted ["answers"](/static/homeworks/00-preliminary-survey-answers.html) to the statistics survey.

# Additional Useful Information

Check back here!