### 2019-09-11 static/r/04-r-practice-answers-pdf.Rmd
Fixed link to survey answers on class 5, added R practice 1.4 answers and links

---
title: "1.4 Data Visualization with ggplot2 - R Practice (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

# Exploring the Data

## 1. We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

```{r}
# first time only
# install.packages("gapminder")

# load gapminder
library(gapminder)

# get help
?gapminder
```

## 2. Let's get a quick look at `gapminder` to see what we're dealing with.

a. Get the `str`ucture of the `gapminder` data.

```{r}
str(gapminder)
```

b. What variables are there?

```{r}
# - country: a factor
# - continent: a factor
# - year: an integer
# - lifeExp: a number
# - gdpPercap: a number
```

c. Look at the `head` of the dataset to get an idea of what the data looks like.

```{r}
head(gapminder)
```

d. Get `summary` statistics of all variables.

```{r}
summary(gapminder)
```

# Simple Plots in Base R

## 3. Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

```{r}
hist(gapminder$gdpPercap)
```

## 4. Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

```{r}
boxplot(gapminder$gdpPercap)
```

## 5. Now make it a *boxplot* by `continent`.^[Hint: use formula notation with`~`.]

```{r}
boxplot(gapminder$gdpPercap~gapminder$continent)

# alternate method
# boxplot(gdpPercap~continent, data = gapminder)
```

## 6. Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

```{r}
plot(gapminder$lifeExp~gapminder$gdpPercap)

# alternate method
# boxplot(lifeExp~gdpPercap, data = gapminder)
```

# Plots with `ggplot2`

## 7. Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

```{r}
# install if you don't have
# install.packages("ggplot2")

# load ggplot2 
library(ggplot2)
```

## 8. Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

```{r}
ggplot(data = gapminder,
       aes(x = continent))+
  geom_bar()
```

## 9. For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap))+
  geom_histogram()
```

## 10. Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `fill`.^[In general, `color` refers to the outside borders of a `geom` (except points), `fill` is the interior of an object.]

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           fill = continent))+
  geom_histogram()
```

## 11. Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           fill = continent))+
  geom_density(alpha=0.4)
```

## 12. Redo your plot from 11 for `lifeExp` instead of `gdpPercap`.

```{r}
ggplot(data = gapminder,
       aes(x = lifeExp,
           fill = continent))+
  geom_density(alpha=0.4)
```

## 13. Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point()
```

## 14. Add some color by mapping `continent` to `color` in your `aes`thetics.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp,
           color = continent))+
  geom_point()
```

## 15. Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp,
           color = continent))+
  geom_point()+
  geom_smooth()
```

## 16. Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent))+
  geom_smooth()
```

## 17. Now add an `aes`thetic to your `point`s to map `pop` to `size`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth()
```

## 18. Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth(aes(color = "black"))

# putting it inside aesthetics tries to map color to something
# in the da ta called "black", since R can't find "black", 
# it will produce some random color

ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth(color = "black")

# putting it outside aesthetics (correctly) sets color to black
```

## 19. Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth(color = "black")+
  facet_wrap(~continent)
```

## 20. Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth(color="black")+
  scale_x_log10()
```

## 21. Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

```{r}
ggplot(data = gapminder,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth(color="black")+
  scale_x_log10()+
  labs(x = "GDP per Capita",
       y = "Life Expectancy",
       color = "Continent",
       size = "Population")
```

## 22. Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 17. (You might want to take a look at your new dataframe to make sure it worked first!)

```{r}
america<-gapminder[gapminder$continent=="Americas",]

# verify this worked
america

ggplot(data = america,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth()
```

## 23. Try this again for the *whole* world, but just for observations in the year 2002.

```{r}
gap_2002<-gapminder[gapminder$year==2002,]

# verify this worked
gap_2002

ggplot(data = gap_2002,
       aes(x = gdpPercap,
           y = lifeExp))+
  geom_point(aes(color = continent, 
                 size = pop))+
  geom_smooth()
```