### 2019-09-11 content/reading/06-reading.Rmd
Update schedule format, readings for 5 and 6

---
title: "1.6 Optimize Your Workflow: Projects and Markdown - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-external-link-square-alt"></i>[Chs. 8, 27, 28, 29, 30 (but especially 27)](http://r4ds.had.co.nz) in @R4DS

# Other Resources

## On R Markdown
- Xie, Yihui, J.J. Allaire, and Garett Grolemund, [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/)
- RStudio, [*R Markdown Reference Guide*](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- RStudio, [*Overview of R Markdown*](https://rmarkdown.rstudio.com/)
- Rstudio, [*R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

## On Optimized Workflow

- Healey, Kieran, [*The Plain Person’s Guide to Plain Text Social Science*](http://plain-text.co/)
- Bryan, Jenny and Jim Hester, [*Happy Git and GitHub for the useR*](https://happygitwithr.com/)