### 2019-09-12 content/class/06-class.Rmd
Update lesson 6 stuff

---
title: "1.6: Optimize Your Workflow: Projects, Markdown, and Git - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 10, 2019*

# Overview

Today I will give you about half the class period to finish your practice problems and I can answer questions.

# Knitting on Trading Room Computers

Last year, IT and I were unable to get the Trading Room computers to knit to pdf. I may have a workaround. We will install a minimal version of $\LaTeX{}$ inside of R Studio. Run the following code:

```{r, eval=F}
tinytex::install_tinytex()
```

Then make a new `.Rmd` and try to `knit` it to `pdf`. It may install several packages (`knitr` and `rmarkdown` at least) and ask you a few prompts. It should then work.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/05-slides.html)

# Practice Problems

Today you will be working on [R practice problems](/r/05-r-practice.html). Check back later for solutions.

# Statistics and Software Survey

I have posted ["answers"](/homeworks/00-preliminary-survey-answers.html) to the statistics survey.

# Additional Useful Information

Check back here!