### 2019-09-12 content/r/06-r-practice.Rmd
Add r practice 6 links

---
title: "1.5 Data Wrangling with tidyverse - R Practice"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

- [<i class="fab fa-markdown"></i> Download R Practice 4 answers as  `.Rmd`](/r/04-r-practice-answers.Rmd)
- [<i class="fab fa-markdown"></i> See my `.Rmd` for today's slides on github](https://github.com/ryansafner/metricsF19/blob/master/static/slides/06-slides.Rmd)

Download these and look through the markdown to see how they make the webpages or slides. Note the slides are a bit advanced - I use a lot of formatting!