### 2019-09-13 content/r/06-r-practice.Rmd

---
title: "1.6 Optimize Workflow: Projects, Markdown, and Git - R Practice"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

There's currently no actual *practice* problems, but you should start taking a look at some actual `.Rmd` files to see how they work, if you are interested:

- [<i class="fas fa-external-link-square-alt"></i> Workflow R Project on GitHub](http://github.com/ryansafner/workflow)

Look at my `Workflow` R project on GitHub (first link). Also, see the 2nd-to-last slide in the slides (updated)! Follow the instructions to download and open the `.RProj` file on your computer (in RStudio). Look around the files (in the bottom-right panel), and open `Example_paper.Rmd` to see how you might write a paper in `R Markdown`.

If you are curious, below are two other examples. The first is the raw `.Rmd` file I wrote that created the (`html`) answer key for R practice 4 (on `ggplot2`). The second is the code for today's slides, written in [`Xaringan`](http://github.com/yihui/xaringan) (a special, `html`-based slides package), hosted on github. Note the slides are quite advanced - I use a lot of html and css formatting to make them pretty!

- [<i class="fab fa-markdown"></i> `04-r-practice-answers.Rmd`](/r/04-r-practice-answers.Rmd)
- [<i class="fab fa-markdown"></i> See my `.Rmd` for today's slides on github](https://github.com/ryansafner/metricsF19/blob/master/static/slides/06-slides.Rmd)