### 2019-09-13 content/assignment/01-problem-set.Rmd
Update slides 6, class 6, reading 7, add problem set 1 files, r practice 5 answers

---
title: "Problem Set 1"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set.pdf)
- [<i class="fab fa-markdown"></i> `Download Markdown`](/homeworks/01-r-problem-set.Rmd)

# Instructions

For this problem set, you may type up your answers to each question (save plots as images and include them with your answers) and also turn in a (commented!) *.R* file of commands for each relevant question. You may email or turn in a paper copy. It is not relevant for this particular problem set, but in the future, I will allow handwritten answers for math/theory questions.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

To minimize confusion, I suggest creating a new `R Project` (e.g. `hw1`) and storing any data and plots in that folder on your computer. See my [example workflow](https://github.com/ryansafner/workflow).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# The Popularity of Baby Names

Install and load the package `babynames`. Get help for `?babynames` to see what the data includes.

## 1. 
### a. 
What are the top 5 boys names for 2017, and what *percent* of overall names is each?

### b.
What are the top 5 *girls* names, and what *percent* of overall names is each?

## 2. 
Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)

## 3. 
Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.

## 4.
Make a line graph of the number of babies with your name over time, `color`ed by `sex`.

## 5.
### a.
Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]

### b.
Now do the same for girls.

## 6. Now let's graph the evolution of the most common names since 1880.
### a.
First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.

### b.
Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.

## 7.
**Bonus (hard!)**: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]

# Political and Economic Freedom Around the World

For the remaining questions, we'll look at the relationship between Economic Freedom and Political Freedom in countries around the world today. Our data for economic freedom comes from the [Fraser Institute](https://www.fraserinstitute.org/economic-freedom/dataset?geozone=world&year=2016&page=dataset), and our data for political freedom comes from [Freedom House](https://freedomhouse.org/content/freedom-world-data-and-resources).

## 8. 
Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]

- [<i class="fas fa-table"></i> `econfreedom.csv`](http://metricsf19.classes.ryansafner.com/data/econfreedom.csv)
- [<i class="fas fa-table"></i> `freedomhouse2018.csv`](http://metricsf19.classes.ryansafner.com/data/freedomhouse2018.csv)

Load them with `df<-read_csv("name_of_the_file.csv")` and save one as `econfreedom` and the other as `polfreedom`. Look at each `tibble` you've created.

## 9. 
The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country.Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.

## 10.
Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).

## 11.
Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.

## 12.
Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]

## 13.
Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]^[Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!]