### 2019-09-13 content/reading/07-reading.Rmd

---
title: "1.6 Optimize Your Workflow: Projects and Markdown - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-book"></i> @Bailey[Math and Probability Background Appendices A-I]

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review Bailey's appendices.