### 2019-09-17 content/reading/07-reading.Rmd

---
title: "2.1 Data 101 and Descriptive Statistics - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-book"></i> @Bailey[Math and Probability Background Appendix A]

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review Bailey's appendices. Today, only A is really useful, the rest will come next

Chapter 2 is optional, but will give you a good overview of using data.