### 2019-09-17 static/homeworks/01-problem-set-answers.Rmd

---
title: "Problem Set 1 (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output:
  html_document:
    df_print: paged
    theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set.pdf)
- [<i class="fab fa-markdown"></i> `04-r-practice-answers.Rmd`](/homeworks/01-r-problem-set.Rmd)

# Instructions

For this problem set, you may type up your answers to each question (save plots as images and include them with your answers) and also turn in a (commented!) *.R* file of commands for each relevant question. You may email or turn in a paper copy. It is not relevant for this particular problem set, but in the future, I will allow handwritten answers for math/theory questions.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

To minimize confusion, I suggest creating a new `R Project` (e.g. `hw1`) and storing any data and plots in that folder on your computer. See my [example workflow](https://github.com/ryansafner/workflow).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# The Popularity of Baby Names

**Install and load the package `babynames`. Get help for `?babynames` to see what the data includes.**

```{r}
library(tidyverse)

# install for first use
# install.packages("babynames")

# load package 
library(babynames)

# explore help
# ?babynames
```

## 1. 
### a. 
**What are the top 5 boys names for 2017, and what *percent* of overall names is each?**

```{r}
# save as a new tibble
top_5_boys_2017 <- babynames %>% # take data
  filter(sex=="M", # filter by males
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_boys_2017
```

The top 5 names are

1. Liam (0.95%)
2. Noah (0.93%)
3. William (0.76%)
4. James (0.72%)
5. Logan (0.71%)

### b.
**What are the top 5 *girls* names, and what *percent* of overall names is each?**

```{r}
# save as a new tibble
top_5_girls_2017 <- babynames %>% # take data
  filter(sex=="F", # filter by females
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_girls_2017
```

The top 5 names are

1. Emma (1.05%)
2. Olivia (0.99%)
3. Ava (0.85%)
4. Isabella (0.81%)
5. Sophia (0.79%)

## 2. 
**Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)**

```{r}
ggplot(data = top_5_boys_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+

  # now I'm just making it pretty
  scale_y_continuous(labels=function(x)paste(x,"%",sep=""))+ # optional, add percent signs
      labs(x = "Name",
         y = "Percent of All Babies With Name",
         title = "Most Popular Boys Names Since 1880",
         fill = "Boy's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # rotate axes!
  theme(legend.position = "") # hide legend

```

```{r}
ggplot(data = top_5_girls_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+
  # now I'm just making it pretty
  scale_y_continuous(labels=function(x)paste(x,"%",sep=""))+ # optional, add percent signs
      labs(x = "Name",
         y = "Percent of All Girls With Name",
         title = "Most Popular Girls Names Since 1880",
         fill = "Girl's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # rotate axes!
  theme(legend.position = "") # hide legend

```

## 3. 
**Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.**

```{r}
babynames %>%
  filter(name == "Ryan") %>%
  count(sex, wt=n) %>%
  mutate(percent = round((n/sum(n)*100),2))
```

## 4.
**Make a line graph of the number of babies with your name over time, `color`ed by `sex`.**
```{r}
babynames %>%
  filter(name == "Ryan") %>%
  ggplot(data = .)+
  aes(x = year,
      y = n,
      color = sex)+
  geom_line(size=1)+
  labs(x = "Year",
       y = "Number of Babies",
       title = "Popularity of Babies Named 'Ryan'",
       color = "Sex",
       caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 5.
### a.
**Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1)
```

### b.
**Now do the same for girls.**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1)
```

## 6. Now let's graph the evolution of the most common names since 1880.
### a.
**First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.**

```{r}
babynames %>%
  group_by(name) %>%
  filter(sex=="M") %>%
  summarize(total=sum(n)) %>%
  arrange(desc(total)) %>%
  slice(1:5)

top_boys_names<-c("James", "John", "Robert", "Michael", "William")
```

```{r}
babynames %>%
  group_by(name) %>%
  filter(sex=="F") %>%
  summarize(total=sum(n)) %>%
  arrange(desc(total)) %>%
  slice(1:5)

top_girls_names<-c("Mary", "Elizabeth", "Patricia", "Jennifer", "Linda")
```

### b.
**Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         name %in% top_boys_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line()+
      labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Boys Names Since 1880",
         color = "Boy's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)

```

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         name %in% top_girls_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line()+
    labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Girls Names Since 1880",
         color = "Girl's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 7.
**_Bonus (hard!)_: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]**

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  filter(Male!=0,
         Female!=0) %>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52) %>%
  mutate(total = Male+Female) %>%
  arrange(desc(total)) %>%
  slice(1:10)
```

---

# Political and Economic Freedom Around the World

**For the remaining questions, we'll look at the relationship between Economic Freedom and Political Freedom in countries around the world today. Our data for economic freedom comes from the [Fraser Institute](https://www.fraserinstitute.org/economic-freedom/dataset?geozone=world&year=2016&page=dataset), and our data for political freedom comes from [Freedom House](https://freedomhouse.org/content/freedom-world-data-and-resources).**

## 8. 
**Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]**

- [<i class="fas fa-table"></i> `econfreedom.csv`](http://metricsf19.classes.ryansafner.com/data/econfreedom.csv)
- [<i class="fas fa-table"></i> `freedomhouse2018.csv`](http://metricsf19.classes.ryansafner.com/data/freedomhouse2018.csv)

**Load them with `df<-read_csv("name_of_the_file.csv")` and save one as `econfreedom` and the other as `polfreedom`. Look at each `tibble` you've created.**

```{r}
# note these file paths will be different for you
polfreedom<-read.csv(here::here("./static/data/freedomhouse2018.csv"))
econfreedom<-read.csv(here::here("./static/data/econfreedom.csv"))

polfreedom

econfreedom
```

## 9. 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country.Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

```{r}
polfreedom<-polfreedom %>%
  select(Country.Territory, Total) %>%
  rename(Country=Country.Territory)
```

## 10.
**Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).**

```{r}
freedom<-left_join(econfreedom, polfreedom, by="Country")
```

## 11.
**Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.**

```{r, plot1, echo=F, fig.height=3, out.width="90%", fig.retina=3}
ggplot(data=freedom, aes(x=ef,y=Total))+
  geom_point(aes(color=continent))+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 12.
**Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]**

```{r, fig.retina=3}
# install.packages("ggrepel")
library(ggrepel)

interest<-filter(freedom, ISO %in% c("CHN", "NOR", "USA"))

ggplot(data=freedom, aes(x=ef,y=Total))+
  geom_point(aes(color=continent))+
  geom_label_repel(data=interest, aes(ef, Total, label=ISO,color=continent),alpha=0.6)+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)

```

## 13.
**Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!**^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]. **Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!**

```{r, fig.retina=3}
ggplot(data=freedom, aes(x=ef,y=gdp))+
  geom_point(aes(color=continent))+
  geom_smooth(data=freedom)+
  geom_label_repel(data=interest, aes(ef, Total, label=ISO,color=continent),alpha=0.6)+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```