### 2019-09-17 static/slides/07-slides.Rmd

---
title: "2.1: Data 101 and Descriptive Statistics"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)

```

# Review From 1.2: Two Big Problems with Data

.pull-left[

- We want to use econometrics to **identify** causal relationships and make **inferences** about them

1. Problem for **identification**: .shout[endogeneity]

2. Problem for **inference**: .shout[randomness]

]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/phhet82tcnuegxp/randomimage.jpg?raw=1)]
]

---

# Review from 1.2: Identification Problem: Endogeneity

.pull-left[

- An independent variable $(X)$ is .shout[exogenous] if its variation is *unrelated* to other factors that affect the dependent variable $(Y)$

- An independent variable $(X)$ is .shout[endogenous] if its variation is *related* to other factors that affect the dependent variable $(Y)$
]

.pull-right[

.center[
![](https://www.dropbox.com/s/v5vwsadw5vs448t/causality.jpg?raw=1)
]
]

---

# Review from 1.2: Inference Problem: Randomness

.pull-left[
- Data is random due to **natural sampling variation**

    - Taking one sample of a population will yield slightly different information than another sample of the same population

- Common in statistics, *easy to fix*

- **Inferential Statistics**: making claims about a wider population using sample data

    - We use common tools and techniques to deal with randomness
]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/bsdtuddzjouwzr1/sampling.jpg?raw=1)
]
]

---

# The Two Problems: Where We're Heading...Ultimately

$$\text{Sample} \underbrace{\rightarrow}_{\text{statistical inference}} \text{Population} \underbrace{\rightarrow}_{\text{causal indentification}} \text{Unobserved Parameters}$$

- We want to .whisper[identify] causal relationships between **population** variables
    - Logically first thing to consider
    - **Endogeneity problem**

- We'll use **sample** *statistics* to .whisper[infer] something about population *parameters*
    - In practice, we'll only ever have a finite *sample distribution* of data
    - We *don't* know the *population distribution* of data
    - **Randomness problem**

---

class: inverse, center, middle
# Data 101

---

# Data 101

.pull-left[

- .hl[Data] are information with context
- .hl[Individuals] are the entities described by a set of data 
    - e.g. persons, households, firms, countries 
- .hl[Variables] are particular characteristics about an individual
    - e.g. age, income, profits, population, GDP, marital status, type of legal institutions
- .hl[Observations] or .hl[cases] are the separate individuals described by a collection of variables
    - e.g. for one individual, we have their age, sex, income, education, etc.  
    - individuals and observations are *not necessarily* the same: 
        - e.g. we can have separate observations on the same individual over time 
]

.pull-right[

![](https://www.dropbox.com/s/gmb5xihskovjrly/individual1.jpg?raw=1)

]

---

# Categorical Data

.pull-left[

- .hl[Categorial data] place an individual into one of several possible *categories*
    - e.g. sex, season, political party
    - may be responses to survey questions
    - can be quantitative (e.g. age, zip code)

- `R` calls these `factors` (we'll deal with them much later in the course)
]

.pull-right[

![](https://www.dropbox.com/s/pgdswagr7b348vu/categoricaldata.png?raw=1)

]

---

# Categorical Data: Visualizing I

.pull-left[

```{r, eval=F}
diamonds %>%
  count(cut) %>%
  mutate(frequency = n / sum(n))
```

```{r, echo=F, results="asis"}
diamonds %>%
  count(cut) %>%
  mutate(frequency = n / sum(n)) %>%
knitr::kable(., format = "html", caption = "Summary of diamonds by cut")
```

]

.pull-right[

- Good way to represent categorical data is with a .hl[frequency table]

- **Count (n)**: total number of individuals in a category

- **Frequency**: proportion of a category relative to all data

]

---

# Categorical Data: Visualizing II

.pull-left[

- Charts and graphs are *always* better ways to visualize data

- A .shout[bar graph] represents categories as bars, with lengths proportional to the count or relative frequency fo each category

```{r, echo=T, eval=F}
ggplot(diamonds, aes(x=cut,
                     fill=cut))+
  geom_bar()+
  guides(fill=F)+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

.pull-right[

```{r, fig.retina=3}
ggplot(diamonds, aes(x=cut, fill=cut))+
  geom_bar()+
  guides(fill=F)+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

---

# Categorical Data: Visualizing III

.pull-left[

- Avoid pie charts!

- People are *not* good at judging 2-d differences (angles, area)

- People *are* good at judging 1-d differences (length)

]

--

.pull-right[

.center[
![](https://www.dropbox.com/s/jk79mb94qohqekb/piechart.jpg?raw=1)
]

]

---

# Categorical Data: Visualizing IV

.pull-left[

- Maybe a *stacked bar chart*

```{r, echo=T, eval=F}
diamonds %>%
  count(cut) %>%
ggplot(., aes(x="", y=n, fill=cut))+
  geom_col()+
  geom_label(aes(x="", y=n, label=cut), position = position_stack(), color="white")+
  guides(fill=F)+
  theme_void()
```

]

.pull-right[

```{r, fig.retina=3}
diamonds %>%
  count(cut) %>%
ggplot(., aes(x="", y=n, fill=cut))+
  geom_col()+
  geom_label(aes(x="", y=n, label=cut), position = position_stack(), color="white")+
  guides(fill=F)+
  theme_void()

```

]

---

# Categorical Data: Visualizing IV

.pull-left[

- Maybe *lollipop chart*

```{r, echo=T, eval=F}
diamonds %>%
  count(cut) %>%
  mutate(cut_name = as.factor(cut)) %>%
ggplot(., aes(x = cut_name, y = n, color = cut))+
 geom_point(stat="identity",
            fill="black",
            size=12)  +
  geom_segment(aes(x = cut_name, y = 0,
                   xend = cut_name,
                   yend = n), size = 2)+
  geom_text(aes(label = n),color="white", size=3) +
  coord_flip()+
  labs(x = "Cut")+
  theme_classic(base_family = "Fira Sans Condensed",
                base_size=20)+
  guides(color = F)
```

]

.pull-right[

```{r, fig.retina=3}
diamonds %>%
  count(cut) %>%
  mutate(cut_name = as.factor(cut)) %>%
ggplot(., aes(x=cut_name, y=n, color=cut))+
 geom_point(stat='identity', fill="black", size=12)  +
  geom_segment(aes(x = cut_name, 
                   y = 0,
                   xend = cut_name,
                   yend = n),
               size = 2)+
  geom_text(aes(label = n),color="white", size=3) +
  coord_flip()+
  labs(x = "Cut")+
  theme_classic(base_family = "Fira Sans Condensed",
                base_size=20)+
  guides(color = F)
```

]

---

# Categorical Data: Visualizing IV

.pull-left[

- Maybe a *treemap*

```{r, echo=T, eval=F}
library(treemapify)
diamonds %>%
  count(cut) %>%
ggplot(., aes(area = n, fill = cut)) +
  geom_treemap() +
  guides(fill = FALSE) +
  geom_treemap_text(aes(label = cut), 
                    colour = "white",
                    place = "center",
                    grow = TRUE)
```

]

.pull-right[

```{r, fig.retina=3}
library(treemapify)
diamonds %>%
  count(cut) %>%
ggplot(., aes(area = n, fill = cut)) +
  geom_treemap() +
  guides(fill = FALSE) +
  geom_treemap_text(aes(label = cut), 
                    colour = "white", place = "center", grow = TRUE)
```

]

---

# Quantitative Data I

.pull-left[

- .hl[Quantitative variables] take on numerical values of equal units that describe an individual
    - Units: points, dollars, inches 
    - Context: GPA, prices, height

- We can mathematically manipulate *only* quantitative data
    - e.g. sum, average, standard deviation

]

.pull-right[

.image-75[![](https://www.dropbox.com/s/1e0mcehgzqgnfs8/mathoperations.jpg?raw=1)]

]

---

# Context is Key!

- How variables are classified depends on the *purpose* of collecting and using the data

--

.content-box-blue[
.blue[**Quick Check**]: What kind of data (categorical or quantitative) does each variable describe?
]

--

1. Age measured in years

--

2. Age measured in ranges (0-9 years, 10-19 years, 20-29 years, etc)

--

3. The date a purchase was made

--

4. Transaction ID

--

5. The amount of money spent on a Super Bowl ad

--

6. Customer ratings

--

7. The number of correct answers on an exam

---

# Discrete Data

.pull-left[

- .hl[Discrete data] are finite, with a countable number of alternatives

- Categorical: e.g. letter grades A, B, C, D, F

- Quantitative: integers, e.g. SAT Score, number of children

]

.pull-right[

![](https://www.dropbox.com/s/qfziu7r0wob0q1n/buildingblocks.jpeg?raw=1)
]

---

# Continuous Data

.pull-left[

- .hl[Continuous data] are infinitely divisible, with an uncountable number of alternatives
    - e.g. weights, temperature, GPA

- Many discrete variables may be treated as if they are continuous
    - e.g. SAT scores, wages

]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/lgjpc9cxmrkio0x/continuous.png?raw=1)
]
]
---

# Discrete or Continuous?

.content-box-blue[
.blue[**Quick Check**]: What kind of data (discrete or continuous) does each variable describe?
]

--

1. Weight in pounds

--

2. Price in dollars

--

3. Grade (Letter)

--

4. Grade (Percentage)

--

5. Temperature

--

6. Amazon Star Rating

--

7. Number of customers

---

# Spreadsheets

.pull-left[

```{r, example.spreadsheet.setup, echo=FALSE}
ID<-c(1,2,3,4,5)
Name<-c("John","Emile","Natalya","Lakisha","Cheng")
Age<-c(23,18,28,31,36)
Sex<-c("Male","Male","Female","Female","Male")
Income<-c(41000,52600,48000,60200,81900)
example<-data.frame(ID,Name,Age,Sex,Income)
```

```{r, example.spreadsheet, results="asis", echo=FALSE}
knitr::kable(example, format="html")
```

]

.pull-right[

- The most common data structure we use is a .hl[spreadsheet]
    - Note, in *R*: a `data.frame` or `tibble`

- A **row** contains data about all variables for a single **individual**

- A **column** contains data about a single **variable** across all individuals

]

---

# Spreadsheets II

- It is common to use some notation like the following:

- Let $\{x_1, x_2, \cdots, x_n\}$ be a simple data series on variable $X$
    - $n$ individual observations
    - $x_i$ is the value of the $i$<sup>th</sup> observation for $i=1,2,\cdots, n$

--

.content-box-blue[
.blue[**Quick Check**]: Let $x$ represent the score on a homework assignment:
$$75, 100, 92, 87, 79, 0, 95$$

1. What is $n$?
2. What is $x_1$?
3. What is $x_6$?
]

---

# Datasets: Cross-Sectional

.pull-left[

```{r, example.spreadsheet2, echo=FALSE, results="asis"}
knitr::kable(example, format="html")
```

]

.pull-right[

- .hl[Cross-sectional data]: observations of individuals at a given point in time

- Each observation is a unique individual

- Simplest and most common data

- A "snapshot" to compare differences across individuals

]

---

# Datasets: Time-Series

.pull-left[

```{r, example.timeseries, echo=FALSE, results="asis"}
Year<-c(1950,1960,1970,1980,1985)
GDP<-c(8.2,9.9,10.2,12.4,13.6)
Unemployment<-c(0.06,0.04,0.08,0.08,0.06)
CPI<-c(100,118,130,190,196)
example.timeseries<-data.frame(Year,GDP,Unemployment,CPI)
knitr::kable(example.timeseries, format="html")
```

]

.pull-right[

- .hl[Time-series data]: observations of the same individuals over time

- Each observation is an individual-year

- Often used for macroeconomics, finance, and forecasting

- Unique challenges for time series

- A "moving picture" to see how individuals change over time

]

---

# Datasets: Panel

.pull-left[

```{r, example.panel.setup, echo=FALSE, results="asis"}
City<-c("Philadelphia","Philadelphia","D.C.","D.C.", "New York")
Year<-c(1986, 1990, 1986, 1990, 1986)
Murders<-c(5, 8, 2, 10, 3)
Population<-c(3.7,4.2,0.250,0.275,6.4)
Unemployment<-c(8.7,7.2,5.4,5.5,9.6)
example.panel<-data.frame(City,Year,Murders,Population,Unemployment)
knitr::kable(example.panel, format="html")
```

]

.pull-right[

- .hl[Panel], or .hl[longitudinal] dataset: a time-series for *each* cross-sectional entity

- Must be the *same* cross-sectional entities over time

- More common today for serious researchers

- Unique challenges for panel data

- A combination of "snapshot" comparisons and differences over time 
]

---

class: inverse, center, middle

# Descriptive Statistics

---

# Variables and Distributions

- Variables take on different values, we can describe a variable's .hl[distribution] (of these values)

- We want to *visualize* and *analyze* distributions to search for meaningful patterns using **statistics**

---

# Two Branches of Statistics

.pull-left[

- Two main branches of statistics:

1. .hl[Descriptive Statistics:] describes or summarizes the properties of a sample

2. .hl[Inferential Statistics:] infers properties about a larger population from the properties of a sample<sup>.red[1]</sup>

]

.pull-right[
.center[
![](https://www.dropbox.com/s/l2s6c0s55lsws6y/statsgraphs.jpg?raw=1)
]
]

.footnote[<sup>.red[1]</sup> We'll encounter inferential statistics mainly in the context of regression later.]

---

# Histograms

.pull-left[

- A common way to present a *quantitative* variable's distribution is a .hl[histogram]
    - The quantitative analog to the bar graph for a categorical variable

- Divide up values into **bins** of a certain size, and count the number of values falling within each bin, representing them visually as bars

]

.pull-right[

```{r, echo=F}
df<-tibble(x=rnorm(500,0,1))

ggplot(df, aes(x = x))+
  geom_histogram(color="white")+
  theme_bw(base_family = "Fira Sans Condensed", base_size=20)
```
]

---

# Histogram: Example

.pull-left[

.content-box-green[
.green[**Example**]: a class of 13 students takes a quiz (out of 100 points) with the following results:

$$\{ 0, 62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95 \}$$
]
]

---

# Histogram: Example

.pull-left[

.content-box-green[
.green[**Example**]: a class of 13 students takes a quiz (out of 100 points) with the following results:

$$\{ 0, 62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95 \}$$
]

```{r histogram-input, echo=T}
quizzes<-tibble(scores = c(0,62,66,71,71,74,76,79,83,86,88,93,95))
```
]

---

# Histogram: Example

.pull-left[

.content-box-green[
.green[**Example**]: a class of 13 students takes a quiz (out of 100 points) with the following results:

$$\{ 0, 62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95 \}$$
]

```{r histogram-1, echo=T, eval=F}
h<-ggplot(quizzes,aes(x=scores))+
  geom_histogram(breaks = seq(0,100,10),
                 color = "black",
                 fill = "#56B4E9")+
  scale_x_continuous(breaks = seq(0,100,10))+
  labs(x = "Scores",
       y = "Number of Students")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
h
```
]

.pull-right[
.center[
```{r, ref.label="histogram-1", fig.retina=3}
```
]
]
---

# Descriptive Statistics

.pull-left[
- We are often interested in the *shape* or *pattern* of a distribution, particularly: 
    - Measures of **center**
    - Measures of **dispersion**
    - **Shape** of distribution 
]

.pull-right[
.center[
![](https://www.dropbox.com/s/l2s6c0s55lsws6y/statsgraphs.jpg?raw=1)
]
]

---

# Mode

- The .shout[mode] of a variable is simply its most frequent value

- A variable can have multiple modes

--

.content-box-green[
.green[**Example**]: a class of 13 students takes a quiz (out of 100 points) with the following results:

$$\{ 0, 62, 66, \mathbf{71}, \mathbf{71}, 74, 76, 79, 83, 86, 88, 93, 95 \}$$
]

---

# Mode

.pull-left[

- There is no dedicated `mode()` function in `R`, surprisingly

- A workaround in `dplyr`:

```{r mode, echo=T, eval=F}
quizzes %>%
  count(scores) %>%
  arrange(desc(n))
```

]

.pull-right[
```{r, ref.label="mode"}

```
]

---

# Multi-Modal Distributions

.pull-left[

- Looking at a histogram, the modes are the "peaks" of the distribution
    - Note: depends on how wide you make the bins!

- May be unimodal, bimodal, trimodal, etc

```{r, echo=T}
tibble(scores=c(0,33,33,33,33,35,62,66,71,71,74,76,79,83,86,88,93,95)) %>%
  count(scores) %>%
  arrange(desc(n))
```

]

.pull-right[
```{r, fig.retina=3}
tibble(scores=c(0,33,33,33,33,35,62,66,71,71,74,76,79,83,86,88,93,95)) %>%
  ggplot(data = .,
         aes(x=scores))+
  geom_histogram(breaks = seq(0,100,10),
                 color = "black",
                 fill = "#56B4E9")+
  scale_x_continuous(breaks = seq(0,100,10))+
  labs(x = "Scores",
       y = "Number of Students")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Symmetry and Skew I

.pull-left[

- A distribution is **symmetric** if it looks roughly the same on either side of the "center"

- The thinner ends (far left and far right) are called the **tails** of a distribution
]

.pull-right[
```{r, fig.retina=3}
symmetric<-data.frame(x=rnorm(500,25,10))

ggplot(symmetric,(aes(x)))+
    geom_histogram(stat="bin",bins=10,color="black",fill="green")+
    theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```
]

---

# Symmetry and Skew I

.pull-left[

- If one tail stretches farther than the other, distribution is **skewed** in the direction of the longer tail
]

.pull-right[
```{r, fig.retina=3}
skew.left<-data.frame(x=rbeta(500,5,1))

ggplot(skew.left,(aes(x)))+
    geom_histogram(stat="bin",bins=10,color="black",fill="indianred")+
    theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```
]

---

# Outliers

.pull-left[

- .shout[Outlier]: extreme value that does not appear part of the general pattern of a distribution

- Can strongly affect descriptive statistics

- Might be the most informative part of the data

- Could be the result of errors

- Should always be explored and discussed!

]

.pull-right[
```{r, fig.retina=3}
h
```
]

---

class: inverse, center, middle
# Measures of Center

---

# Arithmetic Mean (Population)

- The natural measure of the center of a *population*'s distribution is its "average" or .shout[arithmetic mean `\\((\mu)\\)`]

$$\mu=\frac{x_1+x_2+...+x_N}{N} = \frac{1}{N} \sum^N_{i=1} x_i$$

- For $N$ values of variable $x$, "mu" is the sum of all individual $x$ values $(x_i)$ from 1 to $N$, divided by the $N$ number of values<sup>.red[1]</sup>

- See today's class notes for more about the summation operator, $\displaystyle\Sigma$, it'll come up again!

.footnote[<sup>.red[1]</sup> Note the mean need not be an actual value of the data!]

---

# Arithmetic Mean (Sample)

- When we have a *sample*, we compute the .shout[sample mean `\\((\bar{x})\\)`]

$$\bar{x}=\frac{x_1+x_2+...+x_n}{n} = \frac{1}{n} \sum^n_{i=1} x_i$$

- For $n$ values of variable $x$, "x-bar" is the sum of all individual $x$ values $(x_i)$ divided by the $n$ number of values

--

.pull-left[

.content-box-green[
.green[**Example**]:

$$\{0, 62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95\}$$

$$\begin{align*}
\bar{x}&=\frac{1}{13}(0+62+66+71+71+74+76+79+83+86+88+93+95)\\
\bar{x}&=\frac{944}{13}\\
\bar{x}&=72.62\\
\end{align*}$$

]
]

--

.pull-right[

```{r, echo=T}
quizzes %>%
  summarize(mean=mean(scores))
```
]

---

# Arithmetic Mean: Affected by Outliers

- If we drop the outlier (0)

--

.pull-left[

.content-box-green[
.green[**Example**]:

$$\{62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95\}$$

$$\begin{align*}
\bar{x}&=\frac{1}{12}(62+66+71+71+74+76+79+83+86+88+93+95)
&=\frac{944}{12}\\
&=78.67\\
\end{align*}$$
]
]

--

.pull-right[

```{r, echo=T}
quizzes %>%
  filter(scores>0) %>%
  summarize(mean=mean(scores))
```
]

---

# Median

$$\{0, 62, 66, 71, 71, 74, \mathbf{76}, 79, 83, 86, 88, 93, 95\}$$

- The .shout[median] is the midpoint of the distribution
    - 50% to the left of the median, 50% to the right of the median

- Arrange values in numerical order
    - For odd $n$: median is middle observation
    - For even $n$: median is average of two middle observations

---

# Mean, Median, and Outliers

.center[
![](https://www.dropbox.com/s/znl2rdi3f19r6ap/meanoutliers.jpg?raw=1)
]

---

# Mean, Median, Symmetry, Skew I

.pull-left[

- A symmetric distribution has mean $\approx$ median

```{r}
symmetric<-tibble(x=c(1,2, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 5, 6, 6, 7))
```

```{r, echo=T}
symmetric %>%
  summarize(mean = mean(x),
            median = median(x))
```
]

.pull-right[

```{r, fig.retina=3}
ggplot(symmetric,aes(x=x))+
  geom_histogram(breaks=seq(0,7,1),color="black",fill="#56B4E9")+
  geom_vline(aes(xintercept=mean(x)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(x)), color="green", linetype="dotted", size=2)+
  scale_x_continuous(breaks=seq(0,7,1))+
  labs(caption = "Mean: Red dashed line, Median: Green dotted line")+
      theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Mean, Median, Symmetry, Skew II

.pull-left[

- A left-skewed distribution has mean $<$ median

```{r}
leftskew<-data.frame(x=c(1,2,3,4,4,4,5,5,6,6,6,7,7))
```

```{r, echo=T}
leftskew %>%
  summarize(mean = mean(x),
            median = median(x))
```
]

.pull-right[

```{r, fig.retina=3}
ggplot(leftskew,aes(x=x))+
  geom_histogram(breaks=seq(0,7,1),color="black",fill="#56B4E9")+
  geom_vline(aes(xintercept=mean(x)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(x)), color="green", linetype="dotted", size=2)+
  scale_x_continuous(breaks=seq(0,7,1))+
  labs(caption = "Mean: Red dashed line, Median: Green dotted line")+
      theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Mean, Median, Symmetry, Skew III

.pull-left[

- A right-skewed distribution has mean $>$ median

```{r}
rightskew<-tibble(x=c(1,1,2,2,2,3,3,4,4,4,5,6,7))
```

```{r, echo=T}
rightskew %>%
  summarize(mean = mean(x),
            median = median(x))
```
]

.pull-right[

```{r, fig.retina=3}
ggplot(rightskew,aes(x=x))+
  geom_histogram(breaks=seq(0,7,1),color="black",fill="#56B4E9")+
  geom_vline(aes(xintercept=mean(x)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(x)), color="green", linetype="dotted", size=2)+
  scale_x_continuous(breaks=seq(0,7,1))+
  labs(caption = "Mean: Red dashed line, Median: Green dotted line")+
      theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

class: inverse, center, middle
# Measures of Spread

---

# Measures of Spread: Range

- The more *variation* in the data, the less helpful a measure of central tendency will tell us
- Beyond just the center, we also want to measure the spread
- Simplest metric is .shout[range] $=max-min$

---

# Measures of Spread: 5 Number Summary I

A common set of summary statistics about a distribution is known as the .shout["five number summary"]:

.pull-left[
.font90[
1. Minimum value
2. 25<sup>th</sup> percentile $(Q_1$, the median of the first 50% of data)
3. 50<sup>th</sup> percentile (median, $Q_2)$
4. 25<sup>th</sup> percentile $(Q_3$, the median of the last 50% of data)
5. Maximum value
]
]
.pull-right[

.font90[
```{r, echo=T}
# Base R summary command (includes Mean)
summary(quizzes$scores)

quizzes %>% # dplyr
  summarize(Min = min(scores),
            Q1 = quantile(scores, 0.25),
            Median = median(scores),
            Q3 = quantile(scores, 0.75),
            Max = max(scores))
```
]
]
---

# Measures of Spread: 5 Number Summary II

- The $n$<sup>th</sup> percentile of a distribution is the value that places $n$ percent of values beneath it

```{r, echo = T}
quizzes %>%
  summarize("37th percentile" = quantile(scores,0.37))
```

---

# Boxplots I

.pull-left[

- .shout[Boxplots] are a great way to visualize the 5 number summary

- Height of box: $Q_1$ to $Q_3$ (known as .onfire[interquartile range (IQR)], middle 50% of data)
- Line inside box: median (50<sup>th</sup> percentile)
- "Whiskers" identify data within $1.5 \times IQR$
- Points *beyond* whiskers are a common definition of **outliers**
]
.pull-right[

```{r}
quizzes %>%
  ggplot(data = .)+
  aes(x = "",
      y = scores)+
  geom_boxplot()+
    labs(x = "Quiz")+
      theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```

]

---

# Comparisons I

- Boxplots (and five number summaries) are great for comparing two distributions

.content-box-green[
.green[**Example**]:

$$\begin{align*} \text{Quiz 1}&: \{0, 62, 66, 71, 71, 74, 76, 79, 83, 86, 88, 93, 95\}	\\
			\text{Quiz 2}&: \{50, 62, 72, 73, 79, 81, 82, 82, 86, 90, 94, 98, 99\} \\ \end{align*}$$

]

```{r}
quizzes_new <- quizzes %>%
  rename(quiz_1 = scores) %>% # rename to Quiz 1
  mutate(student = 1:13) %>% # add students to track over 2 quizzes
  select(student, everything()) %>% # move students to front
  bind_cols(quiz_2 = c(50, 62, 72, 73, 79, 81, 82, 82, 86, 90, 94, 98, 99))
```

---

# Comparisons II

.pull-left[

```{r, echo=T}
quizzes_new %>% summary(quiz_1)

quizzes_new %>% summary(quiz_2)
```
]

.pull-right[

```{r}
quizzes_long<-gather(quizzes_new, key="student", value="score")

ggplot(data = quizzes_long)+
  aes(x = student,
      y = score,
      fill = student)+
  geom_boxplot()+
    labs(x = "Quiz")+
  scale_x_discrete(labels=c("Quiz 1", "Quiz 2"))+
  scale_fill_viridis_d()+
  guides(fill=F)+
      theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

---

# Aside: Making Nice Summary Tables I

- I don't like the options available for printing out summary statistics

- So I wrote my own `R` function that makes nice summary tables that uses `dplyr` and `tidyr`

- One day I will release as a package; until then the `.R` file is [saved here](/files/summaries.R)

```{r, echo=T}
# loads .R files to full functions from; make sure it's in working directory/Project!
source("../files/summaries.R")

# let's summarize "hwy" and "cty" from mpg dataset
mpg %>%
  summary_table(hwy, cty)
```

---

# Aside: Making Nice Summary Tables II

- And when `knit`ted in `R markdown`:

```{r, echo=T, results="asis"}
mpg %>%
  summary_table(hwy, cty) %>%
  knitr::kable(., format="html")
```

- We'll talk more about using `markdown` and making final products nicer when we discuss your paper project (have you forgotten?)

---

# Deviations

- Every observation .onfire[deviates] from the mean of the data: 
$$deviation_i = x_i-\mu	$$

- There are as many deviations as there are data points $(n)$
- We can measure the *average* or .shout[standard deviation] of a variable from its mean

---

# Variance (Population)

- The .shout[population variance `\\((\sigma^2)\\)`] of a *population* distribution measures the average of the *squared* deviations from the *population* mean

$$\sigma^2 = \frac{1}{N}\displaystyle\sum^N_{i=1} (x_i-\mu)^2$$

- Why do we square deviations?

- What are these units?

---

# Standard Deviation (Population)

- Square root the variance to get the .shout[population standard deviation `\\((\sigma)\\)`], the average deviation from the population mean (in same units as $x$)

$$\sigma=\sqrt{\sigma^2}=\sqrt{\frac{1}{N}\displaystyle\sum^N_{i=1} (x_i-\mu)^2	}$$

---

# Variance (Sample)

- The .shout[sample variance `\\((s^2)\\)`] of a *sample* distribution measures the average of the *squared* deviations from the *sample* mean

$$\sigma^2 = \frac{1}{n-1}\displaystyle\sum^n_{i=1} (x_i-\mu)^2$$

- Why do we divide by $n-1$?

---

# Standard Deviation (Sample)

- Square root the sample variance to get the .shout[sample standard deviation `\\((s)\\)`], the average deviation from the *sample* mean (in same units as $x$)

$$s=\sqrt{s^2}=\sqrt{\frac{1}{n-1}\displaystyle\sum^n_{i=1} (x_i-\mu)^2	}$$

---

# Sample Standard Deviation: Example

.content-box-green[
.green[**Example**]: Calculate the sample standard deviation for the following series:

$$\{2, 4, 6, 8, 10 \}$$

]

--

```{r, echo=T}
sd(c(2,4,6,8,10))
```

---

# The Steps to Calculate sd(), Coded I

```{r,echo =T}
#  first let's save our data in a tibble
sd_example<-tibble(x=c(2,4,6,8,10))

# first find the mean (just so we know)

sd_example %>%
  summarize(mean(x))

# now let's make some more columns:
sd_example <- sd_example %>%
  mutate(deviations = x-mean(x), # take deviations from mean
         deviations_sq = deviations^2) # square them

```

---

# The Steps to Calculate sd(), Coded II

.pull-left[
```{r, echo=T, eval=F}
sd_example # see what we made
```
]

.pull-right[
```{r}
sd_example
```
]
---

# The Steps to Calculate sd(), Coded III

.pull-left[
```{r, echo=T, eval=F}
sd_example %>%
  # sum the squared deviations
  summarize(sum_sq_devs = sum(deviations_sq), 
            # divide by n-1 to get variance
            variance = sum_sq_devs/(n()-1), 
            # square root to get sd
            std_dev = sqrt(variance)) 
```
]
.pull-right[
```{r}
sd_example %>%
  summarize(sum_sq_devs = sum(deviations_sq), 
            variance = sum_sq_devs/(n()-1),
            std_dev = sqrt(variance)) 
```

]

---

# Sample Standard Deviation: You Try

.content-box-blue[
.blue[**You Try**]: Calculate the sample standard deviation for the following series:

$$\{1, 3, 5, 7 \}$$

]

--

```{r, echo=T}
sd(c(1,3,5,7))
```

---

# Descriptive Statistics: Populations vs. Samples

.pull-left[
## Population parameters

- **Population size**: $N$

- **Mean**: $\mu$

- **Variance**: $\sigma^2=\frac{1}{N} \displaystyle\sum^N_{i=1} (x_i-\mu)^2$

- **Standard deviation**: $\sigma = \sqrt{\sigma^2}$
]

.pull-right[
## Sample statistics

- **Population size**: $n$

- **Mean**: $\bar{x}$

- **Variance**: $s^2=\frac{1}{n-1} \displaystyle\sum^n_{i=1} (x_i-\bar{x})^2$

- **Standard deviation**: $s = \sqrt{s^2}$

]