### 2019-09-19 content/assignment/02-problem-set.Rmd
Add class 8 content

---
title: "Problem Set 1"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set-pdf.pdf)
- [<i class="fab fa-markdown"></i> `Download Markdown`](/homeworks/01-problem-set.Rmd)

# Instructions

For this problem set, you may type up your answers to each question (save plots as images and include them with your answers) and also turn in a (commented!) *.R* file of commands for each relevant question. You may email or turn in a paper copy. It is not relevant for this particular problem set, but in the future, I will allow handwritten answers for math/theory questions.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

To minimize confusion, I suggest creating a new `R Project` (e.g. `hw1`) and storing any data and plots in that folder on your computer. See my [example workflow](https://github.com/ryansafner/workflow).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# Theory and Concepts

## 1.

### a.

In your own words, explain what (sample) standard deviation *means*.

### b.

In your own words, explain how (sample) standard deviation is calculated. You may also write the formula, but it is not necessary.