### 2019-09-19 content/class/08-class.Rmd

---
title: "2.2: Random Variables and Distributions - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Thursday, September 19, 2019*

# Overview

Today we finish your crash course/review of basic statistics with random variables and distributions.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/08-slides.html)

# Assignments: Problem Set 1 DUE and Problem Set 2

[Problem Set 1](/assignment/01-problem-set.html) is DUE

Problem set 2 (on classes 2.1-2.2) will be posted shortly, and is due by Thursday September 26.

# Appendix: Graphing Statistical and Mathematical Functions in R

Check back later today!