### 2019-09-20 content/assignment/02-problem-set.Rmd
Added problem set 2

---
title: "Problem Set 2"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 26, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/02-problem-set-pdf.pdf)
- [<i class="fab fa-markdown"></i> `Download Markdown`](/homeworks/02-problem-set.Rmd)

# Instructions

For this problem set, you may submit handwritten answers on a plain sheet of paper, or download and type/handwrite on the PDF.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# Theory and Concepts

## Question 1

### Part A 
In your own words, explain the difference between endogeneity and exogeneity.^[Hint: you may want to look back at class 1.2 for this question.]

### Part B
In your own words, explain how conducting a randomized controlled trial helps to identify the causal connection between two variables.

## Question 2

### Part A

In your own words, explain what (sample) standard deviation *means*.

### Part B

In your own words, explain how (sample) standard deviation *is calculated.* You may also write the formula, but it is not necessary.

# Problems

For the remaining questions, you may use `R` to *verify*, but please calculate all sample statistics by hand and show all work.

## Question 3

Suppose you have a very small class of four students that all take a quiz. Their scores are reported as follows:

$$\{83, 92, 72, 81\}$$

## Part A 
Calculate the median.

## Part B
Calculate the sample mean, $\bar{x}$.

## Part C
Calculate the sample standard deviation, $s$.

## Part D
Make or sketch a rough histogram of this data, with the size of each bin being 10 (i.e. 70's, 80's, 90's, 100's). You can draw this by hand or use `R`.^[If you are using `ggplot`, you want to use `+geom_histogram(breaks=seq(start,end,by))` and add `+scale_x_continuous(breaks=seq(start,end,by))`. For each, it creates bins in the histogram, and ticks on the x axis by creating a `seq`uence starting at `start` (a number), ending at `end` (number), `by` a certain interval (i.e. by `10`s.).] Is this distribution roughly symmetric or skewed? What would we expect about the mean and the median?

## Part E
Suppose instead the person who got the 72 did not show up that day to class, and got a 0 instead. Recalculate the mean and median. What happened and why?

## Question 4

Suppose the probabilities of a visitor to Amazon’s website buying 0, 1, or 2 books are 0.2, 0.4, and 0.4 respectively.

### Part A
Calculate the *expected number* of books a visitor will purchase.

### Part B 
Calculate the *standard deviation* of book purchases.

### Part C 
**Bonus**: try doing this in `R` by making an initial dataframe of the data, and then making new columns to the "table" like we did in class.

## Question 5

Scores on the SAT (out of 1600) are approximately normally distributed with a mean of 500 and standard deviation of 100.

### Part A

What is the probability of getting a score between a 400 and a 600?

### Part B

What is the probability of getting a score between a 300 and a 700?

### Part C

What is the probability of getting *at least* a 700?

### Part D

What is the probability of getting *at most* a 700?

### Part E

What is the probability of getting exactly a 500?

## Question 6

Redo problem 5 by using the `pnorm()` command in `R`.^[Hint: This function has four arguments: 1. the value of the random variable, 2. the mean of the distribution, 3. the sd of the distribution, and 4. `lower.tail` `TRUE` or `FALSE`.]