### 2019-09-24 static/homeworks/02-problem-set-answers.Rmd

---
title: "Problem Set 2 (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output:
  html_document:
    df_print: paged
    theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set.pdf)
- [<i class="fab fa-markdown"></i> `04-r-practice-answers.Rmd`](/homeworks/01-r-problem-set.Rmd)

# Instructions

For this problem set, you may submit handwritten answers on a plain sheet of paper, or download and type/handwrite on the PDF.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code).

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# Theory and Concepts

## 1.

### a. 
In your own words, explain the difference between endogeneity and exogeneity.^[Hint: you may want to look back at class 1.2 for this question.]

### b.
In your own words, explain how conducting a randomized controlled trial helps to identify the causal connection between two variables.

## 2.

### a.

In your own words, explain what (sample) standard deviation *means*.

### b.

In your own words, explain how (sample) standard deviation *is calculated.* You may also write the formula, but it is not necessary.

# Problems

## 3.

Suppose you have a very small class of four students that all take a quiz. Their scores are reported as follows:

$$\{83, 92, 72, 81\}$$

For the remaining questions, you may use `R` to *verify*, but please calculate all sample statistics by hand and show all work.

## a. 
Calculate the median.
```{r}
median(c(83,92,72,81))
```

## b.
Calculate the sample mean, $\bar{x}$.

```{r}
mean(c(83,92,72,81))
```

## c.
Calculate the sample standard deviation, $s$.

```{r}
sd(c(83,92,72,81))
```

## d.
Make or sketch a rough histogram of this data, with the size of each bin being 10 (i.e. 70's, 80's, 90's, 100's). You can draw this by hand or use `R`.^[If you are using `ggplot`, you want to use `+geom_histogram(breaks=seq(start,end,by))` and add `+scale_x_continuous(breaks=seq(start,end,by))`. For each, it creates bins in the histogram, and ticks on the x axis by creating a `seq`uence starting at `start` (a number), ending at `end` (number), `by` a certain interval (i.e. by `10`s.).] Is this distribution roughly symmetric or skewed? What would we expect about the mean and the median?

```{r}
df <- tibble(x = c(83,92,72,81))

ggplot(data = df)+
  aes(x = x)+
  geom_histogram(breaks=seq(0,100,10), color="black")+
  scale_x_continuous(breaks=seq(0,100,10))+
  theme_classic()
```

## e.
Suppose instead the person who got the 72 did not show up that day to class, and got a 0 instead. Recalculate the mean and median. What happened and why?

```{r}
mean(c(83,92,0,81))
median(c(83,92,0,81))
```

## 4.

Suppose the probabilities of a visitor to Amazon’s website buying 0, 1, or 2 books are 0.2, 0.4, and 0.4 respectively.

### a.
Calculate the *expected number* of books a visitor will purchase.

### b. 
Calculate the *standard deviation* of book purchases.

### c. 
**Bonus**: try doing this in `R` by making an initial dataframe of the data, and then making new columns to the "table" like we did in class.

## 5.

Scores on the SAT (out of 1600) are approximately normally distributed with a mean of 500 and standard deviation of 100.

### a.

What is the probability of getting a score between a 400 and a 600?

### b.

What is the probability of getting a score between a 300 and a 700?

### c.

What is the probability of getting *at least* a 700?

### d.

What is the probability of getting *at most* a 700?

### e.

What is the probability of getting exactly a 500?

## 6.

Redo problem 5 by using the `pnorm()` command in `R`.^[Hint: This function has four arguments: 1. the value of the random variable, 2. the mean of the distribution, 3. the sd of the distribution, and 4. `lower.tail` `TRUE` or `FALSE`.]