### 2019-09-24 static/slides/09-slides.Rmd

---
title: "2.3: OLS Linear Regression"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

class: inverse, center, middle
# Exploring Relationships

---

# Bivariate Data and Relationships

- We looked at single variables for descriptive statistics
- Most uses of statistics in economics and business investigate relationships *between* variables

.content-box-green[
.green[**Examples**]

- \# of police & crime rates
- healthcare spending & life expectancy
- government spending & GDP growth
- carbon dioxide emissions & temperatures
]

- We will begin with .whisper[bivariate] data for relationships between $X$ and $Y$
- Immediate aim is to explore .onfire[associations] between variables, quantified with .shout[correlation] and .shoutcaps[linear regression]
- Later we want to develop more sophisticated tools to argue for .onfire[causation]

---

# Bivariate Data: Spreadsheets I

.pull-left[

```{r econfreedomtable, echo=T, eval=F}
econfreedom <- read_csv("../data/econfreedom.csv")
head(econfreedom)
```

]

.pull-right[
```{r, echo=F, ref.label="econfreedomtable",out.width ="90%"}
```
]

- **Rows** are individual observations
- **Columns** are variables on all individuals

---

# Bivariate Data: Spreadsheets II

```{r echo=T}
econfreedom %>%
  glimpse()
```

---

# Bivariate Data: Spreadsheets III

```{r, echo=T, eval=F}
source("../files/summaries.R") # use my summary_table function

econfreedom %>%
    summary_table(ef, gdp)
```

```{r}
source("../files/summaries.R") # use my summary_table function

econfreedom %>%
    summary_table(ef, gdp) %>%
  knitr::kable(., format = "html")
```

---

# Bivariate Data: Scatterplots

.pull-left[

- The best way to visualize an association between two variables is with a .shout[scatterplot]

- Each point: pair of variable values $(x_i,y_i) \in X, Y$ for observation $i$

```{r econfreedom-scatter, eval=F, echo = T}
library("ggplot2")
ggplot(data = econfreedom)+
  aes(x = ef, y = gdp)+
  geom_point(aes(color = continent),
             size = 2)+
  labs(x = "Economic Freedom Index (2014)",
       y = "GDP per Capita (2014 USD)",
       color = "")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)+
  theme(legend.position = "bottom")
```
]

.pull-right[
```{r econfreedom-scatter-out, ref.label="econfreedom-scatter", fig.retina=3, fig.align="center"}
```

]

---

# Associations

.pull-left[

- Look for .onfire[association] between independent and dependent variables

1. **Direction**: is the trend positive or negative?

2. **Form**: is the trend linear, quadratic, something else, or no pattern?

3. **Strength**: is the association strong or weak?

4. **Outliers**: do any observations break the trends above?

]

.pull-right[
```{r, ref.label="econfreedom-scatter", fig.retina=3, fig.align="center"}
```

]

---

class: inverse, center, middle

# Quantifying Relationships

---

# Covariance

- For any two variables, we can measure their .shout[sample covariance, `\\(cov(X,Y)\\)`] or .shout[`\\(s_{X,Y}\\)`] to quantify how they vary *together*<sup>.red[1]</sup>

--

$$s_{X,Y}=E\big[(X-\bar{X})(Y-\bar{Y}) \big]$$

--

- Intuition: if $X$ is above its mean, would we expect $Y$:
    - to be *above* its mean also $(X$ and $Y$ covary *positively*)
    - to be *below* its mean $(X$ and $Y$ covary *negatively*)

--

- Covariance is a common measure, but the units are meaningless, thus we rarely need to use it so **don't worry about learning the formula**

.footnote[<sup>.red[1]</sup> Henceforth we limit all measures to *samples*, for convenience. Population covariance is denoted `\\(\sigma_{X,Y}\\)`]

---

# Covariance, in R

```{r, echo = T}
cov(econfreedom$ef,econfreedom$gdp)
```

```{r, echo = T}
econfreedom %>%
  summarize(cov = cov(ef,gdp))
```

---

# Correlation

- More convenient to *standardize* covariance into a more intuitive concept: .shout[correlation, `\\(\rho\\)`] or .shout[`\\(r\\)`] $\in [-1, 1]$

--

$$r_{X,Y}=\frac{s_{X,Y}}{s_X s_Y}=\frac{cov(X,Y)}{sd(X)sd(Y)}$$

--

- Simply weight covariance by the product of the standard deviations of $X$ and $Y$

--

- Alternatively, take the average<sup>.red[1]</sup> of the product of standardized ($Z$-scores for) each $(x_i,y_i)$ pair:<sup>.red[2]</sup>

--

$$\begin{align*}
r&=\frac{1}{n-1}\sum^n_{i=1}\bigg(\frac{x_i-\bar{X}}{s_X}\bigg)\bigg(\frac{y_i-\bar{Y}}{s_Y}\bigg)\\
r&=\frac{1}{n-1}\sum^n_{i=1}Z_XZ_Y\\
\end{align*}$$

.footnote[<sup>.red[1]</sup> Over `\\(n-1\\)`, since this is a *sample* statistic.

<sup>.red[2]</sup> See today's [class notes page](/class/09-class) for an example of the code for how to calculate correlation "by hand" in R using the second method.]

---

# Correlation: Interpretation

.pull-left[

- Correlation is standardized to 
$$-1 \leq r \leq 1$$
    - Negative values $\implies$ negative association
    - Positive values $\implies$ positive association
    - Correlation of 0 $\implies$ no association
    - As $|r| \rightarrow 1 \implies$ the stronger the association
    - Correlation of $|r|=1 \implies$ perfectly linear
]

.pull-right[

```{r, echo=F}
library("MASS")
library("gridExtra")
library("ggthemes")

# generated data for specified correlation: https://stackoverflow.com/questions/13291308/generate-numbers-with-specific-correlation
# colors chosen from http://colorbrewer2.org/#type=diverging&scheme=RdBu&n=6

# corr of -.75

dfn75 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,-0.75,-0.75,1), ncol = 2),
               empirical = TRUE)
dfn75<-data.frame(dfn75)
dfn75 <- dfn75 %>%
  rename(y=X1,
         x=X2)

corrn75<-ggplot(dfn75, aes(x=x,
                   y=y))+
  geom_point(color="#b2182b")+
  theme_void()+
  labs(title="Correlation: -0.75")

# corr of -.5

dfn50 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,-0.5,-0.5,1), ncol = 2),
               empirical = TRUE)
dfn50<-data.frame(dfn50)
dfn50 <- dfn50 %>%
  rename(y=X1,
         x=X2)

corrn50<-ggplot(dfn50, aes(x=x,
                   y=y))+
  geom_point(color="#ef8a62")+
  theme_void()+
  labs(title="Correlation: -0.50")

# corr of -.25

dfn25 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,-0.25,-0.25,1), ncol = 2),
               empirical = TRUE)
dfn25<-data.frame(dfn25)
dfn25 <- dfn25 %>%
  rename(y=X1,
         x=X2)

corrn25<-ggplot(dfn25, aes(x=x,
                   y=y))+
  geom_point(color="#fddbc7")+
  theme_void()+
  labs(title="Correlation: -0.25")

# corr of 0.25

df25 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,0.25,0.25,1), ncol = 2),
               empirical = TRUE)
df25<-data.frame(df25)
df25 <- df25 %>%
  rename(y=X1,
         x=X2)

corr25<-ggplot(df25, aes(x=x,
                   y=y))+
  geom_point(color="#d1e5f0")+
  theme_void()+
  labs(title="Correlation: 0.25")

# corr of .50

df50 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,0.50,0.50,1), ncol = 2),
               empirical = TRUE)
df50<-data.frame(df50)
df50 <- df50 %>%
  rename(y=X1,
         x=X2)

corr50<-ggplot(df50, aes(x=x,
                   y=y))+
  geom_point(color="#67a9cf")+
  theme_void()+
  labs(title="Correlation: 0.50")

# corr of 0.75

df75 <- mvrnorm(50, mu = c(0,0), Sigma = matrix(c(1,0.75,0.75,1), ncol = 2),
               empirical = TRUE)
df75<-data.frame(df75)
df75 <- df75 %>%
  rename(y=X1,
         x=X2)

corr75<-ggplot(df75, aes(x=x,
                   y=y))+
  geom_point(color="#2166ac")+
  theme_void()+
  labs(title="Correlation: 0.75")

# arrange all on grid
grid.arrange(corrn25, corrn50, corrn75, corr25, corr50, corr75, ncol=3)

```
]

---

# Guess the Correlation!

.center[
![:scale 25%](https://www.dropbox.com/s/70opitsw7jjpfko/guessthecorrelation.png?raw=1)

[Guess the Correlation Game](http://guessthecorrelation.com/index.html)
]

---

# Correlation and Endogeneity

- Your Occasional Reminder: .shoutcaps[Correlation does not imply causation!]
    - I'll show you the difference in a few weeks (when we can actually talk about causation)

- If $X$ and $Y$ are strongly correlated, $X$ can still be **endogenous**!

- See today's class notes page for more on Covariance and Correlation

---

# Always Plot Your Data!

.center[
![](https://d2f99xq7vri1nk.cloudfront.net/DinoSequentialSmaller.gif)
]

---

class: inverse, center, middle

# Linear Regression

---

# Fitting a Line to Data

.pull-left[

- If an association appears linear, we can estimate the equation of a line that would "fit" the data

$$Y = a + bX$$

- Recall a linear equation describing a line contains:<sup>.red[1]</sup>
    - $a$: vertical intercept
    - $b$: slope

.footnote[<sup>.red[1]</sup> Note we'll use different symbols for `\\(a\\)` & `\\(b\\)`, the standard econometric notation.]

]

.pull-right[

```{r, fig.retina=3}
x <- runif(100,1,9)
e <- rnorm(100,0,2.5)
y <- 12-0.5*x+e
df <- tibble(x,y)

p <- ggplot(df, aes(x=x,y=y))+
  geom_point()+
  scale_x_continuous(breaks = seq(0,20,1),
                     limits = c(0,10),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(0,20,2),
                     limits = c(0,20),
                     expand = c(0,0))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

p+geom_abline(slope=-1, intercept=14, color="red", size=1)
  #geom_smooth(method="lm", se=F)
```

]

---

# Fitting a Line to Data

.pull-left[

- If an association appears linear, we can estimate the equation of a line that would "fit" the data

$$Y = a + bX$$

- Recall a linear equation describing a line contains:<sup>.red[1]</sup>
    - $a$: vertical intercept
    - $b$: slope

- How do we choose the equation that *best* fits the data?

.footnote[<sup>.red[1]</sup> Note we'll use different symbols for `\\(a\\)` & `\\(b\\)`, the standard econometric notation.]

]

.pull-right[

```{r, fig.retina=3}
p+geom_abline(slope=-1, intercept=14, color="red", size=1)+
  geom_abline(slope=-0.125, intercept=10, color="orange", size=1)
  #geom_smooth(method="lm", se=F)
```

]

---

# Fitting a Line to Data

.pull-left[

- If an association appears linear, we can estimate the equation of a line that would "fit" the data

$$Y = a + bX$$

- Recall a linear equation describing a line contains:<sup>.red[1]</sup>
    - $a$: vertical intercept
    - $b$: slope

- How do we choose the equation that *best* fits the data?

- This process is called .shout[linear regression]

.footnote[<sup>.red[1]</sup> Note we'll use different symbols for `\\(a\\)` & `\\(b\\)`, the standard econometric notation.]

]

.pull-right[
```{r, fig.retina=3}
p+geom_abline(slope=-1, intercept=14, color="red", size=1)+
  geom_abline(slope=-0.125, intercept=10, color="orange", size=1)+
  geom_smooth(method="lm", se=F, color="green", size=2, fullrange = T)
```

]

---

# Population Linear Regression Model

- Linear regression lets us estimate the slope of the population regression line between $X$ and $Y$

- We can make **inferences** about the population slope coefficient
    - eventually, a causal interpretation
    - $\text{slope}=\frac{\Delta Y}{\Delta X}$: for a 1-unit change in $X$, how many units will this *cause* $Y$ to change?

---

# Class Size Example

.pull-left[

.content-box-green[
.green[**Example**]: What is the relationship between class size and educational performance?
]

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

---

# Class Size Example: Load the Data

```{r, echo=TRUE}
# install.packages("haven") # install for first use
library("haven") # load for importing .dta files
CASchool<-read_dta("../data/caschool.dta")
```

---

# Class Size Example: Look at the Data I

```{r, echo=T}
glimpse(CASchool)
```

---

# Class Size Example: Look at the Data II

```{r, results="asis"}
library(kableExtra)
head(CASchool) %>%
  knitr::kable(.,format="html") %>%
  kable_styling(font_size = 8)
```

---

# Class Size Example: Scatterplot

.pull-left[
```{r, school-scatter, eval=F, echo =T}
scatter<-ggplot(data = CASchool)+
  aes(x = str,
      y = testscr)+
  geom_point(color = "blue")+
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size = 20)
scatter
```
]

.pull-right[
```{r, school-scatter-out, ref.label="school-scatter", fig.retina=3}
```
]

---

# Class Size Example: Slope I

.pull-left[

- If we *change* $(\Delta)$ the class size by an amount, what would we expect the *change* in test scores to be?

$$\beta = \frac{\text{change in test score}}{\text{change in class size}} = \frac{\Delta \text{test score}}{\Delta \text{class size}}$$

- If we knew $\beta$, we could say that changing class size by 1 student will change test scores by $\beta$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

---

# Class Size Example: Slope II

.pull-left[

- Rearranging:

$$\Delta \text{test score} = \beta \times \Delta \text{class size}$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

# Class Size Example: Slope II

.pull-left[

- Rearranging:

$$\Delta \text{test score} = \beta \times \Delta \text{class size}$$

- Suppose $\beta=-0.6$. If we shrank class size by 2 students, our model predicts:

$$\begin{align*}
	\Delta \text{test score} &= -2 \times \beta\\
	\Delta \text{test score} &= -2 \times -0.6\\
	\Delta \text{test score}&= 1.2	\\
\end{align*}$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

---

# Class Size Example: Slope and Average Effect

.pull-left[

$$\text{test score} = \beta_0 + \beta_{1} \times \text{class size}$$

- The line relating class size and test scores has the above equation

- $\beta_0$ is the vertical-intercept, test score where class size is 0

- $\beta_{ClassSize}$ is the .onfire[slope] of the regression line

- This relationship only holds __on average__ for all districts in the population, *individual* districts are also affected by other factors

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

---

# Class Size Example: Marginal Effects

.pull-left[

- To get an equation that holds for *each* district, we need to include other factors

$$\text{test score} = \beta_0 + \beta_1 \text{class size}+\text{other factors}$$

- For now, we will ignore these until Unit 3

- Thus, $\beta_0 + \beta_1 \text{class size}$ gives the **average effect** of class sizes on scores

- Later, we will want to estimate the **marginal effect** (**causal effect**) of each factor on an individual district's test score, holding all other factors constant

]

.pull-right[

.center[
![](https://www.dropbox.com/s/80sb1duv6e2tazd/smallclass.jpg?raw=1)
]
]

---

# Econometric Models Overview

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + u$$

--

- $Y$ is the .shout[dependent variable] of interest
    - AKA "response variable," "regressand," "Left-hand side (LHS) variable"

--

- $X_1$ and $X_2$ are .shout[independent variables]
    - AKA "explanatory variables", "regressors," "Right-hand side (RHS) variables", "covariates"

--

- Our data consists of a spreadsheet of observed values of $(X_{1i}, X_{2i}, Y_i)$

--

- To model, we "regress Y on $X_1$ and $X_2$"

--

- $\beta_0$ and $\beta_1$ are .onfire[parameters] that describe the population relationships between the variables
    - unknown! to be estimated!

--

- $u$ is the random .shout[error term]
    - **'U'nobservable**, we can't measure it, and must model with assumptions about it

---

# The Population Regression Model

.pull-left[

- How do we draw a line through the scatterplot? We do not know the **"true"** $\beta_0$ or $\beta_1$

- We do have data from a *sample* of class sizes and test scores<sup>.red[1]</sup>

- So the real question is, .onfire[how can we estimate `\\(\beta_0\\)` and `\\(\beta_1\\)`?]
]

.pull-right[
```{r, echo=FALSE, fig.retina=3}
scatter
```
]

.footnote[<sup>.red[1]</sup> Data are student-teacher-ratio and average test scores on Stanford 9 Achievement Test for 5th grade students for 420 K-6 and K-8 school districts in California in 1999, (Stock and Watson, 2015: p. 141)]
---

class: inverse, center, middle

# Deriving OLS

---

# Deriving OLS

.pull-left[

- Suppose we have some data points

]

.pull-right[

```{r,echo =F}

x <- runif(25,0,10)
e <- rnorm(25,0,1)
y <- 2+0.5*x+e

df <- tibble(x,y)
p <- ggplot(df,
       aes(x = x,
           y = y))+
  geom_point()+
  scale_x_continuous(limits = c(0,10))+
  scale_y_continuous(limits = c(0,10))+
  #geom_smooth(method="lm")+
  theme_void()

p
```

]

---

# Deriving OLS

.pull-left[

- Suppose we have some data points
- We add a line

]

.pull-right[

```{r,echo =F}

p+geom_smooth(method="lm", se=F)
```

]

---

# Deriving OLS

.pull-left[

- Suppose we have some data points
- We add a line
- The .shout[residual], $\hat{u}$ of each data point is the difference between the .onfire[actual] and the .onfire[predicted] value of $Y$ given $X$:

$$u_i = Y_i - \hat{Y_i}$$

]

.pull-right[

```{r,echo =F}
library("broom")

reg <- lm(y ~ x, data = df)

reg_tidy <- tidy(reg)
reg_aug <- augment(reg)

p2 <- ggplot(reg_aug,
       aes(x = x,
           y = y))+
  geom_point()+
  geom_segment(aes(x=x, xend=x, y=y, yend=.fitted), color="red", linetype="dashed")+
  geom_smooth(method="lm", se=F)+
  scale_x_continuous(limits = c(0,10))+
  scale_y_continuous(limits = c(0,10))+
  theme_void()

p2
```

]

---

# Deriving OLS

.pull-left[

- Suppose we have some data points
- We add a line
- The .shout[residual], $\hat{u}$ of each data point is the difference between the .onfire[actual] and the .onfire[predicted] value of $Y$ given $X$:

$$u_i = Y_i - \hat{Y_i}$$

- We square each residual
]

.pull-right[

```{r,echo =F}
library("broom")

p2+geom_rect(aes(xmin=x, xmax=x+.resid, ymin=y, ymax=y-.resid),alpha=0.25,color="red", fill="red", linetype="dashed")
  #geom_tile(aes(x=x, y=y, width=.resid, height=.resid),alpha=0.125,fill="red",color="red")+
```

]

---

# Deriving OLS

.pull-left[

- Suppose we have some data points
- We add a line
- The .shout[residual], $\hat{u}$ of each data point is the difference between the .onfire[actual] and the .onfire[predicted] value of $Y$ given $X$:

$$u = Y_i - \hat{Y_i}$$

- We square each residual
- Add all of these up: .shout[Sum of Squared Errors (SSE)]

$$SSE = \sum^n_{i=1} u_i^2$$

]

.pull-right[

```{r,echo=F}
p2+geom_rect(aes(xmin=x, xmax=x+.resid, ymin=y, ymax=y-.resid),alpha=0.25,color="red", fill="red", linetype="dashed")
  #geom_tile(aes(x=x, y=y, width=.resid, height=.resid),alpha=0.125,fill="red",color="red")+
```

]

---

# Deriving OLS

.pull-left[

- Suppose we have some data points
- We add a line
- The .shout[residual], $\hat{u}$ of each data point is the difference between the .onfire[actual] and the .onfire[predicted] value of $Y$ given $X$:

$$u_i = Y_i - \hat{Y_i}$$

- We square each residual
- Add all of these up: .shout[Sum of Squared Errors (SSE)]

$$SSE = \sum^n_{i=1} u_i^2$$

- .onfire[The line of best fit *minimizes* SSE]

]

.pull-right[

```{r,echo =F}
p2+geom_rect(aes(xmin=x, xmax=x+.resid, ymin=y, ymax=y-.resid),alpha=0.25,color="red", fill="red", linetype="dashed")
  #geom_tile(aes(x=x, y=y, width=.resid, height=.resid),alpha=0.125,fill="red",color="red")+
```

]

---

### Corplots

```{r}
library(corrplot) # see more at https://github.com/taiyun/corrplot
library(RColorBrewer) # for color scheme used here
M <-cor(mtcars) # need to make a correlation matrix with cor(); can only include numeric variables
corrplot(M, type="upper", 
         method = "circle", # number for showing correlation coefficient
         order="alphabet", 
         col=brewer.pal(n=8, name="RdBu"))
```

---

# *O*rdinary *L*east *S*quares Estimators

- The .shout[Ordinary Least Squares (OLS) estimators] of the unknown population parameters $\beta_0$ and $\beta_1$, solve the calculus problem:

$$\min_{\beta_0, \beta_1} \sum^n_{i=1}[\underbrace{Y_i-(\underbrace{\beta_0+\beta_1 X_i}_{\hat{Y_i}})}_{u}]^2$$

--

- Intuitively, OLS estimators .onfire[minimize the average squared distance between the actual values `\\((Y_i)\\)` and the predicted values `\\((\hat{Y}_i)\\)` along the estimated regression line]

---

# The OLS Regression Line

- The .shout[OLS regression line] or .shout[sample regression line] is the linear function constructed using the OLS estimators:

$$\hat{Y_i}=\hat{\beta_0}+\hat{\beta_1}X_i$$

--

- $\hat{\beta_0}$ and $\hat{\beta_1}$ ("beta 0 hat" & "beta 1 hat") are the .shout[OLS estimators] of population parameters $\beta_0$ and $\beta_1$ using sample data

--

- The .shout[predicted value] of Y given X, based on the regression, is $E(Y_i|X_i)=\hat{Y_i}$

--

- The .onfire[residual] or .onfire[prediction error] for the $i^{th}$ observation is the difference between observed $Y_i$ and its predicted value, $\hat{u_i}=Y_i-\hat{Y_i}$

---

# The OLS Regression Estimators

- The solution to the SSE minimization problem yields:.footnote[See **Class Info Page** for proofs.]

--

$$\hat{\beta}_0=\bar{Y}-\hat{\beta_1}\bar{X}$$

--

$$\hat{\beta}_1=\frac{\displaystyle\sum^n_{i=1}(X_i-\bar{X})(Y_i-\bar{Y})}{\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2}=\frac{s_{XY}}{s^2_X}= \frac{cov(X,Y)}{var(X)}$$

---

class: inverse, center, middle

# Our Class Size Example in R

---

# Class Size Scatterplot (Again)

.pull-left[

```{r school-scatter-2, echo=T, eval=F}
scatter
```

- There is some true (unknown) population relationship: 
$$\text{test score}=\beta_0+\beta_1 \times str$$

- $\beta_1=\frac{\Delta \text{test score}}{\Delta \text{str}}= ??$

]

.pull-right[
```{r, ref.label="school-scatter-2", fig.retina=3}
```

]

---

# Class SIze Scatterplot with Regression Line

.pull-left[

```{r school-scatter-line, echo=T, eval=F}
scatter+
  geom_smooth(method = "lm", color = "red") #<<
```

]

.pull-right[
```{r school-scatter-line-out, echo=F}
scatter+
  geom_smooth(method = "lm", color = "red") #<<
```

]

---

# OLS in R

.left-code[
```{r, echo=T}
# run regression of testscr on str
school_reg <- lm(testscr ~ str, 
                 data = CASchool)
```
]

.right-plot[
- Format for regression is `lm(y ~ x, data = df)`
- `y` is dependent variable (listed first!)
- `~` means "modeled by"
- `x` is the independent variable
- `df` is the dataframe where the data is stored
]

---

# OLS in R II

.left-code[
```{r, echo=T}
# look at reg object
school_reg
```
]
.right-plot[
- Stored as an `lm` object called `school_reg`, a `list` object
]

---

# OLS in R III

.left-code[

```{r, echo=T}
summary(school_reg) # get full summary
```

]

.right-plot[

- Looking at the `summary`, there's a lot of information here!

- These objects are cumbersome, come from a much older, pre-`tidyverse` epoch of `base R`

- Luckily, we now have `tidy` ways of working with regressions!
]

---

# Tidy OLS in R: broom

.left-code[

```{r, echo=T}
# load packages
library(broom)

# tidy regression output
tidy(school_reg)
```
]

.right-plot[

- The `broom` package allows us to tidy up regression objects

- The `tidy()` function creates a tidy tibble of regression output
]

---

# OLS in R IV

.left-code[

```{r, echo=T, eval=F}
# load packages
library(broom)

# tidy regression output
tidy(school_reg,
     conf.int = TRUE) # add CI's #<<
```

```{r, echo=F}
tidy(school_reg, conf.int = TRUE)
```
]

.right-plot[

- The `broom` package allows us to tidy up regression objects

- The `tidy()` function creates a tidy tibble of regression output

- Add `conf.int = TRUE` option to include confidence intervals for each coefficient
]

---

# More broom Tools: glance

- `glance()` shows us a lot of overall regression statistics and diagnostics
    - We'll interpret these in the next lecture and beyond

```{r, echo=T}
# look at regression statistics and diagnostics
glance(school_reg)
```

---

# More broom Tools: augment

.left-code[

```{r, echo=T}
augment(school_reg)
```
]

.right-plot[

- `augment()` creates useful new variables in the stored `lm` object
    - `.fitted` are the fitted (predicted) values from the model, i.e. $\hat{Y}_i$
    - `.resid` are the residuals (errors) from the model, i.e. $\hat{u}_i$
]

---

# Class Size Regression Result I

- Using OLS, we find:
$$\widehat{\text{test score}}=689.9-2.28 \times str$$

---

# Class Size Regression Result II

- There's a great package called `equatiomatic` that prints this equation in `markdown` or $\LaTeX$.

```{r, results="asis"}
library(equatiomatic)
extract_eq(school_reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)
```

--

Here was my code:

```{r, echo = T}
# install.packages("equatiomatic") # install for first use
library(equatiomatic)
extract_eq(school_reg, 
           use_coefs = TRUE,
           coef_digits = 2,
           fix_signs = TRUE)
```

- In `R` chunk in `R markdown`, set `{r, results="asis"}` to print this raw output to be rendered

---

# Class Size Regression: A Data Point

.pull-left[
- One district in our sample is Richmond, CA:

```{r, echo = T}
CASchool %>%
  filter(district=="Richmond Elementary") %>%
  dplyr::select(district, testscr, str) # another package masks "select"
```

- Predicted value: $\widehat{\text{Test Score}}_{\text{Richmond}}=698-2.28(22) \approx 648$

- Residual: $\hat{u}_{Richmond}=672-648 \approx 24$

]

.pull-right[
```{r, echo=FALSE, fig.retina=3}

Richmond <- CASchool %>%
  filter(district=="Richmond Elementary")

ggplot(data = CASchool,
       aes(x = str,
           y = testscr))+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  geom_point(data = Richmond, color="magenta")+
  geom_text(data = Richmond, color="magenta", label="Richmond", vjust=-1)+
  geom_segment(x=22,y=672.4, xend=22,yend=652.4,linetype=2, color="magenta")+ #connect Richmond to regression line to show residual
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size = 20)
```
]