### 2019-09-25 static/homeworks/01-problem-set-answers.Rmd
Add problem set 1 answers, add broom logo in slides 2.3

---
title: "Problem Set 1 (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output:
  html_document:
    df_print: paged
    theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

Answers may be longer than I would deem sufficient on an exam. Some might vary slightly based on points of interest, examples, or personal experience. These suggested answers are designed to give you both the answer and a short explanation of *why* it is the answer.

# The Popularity of Baby Names

**Install and load the package `babynames`. Get help for `?babynames` to see what the data includes.**

```{r}
library(tidyverse)

# install for first use
# install.packages("babynames")

# load package 
library(babynames)

# explore help
# ?babynames
```

## 1. 
### a. 
**What are the top 5 boys names for 2017, and what *percent* of overall names is each?**

```{r}
# save as a new tibble
top_5_boys_2017 <- babynames %>% # take data
  filter(sex=="M", # filter by males
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_boys_2017
```

The top 5 names are

1. Liam (0.95%)
2. Noah (0.93%)
3. William (0.76%)
4. James (0.72%)
5. Logan (0.71%)

### b.
**What are the top 5 *girls* names, and what *percent* of overall names is each?**

```{r}
# save as a new tibble
top_5_girls_2017 <- babynames %>% # take data
  filter(sex=="F", # filter by females
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_girls_2017
```

The top 5 names are

1. Emma (1.05%)
2. Olivia (0.99%)
3. Ava (0.85%)
4. Isabella (0.81%)
5. Sophia (0.79%)

## 2. 
**Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)**

```{r}
ggplot(data = top_5_boys_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+

  # now I'm just making it pretty
  scale_y_continuous(labels=function(x)paste(x,"%",sep=""))+ # optional, add percent signs
      labs(x = "Name",
         y = "Percent of All Babies With Name",
         title = "Most Popular Boys Names Since 1880",
         fill = "Boy's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # rotate axes!
  theme(legend.position = "") # hide legend

```

```{r}
ggplot(data = top_5_girls_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+
  # now I'm just making it pretty
  scale_y_continuous(labels=function(x)paste(x,"%",sep=""))+ # optional, add percent signs
      labs(x = "Name",
         y = "Percent of All Girls With Name",
         title = "Most Popular Girls Names Since 1880",
         fill = "Girl's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # rotate axes!
  theme(legend.position = "") # hide legend

```

## 3. 
**Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.**

```{r}
babynames %>%
  filter(name == "Ryan") %>%
  count(sex, wt=n) %>%
  mutate(percent = round((n/sum(n)*100),2))
```

## 4.
**Make a line graph of the number of babies with your name over time, `color`ed by `sex`.**
```{r}
babynames %>%
  filter(name == "Ryan") %>%
  ggplot(data = .)+
  aes(x = year,
      y = n,
      color = sex)+
  geom_line(size=1)+
  labs(x = "Year",
       y = "Number of Babies",
       title = "Popularity of Babies Named 'Ryan'",
       color = "Sex",
       caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 5.
### a.
**Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1) # take first row only
```

### b.
**Now do the same for girls.**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         year>1980) %>%
  arrange(desc(n))%>%
  slice(1) # take first row only
```

## 6. Now let's graph the evolution of the most common names since 1880.
### a.
**First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.**

```{r}
babynames %>%
  group_by(name) %>%
  filter(sex=="M") %>%
  summarize(total=sum(n)) %>%
  arrange(desc(total)) %>%
  slice(1:5)

# make a vector of the names (we'll need this for our graph below)
top_boys_names<-c("James", "John", "Robert", "Michael", "William")
```

```{r}
babynames %>%
  group_by(name) %>%
  filter(sex=="F") %>%
  summarize(total=sum(n)) %>%
  arrange(desc(total)) %>%
  slice(1:5)

# make a vector of the names (we'll need this for our graph below)
top_girls_names<-c("Mary", "Elizabeth", "Patricia", "Jennifer", "Linda")
```

### b.
**Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.**

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         name %in% top_boys_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line()+
      labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Boys Names Since 1880",
         color = "Boy's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)

```

```{r}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         name %in% top_girls_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line()+
    labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Girls Names Since 1880",
         color = "Girl's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 7.
**_Bonus (hard!)_: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]**

There's a lot to this, so I'll break this up step by step and show you what happens at each major step.

We want to find the names where 48% to 52% of the babies with the name are male, as I defined in the footnote. First let's `mutate` a variable to figure out how many babies with a particular name are male.

To do this, we'll need to make a two variables to count the number of `male`s and `female`s of each name each year. We'll use the `ifelse()` function for each:

1. Make a `male` variable where, for each name in each year, if `sex=="M"`, then count the number of males (`n`) that year, otherwise set it equal to `0`.
2. Make a `female` variable where, for each name in each year, if `sex=="F"`, then count the number of females (`n`) that year, otherwise set it equal to `0`.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0))
```

Now with this variable, we want to count the total number of males and females with each name over the entire dataset. Let's first `group_by(name)` so we'll get one row for every name. We will `summarize()` and take the `sum` of our `male` and of our `female` variables.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))
```

Now, we want to figure out what *fraction* of each name is Male or Female. It doesn't matter which we do here, I'll do Male. `mutate()` a new variable I'll call `perc_male` for the percent of the name being for Male babies. It takes the summed variables we made before, and takes the fraction that are Male, multiplying by 100 to get percents (which isn't necessary, but is easy to read).

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  filter(Male!=0,
         Female!=0) %>%
  mutate(perc_male = (Male/(Male+Female)*100))
```

Right now, it's still in alphabetical order. We want to arrange it by `perc_male`, and more importantly, we want `perc_male` to be between 48 and 52, so let's `filter` accordingly:

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52)
```

This gives us a lot of names, all falling between 48% and 52% male. But we want the most popular names that are in this range. So let's finally `mutate` a new variable called `total` that simply adds the number of `Male` and `Female` babies with a name. Then let's `arrange` our results by `desc(total)` to get the largest first, and then `slice(1:10)` to get the top 10 only.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52) %>%
  mutate(total = Male+Female) %>%
  arrange(desc(total)) %>%
  slice(1:10)
```

---

# Political and Economic Freedom Around the World

**For the remaining questions, we'll look at the relationship between Economic Freedom and Political Freedom in countries around the world today. Our data for economic freedom comes from the [Fraser Institute](https://www.fraserinstitute.org/economic-freedom/dataset?geozone=world&year=2016&page=dataset), and our data for political freedom comes from [Freedom House](https://freedomhouse.org/content/freedom-world-data-and-resources).**

## 8. 
**Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]**

- [<i class="fas fa-table"></i> `econfreedom.csv`](http://metricsf19.classes.ryansafner.com/data/econfreedom.csv)
- [<i class="fas fa-table"></i> `freedomhouse2018.csv`](http://metricsf19.classes.ryansafner.com/data/freedomhouse2018.csv)

**Load them with `df<-read_csv("name_of_the_file.csv")` and save one as `econfreedom` and the other as `polfreedom`. Look at each `tibble` you've created.**

I am creating this document for/from the website, so these are all stored in a folder called `data`, one folder up from my current folder, `homeworks`. To get there, I go up one folder (`..`) and move to `data`, where these `csv` files are stored.

I suggest you either keep the data in the same folder as your `R` working directory (always check with `getwd()`), or create an R Project and store the data files in that same folder.

```{r}
# import data with read_csv from readr

# note these file paths will be different for you
polfreedom<-read_csv("../data/freedomhouse2018.csv")
econfreedom<-read_csv("../data/econfreedom.csv")

# look at each dataframe
polfreedom
econfreedom
```

## 9. 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country/Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

```{r}
polfreedom<-polfreedom %>%
  select(`Country/Territory`, Total) %>%
  rename(Country=`Country/Territory`)
```

## 10.
**Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).**

This one is a bit advanced to explain (but see the last few slides of 1.5 for more), so just copy what I gave you!

```{r}
freedom<-left_join(econfreedom, polfreedom, by="Country")
```

## 11.
**Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.**

```{r, plot1, echo=F, fig.height=3, out.width="90%", fig.retina=3}
ggplot(data=freedom, aes(x=ef,y=Total))+
  geom_point(aes(color=continent))+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 12.
**Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]**

```{r, fig.retina=3}
# install.packages("ggrepel") install for first use 
library(ggrepel) # load

interest<-filter(freedom, ISO %in% c("CHN", "NOR", "USA"))

ggplot(data=freedom, aes(x=ef,y=Total))+
  geom_point(aes(color=continent))+
  geom_label_repel(data=interest, aes(ef, Total, label=ISO,color=continent),alpha=0.6)+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)

```

## 13.
**Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!**^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]. **Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!**

```{r, fig.retina=3}
ggplot(data=freedom, aes(x=ef,y=gdp))+
  geom_point(aes(color=continent))+
  geom_smooth(data=freedom)+
  geom_label_repel(data=interest, aes(ef, Total, label=ISO,color=continent),alpha=0.6)+
  xlab("Economic Freedom Score")+ylab("Political Freedom Score")+theme_bw()+labs(caption="Sources: Frasier Institute, Freedom House")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```