### 2019-09-26 content/class/10-class.Rmd
Add class 10 content

---
title: "2.4: OLS: Goodness of Fit and Bias - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 24, 2019*

# Overview

Today we start looking at *associations* between variables, which we will first attempt to quantify with measures like *covariance* and *correlation*. Then we turn to fitting a line to data via *linear regression*. We overview the basic regression model, the parameters and how they are derived, and see how to work with regressions in `R` with `lm` and the tidyverse package [`broom`](https://broom.tidyverse.org/).

We consider an extended example about class sizes and test scores, which comes from a (Stata) dataset from an old textbook that I used to use, [Stock and Watson, 2007](https://www.pearson.com/us/higher-education/product/Stock-Introduction-to-Econometrics-Update-3rd-Edition/9780133486872.html?tab=resources). Download and follow along with the data from today's example:^[Note this is a `.dta` Stata file. You will need to (install and) load the package `haven` to `read_dta()` Stata files into a dataframe.]

- [<i class="fas fa-table"></i> `caschool.dta`](http://metricsf19.classes.ryansafner.com/data/caschool.dta)

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/09-slides.html)

# Assignments: Problem Set 1 DUE and Problem Set 2

[Problem Set 1](/assignment/01-problem-set.html) answers will be posted shortly, and the graded assignment emailed back to you.^[Sorry I've been too busy to send emails confirming that I recieved your problem set!]

[Problem Set 2](/assignment/01-problem-set.html) (on classes 2.1-2.2) is due by Thursday September 26.

# Appendix: OLS Estimators

## Deriving the OLS Estimators

The population linear regression model is:

$$Y_i=\beta_0+\beta_1 X_i + u _i$$

The errors $(u_i)$ are unobserved, but for candidate values of $\hat{\beta_0}$ and $\hat{\beta_1}$, we can obtain an estimate of the residual. Algebraically, the error is:

$$\hat{u_i}=    Y_i-\hat{\beta_0}-\hat{\beta_1}X_i$$

Recall our goal is to find $\hat{\beta_0}$ and $\hat{\beta_1}$ that *minimizes* the sum of squared errors (SSE):

$$SSE= \sum^n_{i=1} \hat{u_i}^2$$

So our minimization problem is:

$$\min_{\hat{\beta_0}, \hat{\beta_1}} \sum^n_{i=1} (Y_i-\hat{\beta_0}-\hat{\beta_1}X_i)^2$$

Using calculus, we take the partial derivatives and set it equal to 0 to find a minimum. The first order conditions are:

$$\begin{align*}
\frac{\partial SSE}{\partial \hat{\beta_0}}&=-2\displaystyle\sum^n_{i=1} (Y_i-\hat{\beta_0}-\hat{\beta_1} X_i)=0\\
\frac{\partial SSE}{\partial \hat{\beta_1}}&=-2\displaystyle\sum^n_{i=1} (Y_i-\hat{\beta_0}-\hat{\beta_1} X_i)X_i=0\\
\end{align*}$$

### Finding $\hat{\beta_0}$

Working with the first FOC, divide both sides by $-2$:

$$\displaystyle\sum^n_{i=1} (Y_i-\hat{\beta_0}-\hat{\beta_1} X_i)=0$$

Then expand the summation across all terms and divide by $n$:

$$\underbrace{\frac{1}{n}\sum^n_{i=1} Y_i}_{\bar{Y}}-\underbrace{\frac{1}{n}\sum^n_{i=1}\hat{\beta_0}}_{\hat{\beta_0}}-\underbrace{\frac{1}{n}\sum^n_{i=1} \hat{\beta_1} X_i}_{\hat{\beta_1}\bar{X}}=0$$

Note the first term is $\bar{Y}$, the second is $\hat{\beta_0}$, the third is $\hat{\beta_1}\bar{X}$.^[From the [rules about summation operators](https://metricsf19.classes.ryansafner.com/class/07-class/#the-summation-operator), we define the mean of a random variable $X$ as $\bar{X}=\frac{1}{n}\displaystyle\sum_{i=1}^n X_i$. The mean of a constant, like $\beta_0$ or $\beta_1$ is itself.]

So we can rewrite as: 
$$\bar{Y}-\hat{\beta_0}-\beta_1=0$$

Rearranging:

$$\hat{\beta_0}=\bar{Y}-\bar{X}\beta_1$$

### Finding $\hat{\beta_1}$

To find $\hat{\beta_1}$, take the second FOC and divide by $-2$:

$$\displaystyle\sum^n_{i=1} (Y_i-\hat{\beta_0}-\hat{\beta_1} X_i)X_i=0$$

From the formula for $\hat{\beta_0}$, substitute in for $\hat{\beta_0}$:

$$\displaystyle\sum^n_{i=1} \bigg(Y_i-[\bar{Y}-\hat{\beta_1}\bar{X}]-\hat{\beta_1} X_i\bigg)X_i=0$$

Combining similar terms:

$$\displaystyle\sum^n_{i=1} \bigg([Y_i-\bar{Y}]-[X_i-\bar{X}]\hat{\beta_1}\bigg)X_i=0$$

Distribute $X_i$ and expand terms into the subtraction of two sums (and pull out $\hat{\beta_1}$ as a constant in the second sum:

$$\displaystyle\sum^n_{i=1} [Y_i-\bar{Y}]X_i-\hat{\beta_1}\displaystyle\sum^n_{i=1}[X_i-\bar{X}]X_i=0$$

Move the second term to the righthand side:

$$\displaystyle\sum^n_{i=1} [Y_i-\bar{Y}]X_i=\hat{\beta_1}\displaystyle\sum^n_{i=1}[X_i-\bar{X}]X_i$$

Divide to keep just $\hat{\beta_1}$ on the right:

$$\frac{\displaystyle\sum^n_{i=1} [Y_i-\bar{Y}]X_i}{\displaystyle\sum^n_{i=1}[X_i-\bar{X}]X_i}=\hat{\beta_1}$$

Note that from the [rules about summation operators](https://metricsf19.classes.ryansafner.com/class/07-class/#the-summation-operatorproperties):

$$\displaystyle\sum^n_{i=1} [Y_i-\bar{Y}]X_i=\displaystyle\sum^n_{i=1} (Y_i-\bar{Y})(X_i-\bar{X})$$

and:

$$\displaystyle\sum^n_{i=1} [X_i-\bar{X}]X_i=\displaystyle\sum^n_{i=1} (X_i-\bar{X})(X_i-\bar{X})=\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2$$

Plug in these two facts:

$$\frac{\displaystyle\sum^n_{i=1} (Y_i-\bar{Y})(X_i-\bar{X})}{\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2}=\hat{\beta_1}$$

## Algebraic Properties of OLS Estimators

The OLS residuals $\hat{u}$ and predicted values $\hat{Y}$ are chosen by the minimization problem to satisfy:

1. The expected value (average) error is 0: 
$$E(u_i)=\frac{1}{n}\displaystyle \sum_{i=1}^n \hat{u_i}=0$$

2. The covariance between $X$ and the errors is 0:
$$\hat{\sigma}_{X,u}=0$$

Note the first two properties imply strict **exogeneity**. That is, this is only a valid model if $X$ and $u$ are not correlated.

3. The expected predicted value of $Y$ is equal to the expected value of $Y$:
$$\bar{\hat{Y}}=\frac{1}{n} \displaystyle\sum_{i=1}^n \hat{Y_i} = \bar{Y}$$

4. Total sum of squares is equal to the explained sum of squares plus sum of squared errors:
$$\begin{align*}TSS&=ESS+SSE\\
\sum_{i=1}^n (Y_i-\bar{Y})^2&=\sum_{i=1}^n (\hat{Y_i}-\bar{Y})^2 + \sum_{i=1}^n {u}^2\\
\end{align*}$$

Recall $R^2$ is $\frac{ESS}{TSS}$ or $1-SSE$

5. The regression line passes through the point $(\bar{X},\bar{Y})$, i.e. the mean of $X$ and the mean of $Y$.

## Bias in $\hat{\beta_1}$

Begin with the formula we derived for $\hat{\beta_1}$:

$$\hat{\beta_1}=\frac{\displaystyle \sum^n_{i=1} (Y_i-\bar{Y})(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$

Recall from **Rule 6** of summations, we can rewrite the numerator as

$$\begin{align*}
	=&\displaystyle \sum^n_{i=1} (Y_i-\bar{Y})(X_i-\bar{X})\\
	=& \displaystyle \sum^n_{i=1} Y_i(X_i-\bar{X})\\
\end{align*}$$

$$\hat{\beta_1}=\frac{\displaystyle \sum^n_{i=1} Y_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$

We know the true population relationship is expressed as:

$$Y_i=\beta_0+\beta_1 X_i+u_i$$

Substituting this in for $Y_i$ in equation 2:

$$\hat{\beta_1}=\frac{\displaystyle \sum^n_{i=1} (\beta_0+\beta_1X_i+u_i)(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$
Breaking apart the sums in the numerator:

$$\hat{\beta_1}=\frac{\displaystyle \sum^n_{i=1} \beta_0(X_i-\bar{X})+\displaystyle \sum^n_{i=1} \beta_1X_i(X_i-\bar{X})+\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$

We can simplify equation 4 using **Rules 4 and 5** of summations

1. The first term in the numerator $\left[\displaystyle \sum^n_{i=1} \beta_0(X_i-\bar{X})\right]$ has the constant $\beta_0$, which can be pulled out of the summation. This gives us the summation of deviations, which add up to 0 as per **Rule 4**:

$$\begin{align*}
\displaystyle \sum^n_{i=1} \beta_0(X_i-\bar{X})&= \beta_0 \displaystyle \sum^n_{i=1} (X_i-\bar{X})\\
&=\beta_0 (0)\\
&=0\\
\end{align*}$$

2. The second term in the numerator  $\left[\displaystyle \sum^n_{i=1} \beta_1X_i(X_i-\bar{X})\right]$ has the constant $\beta_1$, which can be pulled out of the summation. Additionally, **Rule 5** tells us $\displaystyle \sum^n_{i=1} X_i(X_i-\bar{X})=\displaystyle \sum^n_{i=1}(X_i-\bar{X})^2$:

$$\begin{align*}
\displaystyle \sum^n_{i=1} \beta_1X_1(X_i-\bar{X})&= \beta_1 \displaystyle \sum^n_{i=1} X_i(X_i-\bar{X})\\
&=\beta_1\displaystyle \sum^n_{i=1}(X_i-\bar{X})^2\\
\end{align*}$$

When placed back in the context of being the numerator of a fraction, we can see this term simplifies to just $\beta_1$:

$$\begin{align*}
	\frac{\beta_1\displaystyle \sum^n_{i=1}(X_i-\bar{X})^2}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} &=\frac{\beta_1}{1} \times \frac{\displaystyle \sum^n_{i=1}(X_i-\bar{X})^2}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}\\
	&=\beta_1	\\
\end{align*}$$

Thus, we are left with:

$$\hat{\beta_1}=\beta_1+\frac{\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$

Now, take the expectation of both sides:

$$E[\hat{\beta_1}]=E\left[\beta_1+\frac{\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} \right]$$

We can break this up, using properties of expectations. First, recall $E[a+b]=E[a]+E[b]$, so we can break apart the two terms.

$$E[\hat{\beta_1}]=E[\beta_1]+E\left[\frac{\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} \right]$$

Second, the true population value of $\beta_1$ is a constant, so $E[\beta_1]=\beta_1$.

Third, since we assume $X$ is also "fixed" and not random, the variance of $X$, $\displaystyle\sum_{i=1}^n (X_i-\bar{X})$, in the denominator, is just a constant, and can be brought outside the expectation.

$$E[\hat{\beta_1}]=\beta_1+\frac{E\left[\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})\right]	}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2}$$

Thus, the properties of the equation are primarily driven by the expectation $E\bigg[\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})\bigg]$. We now turn to this term.

Use the property of summation operators to expand the numerator term:

$$\begin{align*}
	\hat{\beta_1}&=\beta_1+\frac{\displaystyle \sum^n_{i=1} u_i(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} \\
		\hat{\beta_1}&=\beta_1+\frac{\displaystyle \sum^n_{i=1} (u_i-\bar{u})(X_i-\bar{X})}{\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} \\
\end{align*}$$

Now divide the numerator and denominator of the second term by $\frac{1}{n}$. Realize this gives us the covariance between $X$ and $u$ in the numerator and variance of $X$ in the denominator, based on their respective definitions.

$$\begin{align*}
	\hat{\beta_1}&=\beta_1+\cfrac{\frac{1}{n}\displaystyle \sum^n_{i=1} (u_i-\bar{u})(X_i-\bar{X})}{\frac{1}{n}\displaystyle\sum^n_{i=1} (X_i-\bar{X})^2} \\
	\hat{\beta_1}&=\beta_1+\cfrac{cov(X,u)}{var(X)} \\
	\hat{\beta_1}&=\beta_1+\cfrac{s_{X,u}}{s_X^2} \\
\end{align*}$$

By the Zero Conditional Mean assumption of OLS, $s_{X,u}=0$.

Alternatively, we can express the bias in terms of correlation instead of covariance:

$$E[\hat{\beta_1}]=\beta_1+\cfrac{cov(X,u)}{var(X)}$$

From the definition of correlation:

$$\begin{align*}
	 corX,u)&=\frac{cov(X,u)}{s_X s_u}\\
	 corX,u)s_Xs_u &=cov(X,u)\\
\end{align*}$$

Plugging this in:

$$\begin{align*}
E[\hat{\beta_1}]&=\beta_1+\frac{cov(X,u)}{var(X)} \\
E[\hat{\beta_1}]&=\beta_1+\frac{\big[corX,u)s_xs_u\big]}{s^2_X} \\
E[\hat{\beta_1}]&=\beta_1+\frac{corX,u)s_u}{s_X} \\
E[\hat{\beta_1}]&=\beta_1+corX,u)\frac{s_u}{s_X} \\
\end{align*}$$

## Proof of the Unbiasedness of $\hat{\beta_1}$

Begin with equation:

$$\hat{\beta_1}=\frac{\sum Y_iX_i}{\sum X_i^2}$$

Substitute for $Y_i$:

$$\hat{\beta_1}=\frac{\sum (\beta_1 X_i+u_i)X_i}{\sum X_i^2}$$

Distribute $X_i$ in the numerator:

$$\hat{\beta_1}=\frac{\sum \beta_1 X_i^2+u_iX_i}{\sum X_i^2}$$

Separate the sum into additive pieces:

$$\hat{\beta_1}=\frac{\sum \beta_1 X_i^2}{\sum X_i^2}+\frac{u_i X_i}{\sum X_i^2}$$

$\beta_1$ is constant, so we can pull it out of the first sum:

$$\hat{\beta_1}=\beta_1 \frac{\sum X_i^2}{\sum X_i^2}+\frac{u_i X_i}{\sum X_i^2}$$

Simplifying the first term, we are left with:

$$\hat{\beta_1}=\beta_1 +\frac{u_i X_i}{\sum X_i^2}$$

Now if we take expectations of both sides:

$$E[\hat{\beta_1}]=E[\beta_1] +E\left[\frac{u_i X_i}{\sum X_i^2}\right]$$

$\beta_1$ is a constant, so the expectation of $\beta_1$ is itself.

$$E[\hat{\beta_1}]=\beta_1 +E\bigg[\frac{u_i X_i}{\sum X_i^2}\bigg]$$

Using the properties of expectations, we can pull out $\frac{1}{\sum X_i^2}$ as a constant:

$$E[\hat{\beta_1}]=\beta_1 +\frac{1}{\sum X_i^2} E\bigg[\sum u_i X_i\bigg]$$

Again using the properties of expectations, we can put the expectation inside the summation operator (the expectation of a sum is the sum of expectations):

$$E[\hat{\beta_1}]=\beta_1 +\frac{1}{\sum X_i^2}\sum E[u_i X_i]$$

Under the exogeneity condition, the correlation between $X_i$ and $u_i$ is 0.

$$E[\hat{\beta_1}]=\beta_1$$