### 2019-09-26 static/slides/10-slides.Rmd

---
title: "2.4: OLS: Goodness of Fit and Bias"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

```{r regression-setup, echo=F, results="hide"}

library(haven)
CASchool<-read_dta("../data/caschool.dta")

# run regression of testscr on str
school_reg <- lm(testscr ~ str, 
                 data = CASchool)

library(broom)
school_reg_tidy <- tidy(school_reg,
     conf.int = TRUE) # add confidence intervals
CASchool_aug <- augment(school_reg)

library(equatiomatic)
extract_eq(school_reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)

```

class: inverse, center, middle
# Goodness of Fit

---

# Models

> "All models are wrong. But some are useful." - George Box

--

# All of Statistics:

$$Observed_i = Model_i + Error_i$$

---

# Goodness of Fit

- How well does a line fit data? How tightly clustered around the line are the data points?

- Quantify .onfire[how much variation in `\\(Y_i\\)` is "explained" by the model]

--

$$\underbrace{Y_i}_{Observed}=\underbrace{\widehat{Y_i}}_{Model}+\underbrace{\hat{u}}_{Error}$$

- Recall OLS estimators are chosen specifically to minimize the .shout[Sum of Squared Errors (SSE)]: $\left(\displaystyle \sum^n_{i=1}\hat{u_i}^2\right)$

---

# Goodness of Fit: $R^2$

- Primary measure<sup>.red[1]</sup> is .shout[regression R-squared], the fraction of variation in $Y$ explained by variation in predicted values

$$R^2 = \frac{\text{ variation in }\widehat{Y_i}}{\text{variation in }Y_i}$$

.footnote[<sup>.red[1]</sup> Sometimes called the .shout["coefficient of determination"]]

---

# Goodness of Fit: $R^2$ Formula

$$R^2 =\frac{ESS}{TSS}$$

--

- .shout[Explained Sum of Squares (ESS)]:<sup>.red[1]</sup> sum of squared deviations of *predicted* values from their mean<sup>.red[2]</sup>

--

$$ESS= \sum^n_{i=1}(\hat{Y_i}-\bar{Y})^2$$

--

- .shout[Total Sum of Squares (TSS)]: sum of squared deviations of *observed* values from their mean

--

$$TSS= \sum^n_{i=1}(Y_i-\bar{Y})^2$$

.footnote[

<sup>1</sup> Sometimes called Model Sum of Squares (MSS) or Regression Sum of Squares (RSS) in other textbooks

<sup>2</sup> It can be shown that $\bar{\hat{Y_i}}=\bar{Y}$]

---

# Goodness of Fit: $R^2$ Formula II

- Equivalently, the complement of the fraction of *unexplained* variation in $Y_i$

$$R^2=1-\frac{SSE}{TSS}$$

- Equivalently, the square of the correlation coefficient between $X$ and $Y$:

--

$$R^2=(r_{X,Y})^2$$

---

# Calculating $R^2$ in R I

- If we wanted to calculate it manually:

```{r, echo=T}
# as squared correlation coefficient

# Base R
cor(CASchool$testscr, CASchool$str)^2

# dplyr
CASchool %>%
  summarize(r_sq = cor(testscr,str)^2)
```

---

# Calculating $R^2$ in R II

- Recall `broom`'s `augment()` command makes a lot of new regression-based values, such as:
    - `.fitted`: predicted values $(\hat{Y_i})$
    - `.resid`: residuals $(\hat{u_i})$

```{r, echo=T}
library(broom)
school_reg %>%
  augment() %>%
  head(., n=5) # show first 5 values
```

---

# Calculating $R^2$ in R III

- We can calculate `R` as the ratio of variances in model vs. actual (i.e. akin to $\frac{ESS}{TSS}$)

```{r, echo = T}
# as ratio of variances
school_reg %>%
  augment() %>%
  summarize(r_sq = var(.fitted)/var(testscr)) # variance of *predicted* testscr over variance of *actual* teestscr
```

---

# Goodness of Fit: Standard Error of the Regression

- The .shout[Standard Error of the Regression], $\hat{\sigma}$ or $\hat{\sigma}_{u}$ is an estimator of the standard deviation of $u_i$

--

$$\hat{\sigma_u}=\sqrt{\frac{SSE}{n-2}}$$

--

- Measures the **average size of the residuals** (distance between a data point and the regression line)
    - **Degrees of Freedom correction** of $n-2$: we use up 2 df to first calculate $\hat{\beta_0}$ and $\hat{\beta_1}$!

---

# Calculating SER in R

.pull-left[

```{r, echo=T, eval=F}
school_reg %>%
  augment() %>%
  summarize(SSE = sum(.resid^2),
            df = n()-1-1,
            SER = sqrt(SSE/df))
```
]

.pull-right[

```{r, echo=F}
school_reg %>%
  augment() %>%
  summarize(SSE = sum(.resid^2),
            df = n()-1-1,
            SER = sqrt(SSE/df))
```
]

--

In large samples (where $n-2$ is very close to $n)$, SER converges to just the standard deviation of the residuals

.pull-left[

```{r, echo=T, eval =F}
school_reg %>%
  augment() %>%
  summarize(sd_resid = sd(.resid))
```
]

.pull-right[
```{r, echo=F}
school_reg %>%
  augment() %>%
  summarize(sd_resid = sd(.resid))
```

]

---

# Goodness of Fit: Looking at R I

.left-code[
- `summary()` command in `Base R` gives:
  - `Multiple R-squared`
  - `Residual standard error` (SER)
    - Calculated with a df of $n-2$

```{r, echo=T, eval=F}
# Base R
summary(school_reg)
```
]

.right-plot[
```{r}
summary(school_reg)
```
]

---

# Goodness of Fit: Looking at R II

```{r, echo=T}
# using broom
library(broom)
glance(school_reg)
```

--

- `r.squared` is `0.05` $\implies$ about 5% of variation in `testscr` is explained by our model
- `sigma` (SER) is `18.6` $\implies$ average test score is about 18.6 points above/below our model's prediction

--

```{r, echo=T}
# extract it if you want with $r.squared
glance(school_reg)$r.squared
```

---

class: inverse, center, middle

# Bias: The Sampling Distributions of the OLS Estimators

---

# Recall: The Two Big Problems with Data

.pull-left[

- We use econometrics to .onfire[identify] causal relationships and make .onfire[inferences] about them

1. Problem for .onfire[identification]: .shout[endogeneity]
    - $X$ is **exogenous** if its variation is *unrelated* to other factors $(u)$ that affect $Y$
    - $X$ is **endogenous** if its variation is *related* to other factors $(u)$ that affect $Y$

2. Problem for .onfire[inference]: .shout[randomness]
    - Data is random due to **natural sampling variation**
    - Taking one sample of a population will yield slightly different information than another sample of the same population

]

.pull-right[
.center[
![:scale 55%](https://www.dropbox.com/s/v5vwsadw5vs448t/causality.jpg?raw=1)

![:scale 55%](https://www.dropbox.com/s/phhet82tcnuegxp/randomimage.jpg?raw=1)]
]

---

# Distributions of the OLS Estimators

- OLS estimators $(\hat{\beta_0}$ and $\hat{\beta_1})$ are computed from a finite (specific) sample of data

- Our OLS model contains **2 sources of randomness**:

--

- .onfire[*Modeled* randomness]: $u$ includes all factors affecting $Y$ *other* than $X$
    - different samples will have different values of those other factors $(u_i)$

--

- .onfire[*Sampling* randomness]: different samples will generate different OLS estimators
    - Thus, $\hat{\beta_0}, \hat{\beta_1}$ are *also* **random variables**, with their own <span class=shout>sampling distribution</span>

---

# Inferential Statistics and Sampling Distributions

.pull-left[

- **Inferential statistics** analyzes a **sample** to make inferences about a much larger (unobservable) **population**

- .shout[Population]: all possible individuals that match some well-defined criterion of interest (people, firms, cities, etc)
  - Characteristics about (relationships between variables describing) populations are called **parameters**

- .shout[Sample]: some portion of the population of interest to *represent the whole*
  - Samples examine part of a population to generate **statistics** used to estimate population parameters

- We almost never can directly study the population, so we *model* it with our samples 
]

.pull-right[
.center[
![](https://www.dropbox.com/s/17cb7ceqipgi8fs/citymodel.jpg?raw=1)
]
]

---

# Sampling Basics

.content-box-green[
.green[**Example**]: Suppose you randomly select 100 people and ask how many hours they spend on the internet each day. You take the mean of your sample, and it comes out to 5.4 hours. 
]

--

- 5.4 hours is a **sample statistic** describing the sample
- We are more interested in the corresponding **parameter**	of the relevant population (e.g. all Americans)

--

- If we take another sample of $n=100$ people, would we get the same number?

--

- Roughly, but probably not exactly

- .shout[Sampling variability] describes the effect of a statistic varying somewhat from sample to sample
    - This is normal, not the result of any error or bias!

---

# I.I.D. Samples

.pull-left[

- If we collect many samples, and each sample is randomly drawn from the population (and then replaced), then the distribution of samples is said to be .onfire[independently and identically distributed (i.i.d.)]

- Each sample is **independent** of each other sample (due to replacement)

- Each sample comes from the **identical** underlying population distribution

]

.pull-right[
.center[
![:scale 90%](https://www.dropbox.com/s/bsdtuddzjouwzr1/sampling.jpg?raw=1)
]
]

---

# The Sampling Distribution of OLS Estimators

.pull-left[
- Calculating OLS estimators for a sample makes the OLS estimators *themselves* random variables:

- Draw of $i$ is random $\implies$ value of each $(X_i,Y_i)$ is random $\implies$ $\hat{\beta_0},\hat{\beta_1}$ are random

- Taking different samples will create different values of $\hat{\beta_0},\hat{\beta_1}$

- Therefore, $\hat{\beta_0},\hat{\beta_1}$ each have a .shout[sampling distribution] across different samples

]

.pull-right[

.center[

![:scale 75%](https://www.dropbox.com/s/phhet82tcnuegxp/randomimage.jpg?raw=1)]

]

---

# The Central Limit Theorem

- .shout[Central Limit Theorem (CLT)]: if we collect samples of size $n$ from the same population and generate a sample statistic (e.g. OLS estimator), then with large enough $n$, the distribution of the sample statistic is approximately normal IF 
    1. $n \geq 30$
    2. Samples come from a *known* normal distribution $\sim N(\mu,\sigma)$

--

- If neither of these are true, we have other methods (coming shortly!)

- One of the most fundamental principles in all of statistics

- Allows for virtually all testing of statistical hypotheses $\rightarrow$ estimating probabilities of values on a normal distribution

---

# The Sampling Distribution of $\hat{\beta_1}$ I

.pull-left[

- The CLT allows us to approximate the sampling distributions of $\hat{\beta_0}$ and $\hat{\beta_1}$ as normal
    - Generally agreed for $n>100$

- We care about $\hat{\beta_1}$ (slope) since it has economic meaning, rarely about $\hat{\beta_0}$ (intercept)

$$\hat{\beta_1} \sim N(E[\hat{\beta_1}], \sigma_{\hat{\beta_1}})$$

]

.pull-right[

```{r, fig.retina=3}
beta_dist<-ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  geom_segment(aes(x=0,xend=0, y=0, yend=0.4), linetype="dashed")+
  scale_x_continuous(breaks = 0,
                     labels = expression(E(hat(beta[1]))))+
  labs(x = expression(hat(beta[1])),
       y = "Probability")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
beta_dist
```

]
---

# The Sampling Distribution of $\hat{\beta_1}$ II

.pull-left[

$$\hat{\beta_1} \sim N(E[\hat{\beta_1}], \sigma_{\hat{\beta_1}})$$

- We want to know:

1. $E[\hat{\beta_1}]$; what is the **center** of the distribution? (today)

2. $\sigma_{\hat{\beta_1}}$; how **precise** is our estimate? (next class)

]

.pull-right[

```{r, fig.retina=3}
beta_dist
```

]

---

class: inverse, center, middle
# Bias and Exogeneity

---

# Assumptions about Errors I

.pull-left[

- In order to talk about $E[\hat{\beta_1}]$, we need to talk about $u$

- Recall: $u$ is a random variable, and we can never measure the error term

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions about Errors II

.pull-left[

- We make .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions about Errors II

.pull-left[

- We make .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions about Errors II

.pull-left[

- We make .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

2. The variance of the residuals is constant, written:
$$var(u)=\sigma^2_{u}$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions about Errors II

.pull-left[

- We make .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

2. The variance of the residuals is constant, written:
$$var(u)=\sigma^2_{u}$$

3. Errors are not correlated across observations: 
$$cor(u_i,u_j)=0 \quad \forall i \neq j$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions about Errors II

.pull-left[

- We make .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

2. The variance of the residuals is constant, written:
$$var(u)=\sigma^2_{u}$$

3. Errors are not correlated across observations: 
$$cor(u_i,u_j)=0 \quad \forall i \neq j$$

4. There is no correlation between $X$ and the error term: 
$$cor(X, u)=0 \text{ or } E[u|X]=0$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions 1 and 2: Errors are i.i.d.

.pull-left[

1. The expected value of the residuals is 0
$$E[u]=0$$

2. The variance of the residuals is constant, written:
$$var(u)=\sigma^2_{u}$$

- The first two assumptions $\implies$ errors are **i.i.d.**, drawn from the same distribution with mean 0 and variance $\sigma^2_{u}$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumption 2: Heteroskedasticity

.pull-left[

- The variance of the residuals is constant, written:
$$var(u)=\sigma^2_{u}$$

- Assumption 2 implies that errors are ".shout[homoskedastic]:
 they have the same variance across $X$

- Often this assumption is violated: errors may be ".shout[heteroskedastic]:
 they do not have the same variance across $X$

- This *is* a problem for **inference**, but we have a simple fix for this (coming shortly)

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumption 3: No Serial Correlation

.pull-left[

- Errors are not correlated across observations: 
$$cor(u_i,u_j)=0 \quad \forall i \neq j$$

- For simple cross-sectional data, this is rarely an issue

- Time-series & panel data nearly always contain .shout[serial correlation] or .shout[autocorrelation] between errors

- e.g. "this week's sales look a lot like last weel's sales, which look like...etc"

- There are fixes to deal with autocorrelation (coming much later)

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumption 4: The Zero Conditional Mean Assumption

.pull-left[

- There is no correlation between $X$ and the error term: 
$$cor(X, u)=0$$

- .onfire[This is the absolute killer assumption, because it assumes **exogeneity**]

- Often called the .shout[Zero Conditional Mean] assumption: 
$$E[u|X]=0$$

- "Does knowing $X$ give me any useful information about $u$?"
    - If yes: model is **endogenous**, **biased** and **not-causal**!

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Exogeneity and Unbiasedness

- $\hat{\beta_1}$ is .onfire[unbiased] iff no systematic difference, on average, between sample values of $\hat{\beta_1}$ and true population $\beta_1$, i.e.

$$E[\hat{\beta_1}]=\beta_1$$

--

- Does *not* mean any sample gives us $\hat{\beta_1}=\beta_1$, only the **estimation procedure** will, *on average*, yield the correct value

- Random errors above and below the true value cancel out (so that on average, $E[\hat{u}|X]=0)$

---

# Sidenote: Estimators of Statistics I

- In statistics, an .shout[estimator] is simply a rule that for calculating a statistic (often about a wider population parameter)

--

.content-box-green[
.green[**Example**]:
We want to estimate the average height (H) of U.S. adults (population) and have a random sample of 100 adults.]

--

- Calculate the mean height of our sample $(\bar{H})$ to estimate the true mean height of the population $(\mu_H)$

- $\bar{H}$ is an **estimator** of $\mu_H$

--

- There are many estimators we *could* use to estimate $\mu_H$
    - How about using the first value in our sample: $H_1$

---

# Sidenote: Estimators of Statistics II

.pull-left[

- What makes one estimator (e.g. $\bar{H}$) better than another (e.g. $H_1$)?<sup>.red[1]</sup>

1. .onfire[Biasedness]: does the estimator give us the true parameter *on average*?

2. .onfire[Efficiency]: an estimator with a smaller variance is better
]

.pull-right[
.center[
![](https://www.dropbox.com/s/3adddurpkp2k22o/biasvariability.png?raw=1)
]
]

.footnote[<sup>.red[1]</sup> Technically, we also care about <span class="onfire">consistency</span>: minimizing uncertainty about the correct value. The Law of Large Numbers, similar to CLT, permits this. We don't need to get too advanced about probability in this class.]

---

# Exogeneity and Unbiasedness I

.pull-left[

- $\mathbf{\hat{\beta_1}}$ is the .shout[Best Linear Unbiased Estimator (BLUE)] estimator of  $\mathbf{\beta_1}$ *when $X$ is exogenous*<sup>.red[1]</sup>

- No systematic difference, on average, between sample values of $\hat{\beta_1}$ and the true population $\beta_1$:

$$E[\hat{\beta_1}]=\beta_1$$

- Does *not* mean that each sample gives us $\hat{\beta_1}=\beta_1$, only the estimation *procedure* will, on average, yield the correct value

]

.pull-right[
.center[
![](https://www.dropbox.com/s/3adddurpkp2k22o/biasvariability.png?raw=1)
]
]

.footnote[<sup>.red[1]</sup> The proof for this is known as the [Gauss-Markov Theorem](https://en.wikipedia.org/wiki/Gauss%E2%80%93Markov_theorem). See today's [class notes](/class/10-class) for a simplified proof.]

---

# Exogeneity and Unbiasedness II

- Recall, an .shout[exogenous] variable $(X)$ is unrelated to other factors affecting $Y$, i.e.:

$$cor(X,u)=0$$

--

- Again, this is called the **Zero Conditional Mean Assumption**

$$E(u|X)=0$$

--

- For any known value of $X$, the expected value of $u$ is 0

- Knowing the value of $X$ must tell us *nothing* about the value of $u$ (anything else relevant to $Y$ other than $X$)

- We can then confidently assert causation: $X \rightarrow Y$

---

# Endogeneity and Bias

- Nearly all independent variables are .shout[endogenous], they **are** related to the error term $u$
$$cor(X,u)\neq 0$$

--

.content-box-green[
.green[**Example**]: Suppose we estimate the following relationship:

$$\text{Violent crimes}_t=\beta_0+\beta_1\text{Ice cream sales}_t+u_t$$

- We find $\hat{\beta_1}>0$

- Does this mean Ice cream sales$\rightarrow$Violent crimes?
]

---

# Endogeneity and Bias III

- The true expected value of $\hat{\beta_1}$ is actually:<sup>.red[1]</sup>

$$E[\hat{\beta_1}]=\beta_1+cor(X,u)\frac{\sigma_u}{\sigma_X}$$

--

- Takeaways:

--

1. If $X$ is exogenous: $cor(X,u)=0$, we're just left with $\beta_1$

--

2. The larger $cor(X,u)$ is, larger .onfire[bias]: $\left(E[\hat{\beta_1}]-\beta_1 \right)$

--

3. We can "sign" the direction of the bias based on $cor(X,u)$
    - **Positive** $cor(X,u)$ overestimates the true $\beta_1$ $(\hat{\beta_1}$ is too high)
    - **Negative** $cor(X,u)$ underestimates the true $\beta_1$ $(\hat{\beta_1}$ is too low)

.footnote[<sup>.red[1]</sup> See today's [class notes](/class/10-class) for proof.]

---

# Endogeneity and Bias: Example I

.content-box-green[
.green[**Example**]: 
$$wages_i=\beta_0+\beta_1 education_i+u$$

- Is this an accurate reflection of $educ \rightarrow wages$?

- Does $E[u|education]=0$?

- What would $E[u|education]>0$ mean?

]

---

# Endogeneity and Bias: Example II

.content-box-green[
.green[**Example**]: 
$$\text{per capita cigarette consumption}=\beta_0+\beta_1 \text{State cig tax rate}+u 	$$

- Is this an accurate reflection of $tax \rightarrow cons$?

- Does $E[u|tax]=0$?

- What would $E[u|tax]>0$ mean?

]

---

# Exogeneity and RCTs

.pull-left[

- Think about an idealized randomized controlled experiment

- Subjects randomly assigned to treatment or control group
    - Implies knowing whether someone is treated $(X)$ tells us nothing about their personal characteristics $(u)$

    - Random assignment makes $u$ independent of $X$, so

$$cor(X,u)=0\text{ and } E[u|X]=0$$

]

.pull-right[
.center[
![:scale 80%](https://www.dropbox.com/s/wqjisnpxaoaa2yy/rct.png?raw=1)
]

]