### 2019-10-01 static/slides/11-slides.Rmd

---
title: "2.5: OLS: Precision and Diagnostics"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

```{r regression-setup, echo=F, results="hide"}

library(haven)
CASchool<-read_dta("../data/caschool.dta")

# run regression of testscr on str
school_reg <- lm(testscr ~ str, 
                 data = CASchool)

library(broom)
school_reg_tidy <- tidy(school_reg,
     conf.int = TRUE) # add confidence intervals
CASchool_aug <- augment(school_reg)

library(equatiomatic)
extract_eq(school_reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)

```

# The Sampling Distribution of $\hat{\beta_1}$

.pull-left[

$$\hat{\beta_1} \sim N(E[\hat{\beta_1}], \sigma_{\hat{\beta_1}})$$

1. $E[\hat{\beta_1}]$; the **center** of the distribution (last class)
    - $E[\hat{\beta_1}]=\beta_1$<sup><span class="red">1</span></sup>

]

.pull-right[

```{r, fig.retina=3}
beta_dist_1<-ggplot(data = tibble(x=-4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, size=2, color="blue")+
  geom_label(x=1, y=dnorm(1), label=expression(sigma[hat(beta[1])]==1), color="blue")+
  geom_segment(aes(x=0,xend=0, y=0, yend=0.4), linetype="dashed")+
  scale_x_continuous(breaks = 0,
                     labels = expression(E(hat(beta[1]))))+
  labs(x = expression(hat(beta[1])),
       y = "Probability")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
beta_dist_1
```

]

---

# The Sampling Distribution of $\hat{\beta_1}$

.pull-left[

$$\hat{\beta_1} \sim N(E[\hat{\beta_1}], \sigma_{\hat{\beta_1}})$$

1. $E[\hat{\beta_1}]$; the **center** of the distribution (last class)
    - $E[\hat{\beta_1}]=\beta_1$<sup><span class="red">1</span></sup>

2. $\sigma_{\hat{\beta_1}}$; how **precise** is our estimate? (today)
    - <span class="shout">Variance $\sigma^2_{\hat{\beta_1}}$</span> or <span class="shout">standard error<sup><span class="red">+</span></sup> $\sigma_{\hat{\beta_1}}$</span>

]

.pull-right[

```{r, fig.retina=3}
beta_dist_1+
  stat_function(fun = dnorm, args=list(mean = 0, sd = 2), size=2, color="red")+
  geom_label(x=2, y=dnorm(2,0,2), label=expression(sigma[hat(beta[1])]==2), color="red")

```

]
.footnote[<sup>.red[1]</sup> Under the 4 assumptions about $u$ (particularly, $cor(X,u)=0)$.

<sup>.red[+]</sup> Standard **"error"** is the analog of standard *deviation* when talking about <br>
the *sampling distribution* of a sample statistic (such as $\bar{X}$ or $\hat{\beta_1})$.
]
---

class: inverse, center, middle

# Variation in $\hat{\beta_1}$

---

# What Affects Variation in $\hat{\beta_1}$

.pull-left[

$$var(\hat{\beta_1})=\frac{(SER)^2}{n \times var(X)}$$

$$se(\hat{\beta_1})=\sqrt{var(\hat{\beta_1})} = \frac{SER}{\sqrt{n} \times sd(X)}$$

]

.pull-right[

- Variation in $\hat{\beta_1}$ is affected by three things:

1. **Goodness of fit of the model $(SER)$**<sup>.red[+]</sup>
    - Larger $SER$ $\rightarrow$ larger $var(\hat{\beta_1})$
2. **Sample size, $n$**
    - Larger $n$ $\rightarrow$ smaller $var(\hat{\beta_1})$
3. **Variance of $X$**
    - Larger $var(X)$ $\rightarrow$ smaller $var(\hat{\beta_1})$

]

.footnote[<sup>.red[+]</sup> Recall from last class, the **S**tandard **E**rror of the **R**egression `\\(\hat{\sigma_u} = \sqrt{\frac{\sum \hat{u_i}^2}{n-2}}\\)`]

---

# Variation in $\hat{\beta_1}$: Goodness of Fit

.pull-left[

```{r, fig.retina=3}
data_1<-tibble(x=rnorm(50,5,1),
               u=rnorm(50,1,1),
               y=3+x+u)

sd_x_1<-lm(y~x, data = data_1) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(std.error)  %>%
  round(.,2) %>%
  as.character()

beta0_1<-lm(y~x, data = data_1) %>%
  tidy() %>%
  slice(1) %>% # get first row (which is intercept, beta 0)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

beta1_1<-lm(y~x, data = data_1) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

ser_1<-lm(y~x, data = data_1) %>%
  glance() %>%
  pull(sigma)  %>%
  round(.,2) %>%
  as.character()

ggplot(data = data_1)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_1, '~+', beta1_1, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_1)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_1)))+
  #geom_text(aes(x=5,y=13.5), label=bquote(sigma[hat(beta[1])] == .(sd_x_1)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With Better Fit",
       subtitle = expression(paste("Lower SER lowers variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

```{r, fig.retina=3}
data_2<-tibble(x=data_1$x,
               u=rnorm(50,1,3),
               y=3+x+u)

sd_x_2<-lm(y~x, data = data_2) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(std.error)  %>%
  round(.,2) %>%
  as.character()

beta0_2<-lm(y~x, data = data_2) %>%
  tidy() %>%
  slice(1) %>% # get first row (which is intercept, beta 0)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

beta1_2<-lm(y~x, data = data_2) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

ser_2<-lm(y~x, data = data_2) %>%
  glance() %>%
  pull(sigma)  %>%
  round(.,2) %>%
  as.character()

ggplot(data = data_2)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_2, '~+', beta1_2, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_2)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_2)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With Worse Fit",
       subtitle = expression(paste("Higher SER raises variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

---

# Variation in $\hat{\beta_1}$: Sample Size

.pull-left[

```{r, fig.retina=3}
# reuse data_1

ggplot(data = data_1)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_1, '~+', beta1_1, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_1)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_1)))+
  #geom_text(aes(x=5,y=13.5), label=bquote(sigma[hat(beta[1])] == .(sd_x_1)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With Fewer Observations",
       subtitle = expression(paste("Smaller n raises variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

```{r, fig.retina=3}
data_3<-tibble(x=rnorm(100,5,1),
               u=rnorm(100,1,1),
               y=3+x+u)

sd_x_3<-lm(y~x, data = data_3) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(std.error)  %>%
  round(.,2) %>%
  as.character()

beta0_3<-lm(y~x, data = data_3) %>%
  tidy() %>%
  slice(1) %>% # get first row (which is intercept, beta 0)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

beta1_3<-lm(y~x, data = data_3) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

ser_3<-lm(y~x, data = data_3) %>%
  glance() %>%
  pull(sigma)  %>%
  round(.,2) %>%
  as.character()

ggplot(data = data_3)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_3, '~+', beta1_3, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_3)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_3)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With More Observations",
       subtitle = expression(paste("Larger n lowers variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

---

# Variation in $\hat{\beta_1}$: Variation in $X$

.pull-left[

```{r, fig.retina=3}
# reuse data_3

sd_X_3<-data_3 %>%
  summarize(sd_x_3=sd(x)) %>%
  round(.,2) %>%
  as.character()

ggplot(data = data_3)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_1, '~+', beta1_1, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_1)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SD(X) ==', sd_X_3)))+
  geom_text(aes(x=7,y=3), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_1)))+
  #geom_text(aes(x=5,y=13.5), label=bquote(sigma[hat(beta[1])] == .(sd_x_1)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With More Variation in X",
       subtitle = expression(paste("Larger ", var(X), " lowers variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

```{r, fig.retina=3}
#subset data_3

data_4<-data_3 %>%
  filter(x>4.5, x<5.5)

sd_X_4<-data_4 %>%
  summarize(sd_x_4=sd(x)) %>%
  round(.,2) %>%
  as.character()

sd_x_4<-lm(y~x, data = data_4) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(std.error)  %>%
  round(.,2) %>%
  as.character()

beta0_4<-lm(y~x, data = data_4) %>%
  tidy() %>%
  slice(1) %>% # get first row (which is intercept, beta 0)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

beta1_4<-lm(y~x, data = data_4) %>%
  tidy() %>%
  slice(2) %>% # get second row (which is x coefficient, beta 1)
  pull(estimate)  %>%
  round(.,2) %>%
  as.character()

ser_4<-lm(y~x, data = data_4) %>%
  glance() %>%
  pull(sigma)  %>%
  round(.,2) %>%
  as.character()

ggplot(data = data_4)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_point(data = data_3, aes(x = x, y = y), color="blue", alpha=0.3)+
  geom_smooth(method = "lm", color = "red")+
  geom_text(aes(x=7,y=6), label=parse(text=paste('~hat(Y)==', beta0_4, '~+', beta1_4, '~X')))+
  geom_text(aes(x=7,y=5), label=parse(text=paste('~SER==', ser_4)))+
  geom_text(aes(x=7,y=4), label=parse(text=paste('~SD(X) ==', sd_X_4)))+
  geom_text(aes(x=7,y=3), label=parse(text=paste('~SE(hat(beta[1])) ==', sd_x_4)))+
  scale_x_continuous(breaks=seq(2,8,2),
                     limits=c(2,8))+
  scale_y_continuous(breaks=seq(3,15,3),
                     limits=c(3,15))+
  labs(x = "X",
       y = "Y",
       title = "Model With Less Variation in X",
       subtitle = expression(paste("Smaller ", var(X), " raises variation in ", hat(beta[1]))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

]

---

class: inverse, center, middle

# Presenting Regression Results

---

# Our Class Size Regression: Base R

.left-code[

```{r, echo=T}
summary(school_reg) # get full summary
```

]

.right-plot[

- How can we present all of this information in a tidy way?

]

---

# Our Class Size Regression: Broom I

- `broom`'s `tidy()` gives us a little bit neater table

```{r, echo=T}
library(broom)
tidy(school_reg)
```

---

# Our Class Size Regression: Broom II

- `broom`'s `glance()` gives us summary statistics about the regression

```{r, echo=T}
library(broom)
glance(school_reg)
```

---

# Presenting Regressions in a Table

.pull-left[

- Professional journals and papers often have a regression table, including:
  - Estimates of $\hat{\beta_0}$ and $\hat{\beta_1}$
  - Standard errors of $\hat{\beta_0}$ and $\hat{\beta_1}$ (often below, in parentheses)
  - Indications of statistical significance (often with asterisks)
  - Measures of regression fit: $R^2$, $SER$, etc

- Later: multiple columns for multiple models
]

.pull-right[
.font80[
.regtable[
```{r}
library(huxtable)
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

---

# Regression Output with huxtable I

.pull-left[

- You will need to first `install.packages("huxtable")`

- Load with `library(huxtable)`

- Command: `huxreg()`

- Main argument is the name of your `lm` object

- Default output is fine, but often we want to customize a bit

```{r, echo=T, eval=F}
# install.packages("huxtable")
library(huxtable)

huxreg(school_reg)
```

]

.pull-right[

.font80[
.regtable[
```{r}
huxreg(school_reg)
```
]
]
]

---

# Regression Output with huxtable II

- Can give title to each column

```{r, echo=T, eval=F}
"Test Score" = school_reg
```

--

- Can change name of coefficients from default

```{r, echo=T, eval=F}
coefs = c("Intercept" = "(Intercept)",
          "STR" = "str")
```

--

- Decide what statistics to include, and rename them

```{r, echo=T, eval=F}
statistics = c("N" = "N",
               "R-Squared" = "r.squared",
               "SER" = "sigma")
```

--

- Choose how many decimal places to round to

```{r, echo=T, eval=F}
number_format = 2
```

---

# Regression Output with huxtable III

.pull-left[

```{r, echo=T, eval=F}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

]

--

.pull-left[
.font80[
.regtable[
```{r}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

---

# Regression Outputs

- `huxtable` is one package you can use
    - See [here for more options](https://cran.r-project.org/web/packages/huxtable/vignettes/huxtable.html)

- I used to only use [`stargazer`](https://cran.r-project.org/web/packages/stargazer/index.html), but as it was originally meant for STATA, it has limits and problems
    - A great [cheetsheat](http://jakeruss.com/cheatsheets/stargazer.html) by my friend Jake Russ

---

class: inverse, center, middle

# Diagnostics about Regression

---

# Diagnostics: Residuals I

- We often look at the residuals of a regression to get more insight about its **goodness of fit** and its **bias**

- Recall `broom`'s `augment` creates some useful new variables
    - `.fitted` are fitted (predicted) values from model, i.e. $\hat{Y}_i$
    - `.resid` are residuals (errors) from model, i.e. $\hat{u}_i$

---

# Diagnostics: Residuals II

- Often a good idea to store in a new object (so we can make some plots)

```{r, echo=T}
aug_reg<-augment(school_reg)
aug_reg %>%
  head()
aug_reg<-augment(school_reg)
aug_reg %>% head()
```

---

# Recap: Assumptions about Errors

.pull-left[

- Recall the .shout[4 critical **assumptions** about `\\(u\\)`]:

1. The expected value of the residuals is 0
$$E[u]=0$$

2. The variance of the residuals over $X$ is constant, written:
$$var(u|X)=\sigma^2_{u}$$

3. Errors are not correlated across observations: 
$$cor(u_i,u_j)=0 \quad \forall i \neq j$$

4. There is no correlation between $X$ and the error term: 
$$cor(X, u)=0 \text{ or } E[u|X]=0$$

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Assumptions 1 and 2: Errors are i.i.d.

- Assumptions 1 and 2 assume that errors are coming from the same (*normal*) distribution
$$u \sim N(0, \sigma_u)$$
  - Assumption 1: $E[u]=0$
  - Assumption 2: $sd(u|X)=\sigma_u$
      - virtually always unknown...

- We often can visually check by plotting a **histogram** of $u$

---

# Plotting Residuals

.pull-left[

```{r residual-hist, echo=T, eval=F}
ggplot(data = aug_reg)+
  aes(x = .resid)+
  geom_histogram(color="white")+
  labs(x = expression(paste("Residual, ", hat(u))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

--

.pull-right[

```{r, ref.label="residual-hist", fig.retina=3}
```

]

---

# Plotting Residuals

.pull-left[

```{r residual-hist-2, echo=T, eval=F}
ggplot(data = aug_reg)+
  aes(x = .resid)+
  geom_histogram(color="white")+
  labs(x = expression(paste("Residual, ", hat(u))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

- Just to check:

```{r, echo=T}
aug_reg %>%
  summarize(e_u = mean(.resid),
            sd_u = sd(.resid))
```

]

--

.pull-right[

```{r, ref.label="residual-hist-2", fig.retina=3}
```

]

---

# Residual Plot

.pull-left[

- We often plot a .shout[residual plot] to see any odd patterns about residuals
    - $x$-axis are $X$ values (`str`)
    - $y$-axis are $u$ values (`.resid`)

```{r residual-plot, eval=F}
ggplot(data = aug_reg)+
  aes(x = str,
      y = .resid)+
  geom_point(color="blue")+
  geom_hline(aes(yintercept = 0), color="red")+
  labs(x = "Student to Teacher Ratio",
       y = expression(paste("Residual, ", hat(u))))+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

--

.pull-right[

```{r, ref.label="residual-plot", fig.retina=3}
```

]

---

class: inverse, center, middle

# Problem: Heteroskedasticity

---

# Homoskedasticity

.pull-left[

- ".shout[Homoskedasticity]:" variance of the residuals over $X$ is constant, written:
$$var(u|X)=\sigma^2_{u}$$

- Knowing the value of $X$ does not affect the variance (spread) of the errors

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Heteroskedasticity I

.pull-left[

- ".shout[Heteroskedasticity]:" variance of the residuals over $X$ is *NOT* constant:
$$var(u|X) \neq \sigma^2_{u}$$

- **This does not cause $\hat{\beta_1}$ to be biased**, but it does cause the standard error of $\hat{\beta_1}$ to be incorrect

- This **does** cause a problem for .onfire[inference]!
]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

# Heteroskedasticity II

- Recall the formula for the standard error of $\hat{\beta_1}$:

$$se(\hat{\beta_1})=\sqrt{var(\hat{\beta_1})} = \frac{SER}{\sqrt{n} \times sd(X)}$$

- This actually *assumes* homoskedasticity

---

# Heteroskedasticity III

- Under heteroskedasticity, the standard error of $\hat{\beta_1}$ mutates to:

$$se(\hat{\beta_1})=\sqrt{\frac{\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2\hat{u}^2}{\big[\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2\big]^2}}$$

- This is a .shout[heteroskedasticity-robust] (or just .shout["robust"]) method of calculating $se(\hat{\beta_1})$

- Don't learn formula, **do learn what heteroskedasticity is and how it affects our model!**

---

# Visualizing Heteroskedasticity I

.pull-left[

- Our original scatterplot with regression line
]

.pull-right[

```{r, fig.height=3.5, fig.retina=3}
ggplot(data = CASchool)+
  aes(x = str,
      y = testscr)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=16)
```
]

---

# Visualizing Heteroskedasticity I

.pull-left[

- Our original scatterplot with regression line

- Does the spread of the errors change over different values of $str$?
    - No: homoskedastic
    - Yes: heteroskedastic
]

.pull-right[

```{r, fig.height=3.5, retina=3}
ggplot(data = CASchool)+
  aes(x = str,
      y = testscr)+
  geom_point(color="blue")+
  geom_vline(xintercept=seq(14,26,2), linetype="dashed", size=1)+
  geom_smooth(method="lm", color="red")+
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=16)
```
]

---

# Visualizing Heteroskedasticity I

.pull-left[

- Our original scatterplot with regression line

- Does the spread of the errors change over different values of $str$?
    - No: homoskedastic
    - Yes: heteroskedastic
]

.pull-right[

```{r, fig.height=3.5, retina=3}
ggplot(data = CASchool)+
  aes(x = str,
      y = testscr)+
  geom_point(color="blue")+
  geom_vline(xintercept=seq(14,26,2), linetype="dashed", size=1)+
  geom_smooth(method="lm", color="red")+
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=16)
```

```{r, fig.height=3.5, retina=3}
aug_reg_het<-aug_reg %>%
  mutate(range = case_when(str>=14 & str<16 ~ "14 to 16",
                           str>=16 & str<18 ~ "16 to 18",
                           str>=18 & str<20 ~ "18 to 20",
                           str>=20 & str<22 ~ "20 to 22",
                           str>=22 & str<24 ~ "22 to 24",
                           str>=24 & str<26 ~ "24 to 26"),
         range = factor(range, levels = c("14 to 16", "16 to 18", "18 to 20", "20 to 22", "22 to 24", "24 to 26"))) # needs to be a factor in order to plot in order

aug_reg_het_sigmas<-aug_reg_het %>%
  group_by(range) %>%
  summarize(sigmas = as.character(round(sd(.resid),2))) %>%
  slice(1:6) # remove NA row 7

ggplot(data = aug_reg_het)+
  aes(x = .resid)+
  geom_density(aes(fill=range), alpha=0.5)+
  geom_label(data=aug_reg_het_sigmas,
             aes(x=0,y=0.005,
                 label=sigmas,
                 color=range),size=5)+
                 #label=parse(text=paste('~sigma==', sigmas))))+
  facet_grid(~range)+
  guides(fill = F,
         color = F)+
  labs(x = "Residuals by STR",
       y = "Test Score")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=16)
```

]

---

# More Obvious Heteroskedasticity

.pull-left[

- Visual cue: data is "fan-shaped"
    - Data points are closer to line in some areas
    - Data points are more spread from line in other areas
]

.pull-right[

```{r, fig.height=3.5, fig.retina=3}
het_data<-tibble(x = runif(500,0,10),
                 y = rnorm(500,x,x))

ggplot(data = het_data)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  scale_x_continuous(breaks=seq(0,10,1))+
    labs(x = "X",
       y = "Y")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```
]

---

# More Obvious Heteroskedasticity

.pull-left[

- Visual cue: data is "fan-shaped"
    - Data points are closer to line in some areas
    - Data points are more spread from line in other areas
]

.pull-right[

```{r, fig.height=3.5, fig.retina=3}
ggplot(data = het_data)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  geom_vline(xintercept=seq(0,10,2), linetype="dashed", size=1)+
  scale_x_continuous(breaks=seq(0,10,1))+
    labs(x = "X",
       y = "Y")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

het_reg<-lm(y~x, data = het_data)
aug_het_reg<-het_reg %>% augment()

aug_het_reg<-aug_het_reg %>%
  mutate(range = case_when(x>=0 & x<2 ~ "0 to 2",
                           x>=2 & x<4 ~ "2 to 4",
                           x>=4 & x<6 ~ "4 to 6",
                           x>=6 & x<8 ~ "6 to 8",
                           x>=8 & x<=10 ~ "8 to 10"),
         range = factor(range, levels = c("0 to 2", "2 to 4", "4 to 6", "6 to 8", "8 to 10"))) # needs to be a factor to be in order in the plot

aug_het_reg_sigmas<-aug_het_reg %>%
  group_by(range) %>%
  summarize(sigmas = as.character(round(sd(.resid),2))) %>%
  slice(1:6) # remove NA row 7

#facets$new = factor(temp$type, levels=c("T","F","P"), labels=c("T","F","P"))

ggplot(data = aug_het_reg)+
  aes(x = .resid)+
  geom_density(aes(fill=range), alpha=0.5)+
  geom_label(data=aug_het_reg_sigmas,
             aes(x=0,y=0.3,
                 label=sigmas,
                 color=range),size=5)+
                 #label=parse(text=paste('~sigma==', sigmas))))+
  facet_grid(~range)+
  guides(fill = F,
         color = F)+
  labs(x = "Residuals by Range of X",
       y = "Density")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```
]

---

# What Might Cause Heteroskedastic Errors?

.pull-left[

$$\widehat{wage_i}=\hat{\beta_0}+\hat{\beta_1}educ_i$$

```{r}
library(wooldridge)
wage_reg<-lm(wage~educ, data=wage1)
huxreg("Wage" = wage_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "Years of Schooling" = "educ"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]

.pull-right[

```{r, fig.height=3.5, fig.retina=3}

wage_plot<-ggplot(data = wage1)+
  aes(x = educ,
      y = wage)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  scale_x_continuous(breaks=seq(0,20,2))+
  scale_y_continuous(labels=scales::dollar)+
    labs(x = "Years of Schooling",
       y = "Hourly Wage")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
wage_plot
```
]

---

# What Might Cause Heteroskedastic Errors?

.pull-left[

$$\widehat{wage_i}=\hat{\beta_0}+\hat{\beta_1}educ_i$$

```{r}
library(wooldridge)
wage_reg<-lm(wage~educ, data=wage1)
huxreg("Wage" = wage_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "Years of Schooling" = "educ"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]

.pull-right[

```{r, fig.height=3.5, fig.retina=3}

wage_reg <- lm(wage ~ educ, data = wage1)
aug_wage_reg<-wage_reg %>%
  augment()
wage_plot+geom_vline(xintercept=c(0,4,8,12,16,18), linetype="dashed", size=1)

```

```{r, fig.height=3.5, fig.retina=3}
aug_wage_reg<-aug_wage_reg %>%
  mutate(range = case_when(educ>=0 & educ<4 ~ "0 to 4",
                           educ>=4 & educ<8 ~ "4 to 8",
                           educ>=8 & educ<12 ~ "8 to 12",
                           educ>=12 & educ<16 ~ "12 to 16",
                           educ>=16 & educ<20 ~ "16 to 18"),
         range = factor(range, levels = c("0 to 4", "4 to 8", "8 to 12", "12 to 16", "16 to 18"))) # needs to be a factor to be in order in the plot

aug_wage_reg_het_sigmas<-aug_wage_reg %>%
  group_by(range) %>%
  summarize(sigmas = as.character(round(sd(.resid),2))) %>%
  slice(1:6) # remove NA row 7

#facets$new = factor(temp$type, levels=c("T","F","P"), labels=c("T","F","P"))

ggplot(data = aug_wage_reg)+
  aes(x = .resid)+
  geom_density(aes(fill=range), alpha=0.5)+
  geom_label(data=aug_wage_reg_het_sigmas,
             aes(x=5,y=0.3,
                 label=sigmas,
                 color=range),size=5)+
                 #label=parse(text=paste('~sigma==', sigmas))))+
  facet_grid(~range)+
  guides(fill = F,
         color = F)+
  labs(x = "Residuals by Years of Schooling",
       y = "Density")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

```

---

# Detecting Heteroskedasticity I

- Several tests to check if data is heteroskedastic 
- One common test is **Breusch-Pagan test** 
- Can use `bptest()` with `lmtest` package in `R`
    - $H_0$: homoskedastic
    - If $p$-value < 0.05, reject $H_0\implies$ heteroskedastic

--

.pull-left[
```{r,eval=F, echo=T}
# install.packages("lmtest")
library("lmtest")
bptest(school_reg)
```
]

.pull-right[

```{r}
library("lmtest") #load lmtest package, install if first time
bptest(school_reg)
```
]

---

# Detecting Heteroskedasticity II

- How about our wage regression?

.pull-left[
```{r,eval=F, echo=T}
# install.packages("lmtest")
library("lmtest")
bptest(wage_reg)
```
]

.pull-right[

```{r}
library("lmtest") #load lmtest package, install if first time
bptest(wage_reg)
```
]

---

# Fixing Heteroskedasticity I

- Heteroskedasticity is easy to fix with software that can calculate .shout[robust] standard errors (using the more complicated formula above)

- Easiest method is to use `estimatr` package
  - `lm_robust()` command (instead of `lm`) to run regression
  - set `se_type="stata"` to calculate robust SEs using the formula above

```{r}
#install.packages("estimatr")
library(estimatr)

school_reg_robust <-lm_robust(testscr ~ str, data = CASchool,
                              se_type = "stata")

school_reg_robust
```

---

# Fixing Heteroskedasticity II

.pull-left[
```{r, echo=T, eval=F}
library(huxtable)
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)

```
]

.pull-right[
.font80[
.regtable[
```{r}
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

---

# Assumption 3: No Serial Correlation

.pull-left[

- Errors are not correlated across observations: 
$$cor(u_i,u_j)=0 \quad \forall i \neq j$$

- For simple cross-sectional data, this is rarely an issue

- Time-series & panel data nearly always contain .shout[serial correlation] or .shout[autocorrelation] between errors

- Errors may be .shout[clustered]
    - **by group**: e.g. all observations from Maryland, all observations from Virginia, etc.
    - **by time**: GDP in 2006 around the world, GDP in 2008 around the world, etc.

- We'll deal with these fixes when we talk about panel data (or time-series if necessary)

]

.pull-right[

.center[
![](https://www.dropbox.com/s/95rh9ow982y5htb/error.png?raw=1)
]
]

---

class: inverse, center, middle
# Outliers

---

# Outliers Can Bias OLS! I

.pull-left[

- .shout[Outliers] can affect the slope (and intercept) of the line and add .shout[bias]
    - May be result of human error (measurement, transcribing, etc)
    - May be meaningful and accurate

- In any case, compare how including/dropping outliers affects regression and always discuss outliers!

]

.pull-right[

```{r, fig.retina=3}
#Just to simplify the dataset, let's make a new, smaller data frame with just the variables: observation, district, testscr, and str 
CA.simple<-subset(CASchool,select=c(observat,district,testscr,str))

#Now let's create two crazy outliers 
outlier<-c(421,"Terrible School 1",800,30)
outlier2<-c(422,"Terrible School 2",850,28)
outlier3<-c(423,"Terrible School 2",820,29)

#And add them to a new dataframe 
CA.outlier<-rbind(CA.simple,outlier,outlier2,outlier3)
CA.outlier$str<-as.numeric(CA.outlier$str) #dont know why it changes class to character, change to numeric
CA.outlier$testscr<-as.numeric(CA.outlier$testscr)
outliers<-subset(CA.outlier, testscr>790)

#Create a scatterplot with the outlier 
ggplot(CA.simple, aes(str,testscr))+
  geom_point(color="blue")+ #the normal data points
  geom_smooth(data=CA.simple,method=lm, color="red")+ #the regression line
  geom_point(data=outliers, color="magenta")+ #the outliers (in magenta)
  geom_smooth(data=CA.outlier, method=lm,color="purple")+ #the regression line with outliers
  xlab("Student to Teacher Ratio")+ylab("Test Score")+
    theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)

```
]
---

# Outliers Can Bias OLS! II

.font80[
.regtable[
```{r}
school_outlier_reg<-lm(testscr~str, data=CA.outlier)

huxreg("No Outliers" = school_reg,
       "Outliers" = school_outlier_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]

---

# Detecting Outliers

- The `car` package has an `outlierTest` command to run on the regression

```{r, echo=T}
library("car")

# Use Bonferonni test 
outlierTest(school_outlier_reg) # will point out which obs #s seem outliers
```