### 2019-10-03 static/r/11-practice-answers.Rmd

---
title: "2.4 OLS: Precision and Diagnostics - R Practice (Answers)"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output:
  html_document:
    df_print: paged
    theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

First things first, load `tidyverse`.

```{r}
library("tidyverse")
```

```{r}
congress<-read_csv("../data/congress-trump-score.csv")
```

```{r}
congress %>%
  count(party)
```

```{r}
glimpse(congress)
```

```{r}
congress %>%
  count(congress)
```

```{r}
congress_tidy <- congress %>%
  filter(congress!=0) %>%
  mutate(agree_pct = agree_pct * 100)
```

```{r}
congress_tidy %>%
  mutate(pol_party = case_when(
    party %in% c("Democrat", "D") ~ "Democrat",
    party %in% c("Republican", "R") ~ "Republican",
    party %in% c("Independent", "I") ~ "Independent"
  )) %>%
  count(pol_party)

congress_tidy <- congress_tidy %>%
  mutate(pol_party = case_when(
    party %in% c("Democrat", "D") ~ "Democrat",
    party %in% c("Republican", "R") ~ "Republican",
    party %in% c("Independent", "I") ~ "Independent"
  ))

```

# Warm Up to `dplyr` With `gapminder`

## 
#### 1. Load `gapminder`. `select()` the variables `year`, `lifeExp`, `country`.

```{r}

party_colors <- c("Democrat" = "blue", "Republican" = "red", "Independent" = "gray")

ggplot(data = congress_tidy)+
  aes(x = net_trump_vote, y=agree_pct)+
  geom_point(aes(color = pol_party))+
  geom_smooth(method="lm", color = "black")+
  scale_colour_manual("Parties", values = party_colors)+
  theme_classic()
```

```{r}
congress_tidy %>%
  summarize(cor = cor(net_trump_vote, agree_pct))

congress_cor<-congress_tidy %>%
  select(agree_pct, net_trump_vote)
```

```{r}
reg<-lm(agree_pct ~ net_trump_vote, data = congress_tidy)
summary(reg)
```

```{r}
library(broom)
reg %>%
  tidy()
```

```{r, results="asis"}
library(equatiomatic)
extract_eq(reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)
```

```{r}
reg %>%
  glance()
```

```{r}
reg_aug<-reg %>%
  augment()
```

```{r}
reg_aug %>%
  head()
```

```{r}
ggplot(data = reg_aug)+
  aes(x = net_trump_vote, y=.resid)+
  geom_point()+
  geom_hline(yintercept = 0, color = "red")+
  theme_classic()
```

```{r}
ggplot(data = reg_aug)+
  aes(x = .resid)+
  geom_histogram(color="white")+
  theme_classic()
```

```{r}
congress %>%
  count(chamber)
```

```{r}
ggplot(data = reg_aug)+
  aes(x = .resid)+
  geom_density()
```

```{r}
library(car)
outlierTest(reg)
```

```{r}
congress_tidy %>%
  slice(1070)
```

```{r}
reg_dfbetas<-as.tibble(dfbetas(reg))#

reg_dfbetas
```

```{r}
reg_dfbetas %>%
  arrange(desc(net_trump_vote))
```

# Check for Heteroskedasticity

```{r}
library(lmtest)
bptest(reg)
```