### 2019-10-03 content/reading/12-reading.Rmd
There's no practice for class 12, shorten slides 12

---
title: "2.6 Inferential Statistics I - Readings"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
---

# Strongly Recommended

- <i class="fas fa-book"></i> [Ch. 4] in @Bailey
- [<i class="fas fa-external-link-square-alt"></i> Ch. 10 (optionally 8-9) in *Modern Dive* ](https://moderndive.netlify.com/10-inference-for-regression.html)
- [<i class="fab fa-youtube"></i> *Last Week Tonight with John Oliver*: Scientific Studies](https://www.youtube.com/watch?v=0Rnq1NpHdmw)
- [<i class="fas fa-external-link-square-alt"></i> Visualizing Probability and Inference ](https://seeing-theory.brown.edu/frequentist-inference/index.html)

Bailey teaches inferential statistics in the classical way (with reference to theoretical $Z$ and $t$ distributions, and $Z$ and $t$ tests). This is all valid. Again, you may wish to brush up with [Khan Academy](https://www.khanacademy.org/math/statistics-probability)^[From sampling distributions through significance tests, for this. Though the whole class is helpful!].

The latter "book" (also free online, like R4DS) uses the `infer` package to run simulations for inferential statistics. Chapter 10 is focused on regression (but I also recommend the chapters leading up to it, which are on inferential statistics broadly, using this method).

Also be sure to watch the excellent and hilarious discussion of the limits and misuses of scientific studies and statistical significance $(p$-values) in the *Last Week Tonight* clip.

The final link is a great website for visualizing basic statistic concepts like probability, distributions, confidence intervals, hypothesis tests, central limit theorem, and regression.