### 2019-10-10 static/slides/midterm-review.Rmd
Tweak review slides

---
title: "Midterm Review Questions"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```
# Question 1

What does **endogenous** mean, in words? What about statistically?

---

# Question 2

If a regression is biased (from endogeneity), what can we learn about the bias?

---

# Question 3

What does heteroskedasticity mean? Does heteroskedasticity **bias** $\hat{\beta_1}$?

---

# Question 4

Is this data likely heteroskedastic or homoskedastic?

```{r}
library(tidyverse)
het_data<-tibble(x = runif(500,0,2),
                 y = rnorm(500,x,x))

ggplot(data = het_data)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
  scale_x_continuous(breaks=seq(0,10,1))+
    labs(x = "X",
       y = "Y")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)
```

---

# Question 5

- What three things impact the variation of $\hat{\beta_1}$? How?

---

# Question 6

What are the four assumptions we make about the error term?

Which is most important?

---

# Question 7

$$Wages_i=\hat{\beta_0}+\hat{\beta_1}Education+u_i$$

What is in $u_i$?

Is $\hat{\beta_1}$ likely biased?

---

# Question 8

What does $R^2$ measure? What does it mean? How do we calculate it?

---

# Question 9

What does $\sigma_u$ (SER) measure? What does it mean?

---

# Question 10

Interpret all of these numbers (except Adjusted R-squared and the last line):

```{r}
summary(lm(y~x, data = het_data))
```

---

# Question 11

Interpret all of these numbers:

```{r}
reg<-lm(y~x, data = het_data)
library(huxtable)
huxreg(reg,        statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"))
```

---

# Question 12

Suppose $Y$ is normally distributed with a mean of 10 and a standard error of 5. What is the probability that $Y$ is between 5 and 15?

---

# Question 13

Explain what a $Z$-score means.

---

# Question 14

Explain what a $p$-value means. :)

---

# Question 15

We run the following hypothesis test at $\alpha=0.05$:

$$\begin{align*}
H_0: \, & \beta_1=0\\
H_1: \, & \beta_1 \neq 0 \\
\end{align*}$$

Is this test one-sided or two-sided?

We find the $p$-value is 0.02. What is our conclusion? Be specific and precise in your wording!