### 2019-10-29 static/slides/15-slides.Rmd

---
title: "3.3: Model Specification Strategies"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

```{r regression-setup, echo=F, results="hide"}

library(haven)
CASchool<-read_dta("../data/caschool.dta")

# run regression of testscr on str
school_reg <- lm(testscr ~ str, 
                 data = CASchool)

library(broom)
school_reg_tidy <- tidy(school_reg,
     conf.int = TRUE) # add confidence intervals
CASchool_aug <- augment(school_reg)

library(equatiomatic)
extract_eq(school_reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)

```

# Model Specification

- The big challenge in applied econometrics is choosing how to **specify a model** to regress

- Every dataset is different, every study has a different goal
    - there is no bright line rule, only a set of guidelines and skills that you can only learn by doing!

- But here are some helpful tips and frequent problems (and solutions)

---

# Model Specification: Process

**1. Identify your question of interest:** what do you want to know? What marginal effect(s) do you want to estimate?

--

**2. Think about possible sources of endogeneity:** what *other* variables would cause  .shout[omitted variable bias] if we left them out? Can we get data on them too? 
  - Again: must BOTH (1) affect $Y$ AND (2) be correlated with $X$
  - This requires much of your economic intuitions: $R^2$ and statistical measures cannot tell you everything!

--

**3. Run multiple models and check the robustness of your results:** does the size (or direction) of your marginal effect(s) of interest change as you change your model (i.e. add more variables)?

--

**4. Interpret your results**
  - Are they statistically significant? 
  - Regardless of *statistical* significance, are they **economically meaningful**? 
      - Why should we care? 
      - How big is "big"?

---

# Proxy Variables

- Ideally, we would want a randomized control experiment to assign individuals to treatment

--

- But with observational data, $u_i$ depends on other factors
    - *If* we can observe and measure these factors, then include them in the regression

--

- If we can't directly measure them, often we can include variables *correlated* with these variables to .shout[proxy] for the effects of them!

---

# Proxy Variables: Example I

.content-box-green[
.green[**Example**]: Consider test scores and class sizes again. What about learning opportunities outside of school?
]

--

- Probably a bias-causing omitted variable (affects test score and correlated with class size) but we can't measure it!

--

- Suppose we *can* measure a variable $V$, and significantly, $corr(V,Z)\neq 0$
  - e.g. we have data on the percent of students who get a free or subsidized lunch (`meal_pct`)

--

- This is a good .shout[proxy] for income-determined learning opportunities outside of school
  - `meal_pct` is correlated with Income (which is likely correlated with both class size and test score)
  - A good *indirect* measure of Income

---

# Proxy Variables: Example II

- We assumed we don't have data on average district income, we would expect `meal_pct` to be strongly negatively correlated with income

--

- Just kidding, we do have data on `avginc`, but we'll only use it to confirm our suspicion:

--

```{r, echo=T}
CASchool %>%
  select(testscr, str, el_pct, avginc,meal_pct) %>%
  cor()
```

--

- We can see `meal_pct` is strongly (negatively) correlated with income, as expected

---

# Proxy Variables: Example III

```{r, echo=T}
proxyreg<-lm(testscr~str+el_pct+meal_pct, data=CASchool)
summary(proxyreg)
```

---

# Interpreting Control Variables

$$\widehat{\text{Test Score}}_i=700.15-1.00 \, STR_i-0.122 \, \%EL_i - 0.547 \, meal\%_i$$

--

- Is `meal%` causal?

--

- Getting rid of programs in districts where a large percentage of students need them would boost test scores A LOT! (So probably not causal...)

--

- `meal%` likely correlated with other things in $u_i$ (like *outside learning opportunities*!). 
    - In fact, that's *exactly why* we included it as a variable!

--

- .whisper[We don't need the OLS estimate on `meal%` to be unbiased!]
    - **We only care about getting the estimate on `str` to be unbiased!**

---

# Control Variables

- A .shout[control variable] is a regressor variable **note** of interest, but included to hold factors constant that, if omitted, would bias the causal effect of interest

--

- The control variable may still be correlated with omitted causal factors in $u$
    - Estimators $(\hat{\beta}$'s) on control variables can be biased and that is OK!
    - So long as we have unbiased estimators $(\hat{\beta}$'s) on the regressors we do care about!

--

- Control variables allow us to proceed *as if* $X$ were randomly assigned

---

# Coefficient Plots I

- Sometimes it is helpful to make a **coefficient plot** to visualize the different marginal effects

--

- Take `broom` and `tidy()` your regression with `conf.int=TRUE` (and save as an object)

```{r, echo=T}
# run regression
reg<-lm(testscr~str+el_pct+meal_pct, data=CASchool)

# load broom
library(broom)
#  tidy regression with confidence intervals
# save as "reg_tidy"
reg_tidy <- tidy(reg, conf.int = TRUE)
```

---

# Coefficient Plots II

```{r, echo=T}
# look at it
reg_tidy

# intercept is too big relative to others! let's leave it out
reg_tidy_no_int<-reg_tidy %>%
  filter(term!="(Intercept)")
```

---

# Coefficient Plots III

.pull-left[

```{r coef-plot, echo=T, eval=F}
ggplot(data = reg_tidy_no_int)+
  aes(x = estimate,
      y = term,
      color = term)+
    geom_point()+ # point for estimate
    # now make "error bars" using conf. int.
    geom_segment(aes(x = conf.low,
                     xend = conf.high,
                     y=term,
                     yend=term))+
  # add line at 0
  geom_vline(xintercept=0, size=1, color="black", linetype="dashed")+
  scale_x_continuous(breaks=seq(-1.5,0.5,0.5),
                     limits=c(-1.5,0.5))+
  labs(x = expression(paste("Marginal effect, ", hat(beta[j]))),
       y = "Variable")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

```{r, ref.label="coef-plot"}

```
]

---

# Model Specification Warning: $R^2$

.red[Do NOT just try to maximize `\\(R^2\\)` or `\\(\bar{R}^2\\)`]

- A high $R^2$ or $\bar{R}^2$ means that the regressors explain variation in $Y$

- A high $R^2$ or $\bar{R}^2$ does *NOT* mean you have eliminated omitted variable bias

- A high $R^2$ or $\bar{R}^2$ does *NOT* mean you have an unbiased estimate of a causal effect

- A high $R^2$ or $\bar{R}^2$ does *NOT* mean included variables are statistically significant

---

# Model Specification: Goal

- We care mostly about measuring a causal effect

- This involves making the respective $\hat{\beta_j}$'s as unbiased as possible (because $\beta_j$ is the causal effect of $X_j \rightarrow Y$!

- You will not get high $R^2$'s in social science
  - even if you do, you are only explaining *variation* in $Y$ with the model, not necessarily the *causal effect*!