### 2019-10-31 content/class/15-class.Rmd
Add class 16 content

---
title: "3.3: Model Specification Strategies - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, October 29, 2019*

# Overview

Today we will simply talk about the process of specifying a model: how to pick $X$ variables to add to your regression. Recall, they should fit the two criteria for omitted variable bias. One additional issue we discuss is the use of **proxy** variables: where you know omitting a variable would cause bias, but you don't have data on it, just data on a variable *related* to it; and of **control** variables (including proxies), where the OLS estimator $(\hat{\beta_j})$ for an independent variable is *not causal*, but just helps remove bias on the independent variables you *do* care about being causal.

For the rest of class, we will work on practice problems for multivariate regression.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/15-slides.html)

# Practice Problems

Today you will be working on [R practice problems](/r/15-r-practice.html). Check back later for solutions.