### 2019-10-31 content/class/16-class.Rmd

---
title: "3.4: Multivariate Regression Examples - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Thursday, October 31, 2019*

# Overview

Today we will continue working on practice problems from last class. Afterwards, we will explore some actual empirical economics papers that use (relatively) simple multivariate regressions. See them in today's [Readings page](/readings/16-reading).

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/16-slides.html)

# Practice Problems

Today you will finish working on [R practice problems](/r/15-r-practice.html) from last class. Check back later for solutions.