### 2019-10-31 static/slides/16-slides.Rmd

---
title: "3.4: Examples"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

class: inverse, center, middle

# Sacerdote (2001) on Peer Effects

---

# Paper Motivation I

.pull-left[

- What determines student outcomes in college? (GPAs, fraternity enrollment, alcohol/drug use, etc)

- Effects of peer groups

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

# Paper Motivation II

.pull-left[

- "Standard" way to estimate peer effects: regress student $i$'s outcomes/behavior on *other students'* outcomes/behavior

$$GPA_i = \beta_0+\beta_1 OwnBehavior_i + \beta_2 RoommateBehavior_i + u_i$$

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

# Paper Motivation II

.pull-left[

- "Standard" way to estimate peer effects: regress student $i$'s outcomes/behavior on *other students'* outcomes/behavior

$$GPA_i = \beta_0+\beta_1 OwnBehavior_i + \beta_2 RoommateBehavior_i + u_i$$

- Problems with this approach: 
    1. Individuals **self-select** into peer groups
    2. If two roommates $A$ and $B$ influence each other, how do we disentangle causal effect of $B \rightarrow A$ vs. $A \rightarrow B$?
    3. Are peer effects actually driven by students' own backgrounds, or by their actual choices?

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

# Paper Motivation II

.pull-left[

$$\begin{align*}
			corr(OwnBehavior, u) &\neq 0	\\
			corr(RoomateBehavior, u)& \neq 0 \\
			E[u|OwnBehavior,RoommateBehavior] & \neq 0 	\\
\end{align*}$$

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

# Sacerdotes' Identification Strategy

.pull-left[

- Freshmen entering Dartmouth College are **randomly** assigned to dorms & roommates

- Removes self-selection of peer groups by shared characteristics

- **Random assignment**: roommate A's background characteristics are uncorrelated with roommate B's background characteristics

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

# Relevant Institutions

- Freshmen entering Dartmouth College are randomly assigned to dorms & roommates
- Each incoming freshman fills out a questionnaire:
    1. I smoke. (Y/N)
    2. I like to listen to music while studying. (Y/N)
    3. I keep late hours. (Y/N)
    4. I am more neat than messy. (Y/N)
    5. I am (Male/Female). 	
- There are $2^5=32$ combinatorial possibilities of answers to the questions
- Students are assigned to roommates/dorms **at random**, conditional on their combination of answers to the 5 survey answers

---

# The Data

- Data from Dartmouth's database of students: history of dorm assignments & term-by-term academic performance

- Data on pre-treatment characteristics (SAT scores, high school class rank, private/public HS, home state, academic index)

- Outcome variables: GPA, time to graduation, frat membership, major choice, participation in athletics

- Survey of Incoming Freshman: if student drank beer in last year and expectation of graduating with honors

- Sample of 1589 students

- Create dummy variable for each block to control for covariates (we'll talk later about **dummy variables** and **fixed effects** like this)

---

# The Data: Summary Statistics I

.center[
![](https://www.dropbox.com/s/3fuhj7kbz0ptnnf/sacerdote1.png?raw=1)
]

---

# The Data: Summary Statistics II

.center[
![](https://www.dropbox.com/s/uhfupj0mg9p8m0h/sacerdote2.png?raw=1)
]

---

# The Data: Demonstrating Random Assignment

.center[
![](https://www.dropbox.com/s/bfw0v2wqnqw3j89/sacerdote3.png?raw=1)
]

---

# The Data: Demonstrating Random Assignment

.center[
![](https://www.dropbox.com/s/ull5y5x691hn54r/sacerdote3c.png?raw=1)
]

---

# Sacerdote's Empirical Model (Basically)

$$GPA_i=\beta_0+\beta_1ACA_i+\beta_2ACA_j+u_i$$

- Student $i$ and roommate $j$

- $ACA$: Index of academic performance (broken down into different metrics)

- Other outcomes of interest (besides $GPA$: graduation, major, fraternity, athlete

---

# Regression Results

.center[
![](https://www.dropbox.com/s/jwpt4n37qq0fr3h/sacerdote5.png?raw=1)
]

--

- For every 1 point increase (decrease) in your *roommate*'s GPA, *your* GPA increases (decreases) about 0.12 points

--

- If you would have been a 3.0 student with a 3.0 roommate, but you were assigned to a 2.0 roommate, your GPA would be 2.88

---

# Regression Results: Academic Performance

.center[
![](https://www.dropbox.com/s/d2dnjjxbivd9e8a/sacerdote6.png?raw=1)
]

---

# Regression Results: Social Outcomes

.center[
![](https://www.dropbox.com/s/1p0jdtio7qjxbc1/sacerdote9.png?raw=1)
]

---

# Conclusions

.pull-left[

- Peer effects are very strong!

- Important influences in Freshman year performance (GPA) and activities (joining a social organization)

- Not important for choosing a major

]

.pull-right[
.center[
![](https://www.dropbox.com/s/bdy9gn83dgbc1s3/universitypeers.jpg?raw=1)
]
]
.source[Sacerdote, Bruce, (2001), "Peer Effects with Random Assignment: Results from Dartmouth Roommates" *Quarterly Journal of Economics* 116(2):681-704]

---

.class: inverse, center, middle
# Duggan and Levitt (2002) on Corruption in Sumo Wrestling

---

# Paper Motivation I

.pull-left[

- 2000 year history, national sport of Japan, extremely ritualistic

- Japan is a country with low corruption (CPI: 75, Rank 18th best)

- Good data available

- Situation is ripe for cheating! So when/why does it happen?

]

.pull-right[
.center[
![](https://www.dropbox.com/s/iu3ua5wbh2ulv0d/sumo.jpg?raw=1)
]
]
.source[Duggan, Mark and Steven D. Levitt, (2002), "Winning isn't Everything: Corruption in Sumo Wrestling" *American Economic Review* 92(5):1594-1605]

---

# Relevant Institutions I

.pull-left[

- Tournaments (*basho*), 66 wrestlers (*rikishi*), 15 bouts each

- Wrestlers with 8+ wins (*kachi-koshi*) *move up* in rankings (*banzuke*)

- Those with a losing record ($<$8 wins) (*maki-koshi*) *fall* in rankings

]

.pull-right[
.center[
![](https://www.dropbox.com/s/iu3ua5wbh2ulv0d/sumo.jpg?raw=1)
]
]
.source[Duggan, Mark and Steven D. Levitt, (2002), "Winning isn't Everything: Corruption in Sumo Wrestling" *American Economic Review* 92(5):1594-1605]

---

# Relevant Institutions II

.pull-left[

- A marginal win generates a 2.5 rank increase

- But movement from 7 to 8 wins produces almost an 11 rank increase!

- Rank signals prestige, moving up a single rank is worth about $3,000/year

- Top 5th-10th ranked wrestlers make $250,000/year

]

.pull-right[
.center[
![](https://www.dropbox.com/s/23hzngfz77h7rtb/sumo1.png?raw=1)
]
]
.source[Duggan, Mark and Steven D. Levitt, (2002), "Winning isn't Everything: Corruption in Sumo Wrestling" *American Economic Review* 92(5):1594-1605]

---

# Relevant Institutions III

.pull-left[

- Consider 2 wrestlers: .red[A (8-6)] vs .blue[B (7-7)] going into final (15th) match

- Return to winning for .blue[B (7-7)] is much higher than for .red[A (8-6)]

- .red[A (8-6)] throws the match to .blue[B (7-7)], who must return the favor in later tournaments if .red[A] finds himself in the same 7-7 position

]

.pull-right[
.center[
![](https://www.dropbox.com/s/iu3ua5wbh2ulv0d/sumo.jpg?raw=1)
]
]
.source[Duggan, Mark and Steven D. Levitt, (2002), "Winning isn't Everything: Corruption in Sumo Wrestling" *American Economic Review* 92(5):1594-1605]

---

# The Data

- All official top-rank sumo matches from January 1989-January 2000

- Six tournaments per year, nearly 70 wrestlers per tournament

- Tournaments last 15 days with one match per wrestler

- 64,000 wrestler-matches

---

# The Theoretical vs. Actual Probability of Winning

.pull-left[
- Theoretical (binomial) probability of winning 8 times: 19.6%

- Actual probability (from data): 26%

- Much higher probability for 8 wins than it should be! (& lower for 7)
  - Could this be from rampant cheating!?

]

.pull-right[
.center[
![](https://www.dropbox.com/s/c8tf3yx6thuil89/sumo2.png?raw=1)
]
]

---

# The Model (Slightly Modified)

$$\text{Win}_{ijtd}=\beta_1 Bubble_{ijtd}+ \beta_2 Rankdiff_{ijt}+\lambda_{ij}+\delta_{it}+u_{ijtd}$$

- $Win=1$ if wrestler $i$ beats wrestler $j$ in tournament $t$ on day $d$

- $Bubble=1$ if wrestler ($i$) is on margin (7-7), -1 if opponent ($j$) is on margin, $=0$ if neither are on margin

- $Rankdiff$: difference in rank between wrestlers

- Wrestler $\lambda$ and tournament $\delta$ fixed effects

---

# Initial Results

.center[
![](https://www.dropbox.com/s/bamwd6eq97rugou/sumo4.png?raw=1)

- Frequency of rigging increases as tournament nears end (day 15)

- On day 15, 7-7 wrestlers on margin win 25% more often than they otherwise should
]

---

# Interpretation of Initial Results

- Two alternative hypotheses to explain results:
  1. Match rigging (corruption)
  2. Effort: wrestlers on margin (7-7) just fight harder! Wrestlers with 8 wins are more complacent (already made it)

- To test, look for evidence of **reciprocity** agreements over time 
  - If these tacit agreements to rig matches exist, wrestlers from stable A should have very high win rates when on the margin against wrestlers from stable B, and vice versa

---

# Interaction Effects

.pull-left[
.center[
![](https://www.dropbox.com/s/qnh3i0pwogiz1y6/sumo5.png?raw=1)
]
]

.pull-right[
- Last row (before $R^2$): wrestler's success strongly increases with overall success rates of playing wrestlers on the bubble from other stables

- For each 10% increase in success in other bubble matches between these two stables, the wrestler on the bubble is 2.7% more likely to win
]

---

# Reciprocity?

.center[
![](https://www.dropbox.com/s/z2l381t0yqaykki/sumo6.png?raw=1)
]

- Wrestlers who won a bubble match previously tend to do worse when playing same opponent - throw the match to them in reciprocity!

---

class: inverse, center, middle
# Fisman and Miguel (2007)

---

# Paper Motivation

.pull-left[

- What determines the level of corruption?
    1. Legal environment of country
    2. Culture or social norms

- How to identify the true source(s)?

]

.pull-right[
.center[
![](https://www.dropbox.com/s/92fk22dz2zctozu/rentseeking.png?raw=1)
]
]
.source[Fisman, Raymond and Edward Miguel, (2007), "Corruption, Norms, and Legal Enforcement: Evidence from Diplomatic Parking Tickets," *Journal of Political Economy* 115(6): 1020-1048]

---

# Relevant Institutions

.pull-left[

- U.N. Diplomats are given immunity from prosecution or lawsuits in the U.S.

- Reciprocal agreements with other countries, designed to protect diplomats in unfriendly environments

- Diplomatic license plates in NYC are identified, get ticket, but no way to enforce

- "The best free parking pass in town"

]

.pull-right[
.center[
![](https://www.dropbox.com/s/qk19adm49gy5plj/parkingticket.jpg?raw=1)

![](https://www.dropbox.com/s/qtgg9hkmpecm4dw/diplomatplates.jpg?raw=1)
]
]

.source[Fisman, Raymond and Edward Miguel, (2007), "Corruption, Norms, and Legal Enforcement: Evidence from Diplomatic Parking Tickets," *Journal of Political Economy* 115(6): 1020-1048]

---

# The Data

- Between 11/1987 and 12/2002, 150,000 unpaid parking tickets, fines totaling $18,000,000

- 30 Days to pay a fine, afterwards a 110% penalty. After 70 days, recorded as unpaid violation

- Individual violation-level data: license plate, name, country of origin, date & time of violation, fine, amount paid (if any)

- 43% were violations of "no standing/loading zone"

- 20% of cases, the car was registered to the diplomatic mission (not personal)

- Scale fines by the size of the country's mission

---

# Identification Strategy I

- Becker's (1968) rational crime model says with no punishment$\implies$ rational for all diplomats to never pay parking fines

- But large variation in data! Unpaid fines are strongly correlated with country's score on corruption index!

- Home country corruption norms are an important predictor of diplomats breaking the law
    - Low corruption countries' diplomats tend to pay the fine even where they are not legally compelled to
    - High corruption countries' diplomats rack up massive fines

---

# Identification Strategy II

.pull-left[
- **Natural experiment**: post-9/11, NYC began cracking down on enforcement

- Diplomats with 3+ unpaid parking tickets had diplomat plates revoked

- Led to immediate 98% reduction in unpaid parking tickets

- So enforcement matters as well as corruption norms
]

.pull-right[
.center[
![](https://www.dropbox.com/s/xe59fx3ced0cva9/parking1.png?raw=1)
]
]
---

# Data I

.center[
![](https://www.dropbox.com/s/5ndr02pwjvv6zr3/parking2.png?raw=1)
]

---

# Data II

.center[
![](https://www.dropbox.com/s/izkvkpu2izi78sr/parking3.png?raw=1)
]

---

# Empirical Model

$$\text{Unpaid Violations}=\beta_0+\beta_1\text{Corruption}+\beta_2\text{Enforcement}+\beta_3\text{Diplomats}+...+\beta_k\text{Controls}$$

---

# Results I

.center[
![](https://www.dropbox.com/s/rf6r3g66u3aaulr/parking4.png?raw=1)
]

---
# Results II

.center[
![](https://www.dropbox.com/s/2drdlc105j2i15m/parking7.png?raw=1)
]

---
# Results III

.center[
![](https://www.dropbox.com/s/t1s9e3xb3kkn428/parking5.png?raw=1)
]