### 2019-11-05 content/class/17-class.Rmd
Add class 17 content

---
title: "3.5: Causal Inference and DAGs - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, November 5, 2019*

# Overview

Today we return to a more nuanced discussion of causality, given what we now know about multivariate regression. We also introduce a new tool for thinking about simple causal models, the **directed acyclic graph (DAG)*. These are a hip new trend for thinking about causal inference, so new and trendy that they aren't really in any mainstream textbooks yet!

# New Packages Mentioned

- `ggdag`: for drawing DAGs in `ggplot`
- [`dagitty.net`](http://dagitty.net): not a package in R, but an online tool to help you work with DAGs (see the cheat sheets in today's [readings page](/reading/17-reading))

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/17-slides.html)