### 2019-11-12 content/class/18-class.Rmd
Add class 18 content

---
title: "3.6: Regression with Categorical Data - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, November 12, 2019*

# Overview

Today we look at data that is categorical.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/18-slides.html)

# Problem Set 4 Due Tuesday Nov 19

[Problm Set 4](/assignment/04-problem-set) (on classes 3.1-3.5) is due by Tuesday November 19.