### 2019-11-12 static/slides/18-slides.Rmd

---
title: "3.6: Regression with Categorical Data"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))

library(wooldridge)
wages<-wooldridge::wage1
wages<-wages %>%
  mutate(Gender = factor(ifelse(female==0,
                         "Male",
                         "Female")))
```

# Categorical Data

.pull-left[

- .hl[Categorical data] place an individual into one of several possible *categories*
    - e.g. sex, season, political party
    - may be responses to survey questions
    - can be quantitative (e.g. age, zip code)

- `R` calls these `factors` (we'll deal with them much later in the course)
]

.pull-right[

![](https://www.dropbox.com/s/pgdswagr7b348vu/categoricaldata.png?raw=1)

]

---

# Example Research Question

.pull-left[

.content-box-green[
.green[**Example**]: do men earn higher wages, on average, than women? If so, how much?
]

]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/481uod9g3n1w7kr/genderpaygap.jpg?raw=1)
]
]

---

# The Pure Statistics of Comparing Group Means

.pull-left[

- Using basic statistics, we can test for a statistically significant difference in group means with a **t-test**<sup>.red[1]</sup>

- Define:

- .blue[`\\(Y_M\\)`]: average earnings of a sample of .blue[`\\(n_M\\)`] men

- .magenta[`\\(Y_W\\)`]: average earnings of a sample of .magenta[`\\(n_W\\)`] women

- **Difference** in group averages: $d=$ .blue[`\\(\bar{Y}_M\\)`] $-$ .magenta[`\\(\bar{Y}_W\\)`]

- The hypothesis test is:
  - $H_0: d=0$
  - $H_1: d \neq 0$

]

.pull-right[
.center[
![:scale 100%](https://www.dropbox.com/s/481uod9g3n1w7kr/genderpaygap.jpg?raw=1)
]
]

.footnote[ <sup>.red[1]</sup> See today's class notes page for this example]

---

# Comparing Groups with Regression

- In a regression, we can easily compare across groups via a .shout[dummy variable]<sup>.red[1]</sup>

- Dummy variable *only* $=0$ or $=1$, depending on if a condition is met

- Signifies whether an observation belongs to a category or not

.footnote[<sup>.red[1]</sup> Also called a .shout[binary variable] or .shout[dichotomous variable]]

--

.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i \quad \quad \text{ where } Female_i =
 		 \begin{cases}
    		1 & \text{if person } i \text{ is } Female \\
   			0 & \text{if person } i \text{ is } Male\\
  		\end{cases}$$

]

--

- Again, $\hat{\beta_1}$ makes less sense as the "slope" of a line in this context

---

# Comparing Groups in Regression: Scatterplot

.pull-left[

- `Female` is our dummy $x$-variable

- Hard to see relationships because of **overplotting**

]

.pull-right[
```{r, fig.retina=3}
colors<-c(Female = "pink", Male = "blue")

ggplot(data = wages)+
  aes(x = female,
      y = wage)+
  geom_point(aes(color = Gender))+
  geom_smooth(method = "lm", color = "black")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Female",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Comparing Groups in Regression: Scatterplot

.pull-left[

- `Female` is our dummy $x$-variable

- Hard to see relationships because of **overplotting**

- Use `geom_jitter()` instead of `geom_point()` to *randomly* nudge points
  - *Only* for plotting purposes, does not affect actual data, regression, etc.

]

.pull-right[
```{r, fig.retina=3}
colors<-c(Female = "pink", Male = "blue")

ggplot(data = wages)+
  aes(x = female,
      y = wage)+
  geom_jitter(aes(color = Gender),
              width=0.05)+
  geom_smooth(method = "lm", color = "black")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Female",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Comparing Groups in Regression: Scatterplot

.pull-left[

- `Female` is our dummy $x$-variable

- Hard to see relationships because of **overplotting**

- Use `geom_jitter()` instead of `geom_point()` to *randomly* nudge points
  - *Only* for plotting purposes, does not affect actual data, regression, etc.

]

.pull-right[
```{r, fig.retina=3}
male_avg<-wages %>%
  filter(female==0) %>%
  summarize(mean(wage)) %>%
  pull() # get just a number (not a dataframe)

female_avg<-wages %>%
  filter(female==1) %>%
  summarize(mean(wage)) %>%
  pull() # get just a number (not a dataframe)

ggplot(data = wages)+
  aes(x = female,
      y = wage)+
  geom_jitter(aes(color = Gender),
              width=0.05)+
  geom_smooth(method = "lm", color = "black")+
  geom_hline(yintercept=male_avg,linetype="dashed",color="blue")+
  geom_label(x = 0.5, y=male_avg, label="Average for Men", color="blue")+
  geom_hline(yintercept=female_avg,linetype="dashed",color="red")+
  geom_label(x = 0.5, y=female_avg, label="Average for Women", color="red")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Female",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Dummy Variables as Group Means

$$\hat{Y_i}=\hat{\beta_0}+\hat{\beta_1} D_i \quad \text{ where }D_i=\{0,1\}$$

--

- .red[When `\\(D_i=0\\)` (Control group):]
    - $\hat{Y_i}=\hat{\beta_0}$
    - $E[Y|D_i=0]=\hat{\beta_0}$ $\iff$ the mean of $Y$ when $D_i=0$
--

- .blue[When `\\(D_i=1\\)` (Treatment group):]
    - $\hat{Y_i}=\hat{\beta_0}+\hat{\beta_1} D_i$
    - $E[Y|D_i=1]=\hat{\beta_0}+\hat{\beta_1}$ $\iff$ the mean of $Y$ when $D_i=1$

--

- So the **difference** in group means:

$$\begin{align*}
		&=E[Y_i|D_i=1]-E[Y_i|D_i=0]\\
		&=(\hat{\beta_0}+\hat{\beta_1})-(\hat{\beta_0})\\
		&=\hat{\beta_1}\\
\end{align*}$$

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men: 
]

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men:
$$E[Wage|Female=0]=\hat{\beta_0}$$
]

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men: 
$$E[Wage|Female=0]=\hat{\beta_0}$$

- Mean wage for women: 
]

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men:
$$E[Wage|Female=0]=\hat{\beta_0}$$

- Mean wage for women:
$$E[Wage|Female=1]=\hat{\beta_0}+\hat{\beta_1}$$
]

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men:
$$E[Wage|Female=0]=\hat{\beta_0}$$

- Mean wage for women:
$$E[Wage|Female=1]=\hat{\beta_0}+\hat{\beta_1}$$

- Difference in wage between men & women:
]

---

# Dummy Variables as Group Means: Our Example

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Female_i$$

$$\text{where } Female_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is }Female \\
   			0 & \text{if } i \text{ is } Male\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men: 
$$E[Wage|Female=0]=\hat{\beta_0}$$

- Mean wage for women: 
$$E[Wage|Female=1]=\hat{\beta_0}+\hat{\beta_1}$$

- Difference in wage between men & women: 
$$d = \hat{\beta_1}$$
]

---

# Comparing Groups in Regression: Scatterplot

.pull-left[

- `Female` is our dummy $x$-variable

- Hard to see relationships because of **overplotting**

- Use `geom_jitter()` instead of `geom_point()` to *randomly* nudge points
  - *Only* for plotting purposes, does not affect actual data, regression, etc.

]

.pull-right[
```{r, fig.retina=3}
ggplot(data = wages)+
  aes(x = female,
      y = wage)+
  geom_jitter(aes(color = Gender),
              width=0.05)+
  geom_smooth(method = "lm", color="black")+
  geom_hline(yintercept=male_avg,linetype="dashed",color="blue")+
  geom_label(x = 0.5, y=male_avg, label=expression(hat(beta[0])), color="blue")+
  geom_hline(yintercept=female_avg,linetype="dashed",color="red")+
  geom_label(x = 0.5, y=female_avg, label=expression(hat(beta[0])+hat(beta[1])), color="red")+
  geom_label(x = 0.75, y=5.5, label=expression(paste(hat(beta[1]),"= difference")), color="black")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Female",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# The Data

```{r, echo=T}
# from wooldridge package
library(wooldridge)

# save as a dataframe
wages<-wooldridge::wage1

head(wages)
```

---

# Get Group Averages & Std. Devs.

.pull-left[

```{r, echo=T}
# Summarize for Men

wages %>%
  filter(female==0) %>%
  summarize(mean = mean(wage),
            sd = sd(wage))
```

]

.pull-right[

```{r, echo=T}
# Summarize for Women

wages %>%
  filter(female==1) %>%
  summarize(mean = mean(wage),
            sd = sd(wage))
```

]

---

# Visualize Differences

```{r, fig.retina=3}
fill_colors<-c("0" = "blue", "1" = "red")

ggplot(data = wages)+
  aes(x = wage,
      fill = factor(female))+
  geom_density(alpha=0.5)+
  scale_x_continuous(labels=scales::dollar)+
  labs(x = "wage",
       y = "Density")+
  scale_fill_manual("Gender", labels=c("Men","Women"), values = fill_colors)+
    theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```

---

# The Regression I

```{r, echo=T}
femalereg<-lm(wage~female, data=wages)
summary(femalereg)
```

---

# The Regression II

```{r, echo=T}
library(broom)
tidy(femalereg)
```

---

# Dummy Regression vs. Group Means

.pull-left[

From tabulation of group means

| Gender | Avg. Wage | Std. Dev.   | $n$   |
|--------|-------------|-----------|-------|
| Female | $4.59$    | $2.33$      | $252$ |
| Male   | $7.10$    | $4.16$      | $274$ |
| Difference | $2.51$ | $0.30$ | $-$ |

From $t$-test of difference in group means

]

.pull-right[
```{r}
tidy(femalereg)
```

]

---

class: inverse, center, middle

# Recoding Dummies

---

# Recoding Dummies

.content-box-green[
.green[**Example**]:

- Suppose instead of $female$ we had used:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Male_i \quad \quad \text{ where } Male_i =
 		 \begin{cases}
    		1 & \text{if person } i \text{ is } Male \\
   			0 & \text{if person } i \text{ is } Female\\
  		\end{cases}$$

]

---

# Recoding Dummies with Data

```{r}
wages<-wages %>%
  mutate(male = ifelse(female == 0, # condition: is female equal to 0?
                       1, # if true: code as "1"
                       0)) # if false: code as "0"

# verify it worked
wages %>% 
  select(wage, female, male) %>%
  head()
```

---

# Scatterplot with Male

.pull-left[

- `Female` is our dummy $x$-variable

- Hard to see relationships because of **overplotting**

- Use `geom_jitter()` instead of `geom_point()` to *randomly* nudge points
  - *Only* for plotting purposes, does not affect actual data, regression, etc.

]

.pull-right[
```{r, fig.retina=3}
wages<-wages %>%
  mutate(Gender = factor(ifelse(female==0,
                         "Male",
                         "Female")))

ggplot(data = wages)+
  aes(x = male,
      y = wage)+
  geom_jitter(aes(color = Gender),
              width=0.05)+
  geom_smooth(method = "lm", color = "black")+
  geom_hline(yintercept=male_avg,linetype="dashed",color="blue")+
  geom_label(x = 0.5, y=male_avg, label="Average for Men", color="blue")+
  geom_hline(yintercept=female_avg,linetype="dashed",color="red")+
  geom_label(x = 0.5, y=female_avg, label="Average for Women", color="red")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Male",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# Dummy Variables as Group Means: With Male

.pull-left[
.content-box-green[
.green[**Example**]:

$$\widehat{Wage_i}=\hat{\beta_0}+\hat{\beta_1}Male_i$$

$$\text{where } Male_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is } Male \\
   			0 & \text{if } i \text{ is } Female\\
  		\end{cases}$$
]
]

.pull-right[

- Mean wage for men: 
$$E[Wage|Male=1]=\hat{\beta_0}+\hat{\beta_1}$$

- Mean wage for women: 
$$E[Wage|Male=0]=\hat{\beta_0}$$

- Difference in wage between men & women: 
$$d = \hat{\beta_1}$$
]

---

# Scatterplot with Male

.pull-left[

]

.pull-right[
```{r, fig.retina=3}
ggplot(data = wages)+
  aes(x = male,
      y = wage)+
  geom_jitter(aes(color = Gender),
              width=0.05)+
  geom_smooth(method = "lm", color="black")+
  geom_hline(yintercept=male_avg,linetype="dashed",color="blue")+
  geom_label(x = 0.5, y=male_avg, label=expression(hat(beta[0])+hat(beta[1])), color="blue")+
  geom_hline(yintercept=female_avg,linetype="dashed",color="red")+
  geom_label(x = 0.5, y=female_avg, label=expression(hat(beta[0])), color="red")+
  geom_label(x = 0.75, y=5.5, label=expression(paste(hat(beta[1]),"= difference")), color="black")+
  scale_x_continuous(breaks=c(0,1),
                   labels=c(0,1))+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Male",
       y = "Wage")+
  guides(color=F)+ # hide legend
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# The Regression with Male I

```{r, echo=T}
malereg<-lm(wage~male, data=wages)
summary(malereg)
```

---

# The Regression with Male II

```{r, echo=T}
library(broom)
tidy(malereg)
```

---

# The Dummy Regression: Male or Female

.pull-left[

```{r}
library(huxtable)
huxreg(malereg, femalereg,
       coefs = c("Constant" = "(Intercept)",
                 "Female" = "female",
                 "Male" = "male"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

]

.pull-right[

- Note it doesn't matter if we use `male` or `female`, males always earn $2.51 more than females

- Compare the constant (average for the $D=0$ group)

- Should you use `male` AND `female`? We'll come to that...

]

---

class: inverse, center, middle

# Categorical Variables (More than 2 Categories)

---

# Categorical Variables with More than 2 Categories

- A .shout[categorical variable] expresses membership in a category, where there is no ranking or hierarchy of the categories
    - We've looked at categorical variables with 2 categories only
    - e.g. Male/Female, Spring/Summer/Fall/Winter, Democratic/Republican/Independent

--

- Might be an .shout[ordinal variable] expresses rank or an ordering of data, but not necessarily their relative magnitude
    - e.g. Order of finalists in a competition (1st, 2nd, 3rd)
    - e.g. Highest education attained (1=elementary school, 2=high school, 3=bachelor's degree, 4=graduate degree)

---

# Using Categorical Variables in Regression I

.content-box-green[
.green[**Example**]: How do wages vary by region of the country? Let $Region_i=\{Northeast, \, Midwest, \, South, \, West\}$
]

--

- Can we run the following regression?

$$\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Region_i$$

---

# Using Categorical Variables in Regression II

.content-box-green[
.green[**Example**]: How do wages vary by region of the country?

Code region numerically: $$Region_i=
\begin{cases}1 & \text{if } i \text{ is in }Northeast \\
   			2 & \text{if } i \text{ is in } Midwest\\
    		3 & \text{if } i \text{ is in } South \\
   			4 & \text{if } i \text{ is in } West\\ \end{cases}$$
]

--

- Can we run the following regression?

$$\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Region_i$$

---

# Using Categorical Variables in Regression III

.content-box-green[
.green[**Example**]: How do wages vary by region of the country?

Create a dummy variable for *each* region:

- $Northeast_i = 1$ if $i$ is in Northeast, otherwise $=0$
- $Midwest_i = 1$ if $i$ is in Midwest, otherwise $=0$
- $South_i = 1$ if $i$ is in South, otherwise $=0$
- $West_i = 1$ if $i$ is in West, otherwise $=0$

]

--

- Can we run the following regression?

$$\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i+\hat{\beta_4}West_i$$

--

- For every $i: \, Northeast_i+Midwest_i+South_i+West_i=1$!

---

# The Dummy Variable Trap

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i+\hat{\beta_4}West_i$

]

- If we include *all* possible categories, they are .onfire[perfectly multicollinear], an exact linear function of one another:

$$Northeast_i+Midwest_i+South_i+West_i=1 \quad \forall i$$

- This is known as the .shout[dummy variable trap], a common source of perfect multicollinearity

---

# The Reference Category

- To avoid the dummy variable trap, always omit one category from the regression, known as the .onfire[reference category]

- It does not matter which category we omit!

- .onfire[Coefficients on each dummy variable measure the *difference* between the *reference* category and each category dummy]

---

# The Reference Category: Example

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i$
]

- $West_i$ is omitted (arbitrarily chosen)

--

- $\hat{\beta_0}$:

---

# The Reference Category: Example

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i$
]

- $West_i$ is omitted (arbitrarily chosen)

- $\hat{\beta_0}$: average wage for $i$ in the West

--

- $\hat{\beta_1}$:

---

# The Reference Category: Example

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i$
]

- $West_i$ is omitted (arbitrarily chosen)

- $\hat{\beta_0}$: average wage for $i$ in the West

- $\hat{\beta_1}$: difference between West and Northeast

--

- $\hat{\beta_2}$:

---

# The Reference Category: Example

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i$
]

- $West_i$ is omitted (arbitrarily chosen)

- $\hat{\beta_0}$: average wage for $i$ in the West

- $\hat{\beta_1}$: difference between West and Northeast

- $\hat{\beta_2}$: difference between West and Midwest

--

- $\hat{\beta_3}$:

---

# The Reference Category: Example

.content-box-green[
.green[**Example**]: $\widehat{Wages_i}=\hat{\beta_0}+\hat{\beta_1}Northeast_i+\hat{\beta_2}Midwest_i+\hat{\beta_3}South_i$
]

- $West_i$ is omitted (arbitrarily chosen)

- $\hat{\beta_0}$: average wage for $i$ in the West

- $\hat{\beta_1}$: difference between West and Northeast

- $\hat{\beta_2}$: difference between West and Midwest

- $\hat{\beta_3}$: difference between West and South

---

# Dummy Variable Trap in R

```{r}
# create northeast dummy as the final category
wages$noreast<-ifelse((wages$northcen==0 & wages$south==0 & wages$west==0),1,0)
```

.font90[
```{r, echo=T}
summary(lm(wage~noreast+northcen+south+west, data=wages))
```
]

---

# Using Different Reference Categories in R

```{r}
# create northeast dummy as the final category
wages$noreast<-ifelse((wages$northcen==0 & wages$south==0 & wages$west==0),1,0)
```

```{r dummyregs, echo=T, eval=F}
# let's run 4 regressions, each one we omit a different region
no_noreast_reg<-lm(wage~northcen+south+west, data = wages)
no_northcen_reg<-lm(wage~noreast+south+west, data = wages)
no_south_reg<-lm(wage~noreast+northcen+west, data = wages)
no_west_reg<-lm(wage~noreast+northcen+south, data = wages)

# now make an output table
library(huxtable)
huxreg(no_noreast_reg,
       no_northcen_reg,
       no_south_reg,
       no_west_reg,
       coefs = c("Constant" = "(Intercept)",
                 "Northeast" = "noreast",
                 "Midwest" = "northcen",
                 "South" = "south",
                 "West" = "west"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 3)
```

---

# Using Different Reference Categories in R II

.pull-left[
.font90[
```{r, ref.label="dummyregs"}

```
]
]
.pull-right[

- Constant is alsways average wage for reference (omitted) region

- Compare coefficients between Midwest in (1) and Northeast in (2)...

- Compare coefficients between West in (3) and South in (4)...

- Does not matter which region we omit!
  - Same $R^2$, SER, coefficients give same results

]

---

# Dummy *Dependent* (Y) Variables

- In many contexts, we will want to have our *dependent* ($Y$) variable be a dummy variable

--

.content-box-green[
.green[**Example**]: 
$$\widehat{Admitted_i}=\hat{\beta_0}+\hat{\beta_1}GPA_i \quad \text{ where } Admitted_i =
 		 \begin{cases}
    		1 & \text{if } i \text{ is Admitted} \\
   			0 & \text{if } i \text{ is  Not Admitted}\\
  		\end{cases}$$
]

--

- A model where $Y$ is a dummy is called a .shout[linear probability model], as it measures the .onfire[probability of `\\(Y\\)` occuring `\\((=1)\\)` given the X's, i.e. `\\(P(Y_i=1|X_1, \cdots, X_k)\\)`]
    - e.g. the probability person $i$ is Admitted to a program with a given GPA

--

- Requires special tools to properly interpret and extend this (**logit**, **probit**, etc)

- Feel free to write papers that have dummy $Y$ variables (but you may have to ask me some more questions)!