### 2019-11-12 content/class/18-class.Rmd
Add r practice 17 answers

---
title: "3.6: Regression with Categorical Data - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, November 12, 2019*

# Overview

Today we look at how to use data that is categorical (i.e. variables that indicate an observation's membership in a particular group or category). We introduce them into regression models as **dummy variables** that can equal 0 or 1: where 1 indicates membership in a category, and 0 indicates non-membership.

We also look at what happens when categorical variables have more than two values: for regression, we introduce a dummy variable for each possible category - but be sure to leave out one reference category to avoid the dummy variable trap.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/18-slides.html)

# Problem Set 4 Due Tuesday Nov 19

[Problm Set 4](/assignment/04-problem-set) (on classes 3.1-3.5) is due by Tuesday November 19.