### 2019-11-19 content/class/19-class.Rmd

---
title: "3.7: Regression with Interaction Effects - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Thursday, November 14, 2019*

# Overview

Another way to visualize what we've been covering:

```{r, fig.retina=3}
library(mosaic)

control=function(x){2+0.5*x}
treat=function(x){4+x}

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=control, geom="line", size=2, color = "red")+
    geom_label(aes(x=5,y=control(5)), color = "red", label=expression(paste("slope=",hat(beta[1]))), size = 5)+
  stat_function(fun=treat, geom="line", size=2, color = "blue")+
    geom_label(aes(x=5,y=treat(5)), color = "blue", label=expression(paste("slope=",hat(beta[1])+hat(beta[3]))), size = 5)+
    scale_x_continuous(breaks=NULL,
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=c(2,4),
                     labels=c(expression(hat(beta[0])),expression(hat(beta[0])+hat(beta[2]))),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  labs(x = "",
       y = "")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)

```

$$\widehat{\text{wage}}_i=\hat{\beta_0}+\hat{\beta_1} \, \text{experience}_i$$

$$\widehat{\text{wage}}_i=\hat{\beta_0}+\hat{\beta_1} \, \text{experience}_i+\hat{\beta_2} \, \text{gender}_i$$

$$\widehat{\text{wage}}_i=\hat{\beta_0}+\hat{\beta_1} \, \text{experience}_i+\hat{\beta_2} \, \text{gender}_i+ \hat{\beta_3}(\text{experience}_i \times \text{gender}_i)$$

```{r}
library(tidyverse)
library(wooldridge)
# Our data comes from wage1 in the wooldridge package

wages<-wooldridge::wage1
```

```{r, fig.retina=3}
ggplot(data = wages)+
  aes(x = exper,
      y = wage)+
  geom_point()+
  geom_smooth(method="lm")+
  scale_y_continuous(labels=scales::dollar)+
  labs(x = "Experience (years)",
       y = "Wage")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```

```{r}
library(broom)
gender_reg<-lm(wage~exper+female, data = wages) %>%
  tidy()

gender_reg

slope<-gender_reg %>%
  filter(term=="exper") %>%
  pull(estimate)
slope

male_intercept<-gender_reg %>%
  filter(term=="(Intercept)")%>%
  pull(estimate)

female_intercept<-gender_reg %>%
  filter(term=="female") %>%
  pull(estimate)

female_intercept<-female_intercept+male_intercept
```

```{r, fig.retina=3}
colors = c("1" = "pink", "0" = "blue")

ggplot(data = wages)+
  aes(x = exper,
      y = wage,
      color = as.factor(female))+
  geom_point()+
  geom_abline(slope=slope, intercept = male_intercept, size=2, color = "blue")+
  geom_abline(slope=slope, intercept = female_intercept, size=2, color = "pink")+
  geom_label(x=2.5,y=6.63, color="blue", label="6.63")+
  geom_label(x=2.5,y=4.15, color="pink", label="4.15")+
  scale_y_continuous(labels=scales::dollar)+
  scale_color_manual("Gender", labels=c("Men","Women"), values = colors)+
  guides(color=F)+
  labs(x = "Experience (years)",
       y = "Wage",
       title = "Parallel Slopes Model",
       subtitle = "wage = 6.63+0.03 Experience-2.48 Female")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```
]

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/19-slides.html)

# Problem Set 4 Due Thursday Nov 19

[Problm Set 4](/assignment/04-problem-set) (on classes 3.1-3.5) is due by Thursday November 21