### 2019-11-19 static/slides/20-slides.Rmd

---
title: "3.8: Polynomial Regression"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

]

--

.pull-right[

```{r, fig.retina=3}
library(gapminder)
p1<-ggplot(data = gapminder)+
  aes(x = gdpPercap,
      y = lifeExp)+
  geom_point(color="blue", alpha=0.5)+
  scale_x_continuous(labels=scales::dollar,
                     breaks=seq(0,120000,20000))+
  scale_y_continuous(breaks=seq(0,90,10),
                     limits=c(0,90))+
  labs(x = "GDP per Capita",
       y = "Life Expectancy (Years)")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
p1
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

]

.pull-right[

```{r, fig.retina=3}
p1+geom_smooth(method="lm", color="red")
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

]

.pull-right[

```{r, fig.retina=3}
p1+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color = "green")
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

- Get rid of the outliers $(>60,000)$

]

.pull-right[

```{r, fig.retina=3}
gap_60<-gapminder %>%
  filter(gdpPercap<60000)

p2<-ggplot(data = gap_60)+
  aes(x = gdpPercap,
      y = lifeExp)+
  geom_point(color="blue", alpha=0.5)+
  scale_x_continuous(labels=scales::dollar,
                     breaks=seq(0,60000,20000))+
  scale_y_continuous(breaks=seq(0,90,10),
                     limits=c(0,90))+
  labs(x = "GDP per Capita",
       y = "Life Expectancy (Years)")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
p2
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

- Get rid of the outliers $(>60,000)$

]

.pull-right[

```{r, fig.retina=3}
p2+geom_smooth(method="lm", color="red")
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

- Get rid of the outliers $(>60,000)$

]

.pull-right[

```{r, fig.retina=3}
p2+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color = "green")
```
]

---

# *Linear* Regression

.pull-left[

- OLS is commonly known as "**_linear_ regression**" as it fits a *straight line* to data points

- Often, data and relationships between variables may *not* be linear

- Get rid of the outliers $(>60,000)$

]

.pull-right[

```{r, fig.retina=3}
p2+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color = "green")+
  stat_smooth(method = "lm", formula=y~log(x), color="orange", size=2)
```
]

---

# Nonlinear Effects in Linear Regression

- Despite being "linear regression", OLS can handle this with an easy fix

- OLS requires all *parameters* (i.e. the $\beta$'s) to be linear, the *regressors* $(X$'s) can be nonlinear:

--
$$Y_i=\beta_0+\beta_1 X_i^2 \quad  \mathbf{\checkmark}$$
--

$$Y_i=\beta_0+\beta_1^2X_i \quad  \mathbf{X}$$

--

$$Y_i=\beta_0+\beta_1 \sqrt{X_i} \quad  \mathbf{\checkmark}$$

--

$$Y_i=\beta_0+\sqrt{\beta_1} X_i \quad  \mathbf{X} $$

--

$$Y_i=\beta_0+\beta_1 (X_{1i} \times X_{2i}) \quad  \mathbf{\checkmark}$$

--

$$Y_i=\beta_0+\beta_1 ln(X_i) \quad  \mathbf{\checkmark}$$
--

- In the end, each $X$ is always just a number in the data, OLS can always estimate parameters for it

- *Plotting* the modelled points $(X_i, \hat{Y_i})$ can result in a curve!

---

# Sources of Nonlinearities I

- Effect of $X_1 \rightarrow Y$ might be nonlinear if:

--

1. $X_1 \rightarrow Y$ is different for different levels of $X_1$
  - e.g. diminishing returns: $\uparrow X_1$ increases $Y$ at a *decreasing* rate
  - e.g. increasing returns: $\uparrow X_1$ increases $Y$ at an *increasing* rate

--

2. $X_1 \rightarrow Y$ is different for different levels of $X_2$
  - e.g. interaction effects (last lesson)

---

# Sources of Nonlinearities II

.pull-left[

- **Linear**: 
  - slope $(\hat{\beta_1})$, $\frac{\Delta X}{\Delta Y}$ same for all $X$

]
.pull-right[

```{r fig.retina=3}

line=function(x){10-x}

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=line, geom="line", size=2, color="blue")+
  geom_segment(aes(x=5, xend=5, y=line(5), yend=line(6)), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=5, xend=6, y=line(6), yend=line(6)), linetype="dotted", color="black", size=1)+
  scale_x_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```
]

---

# Sources of Nonlinearities II

.pull-left[

- **Polynomial**:
  - slope $(\hat{\beta_1})$, $\frac{\Delta X}{\Delta Y}$ *depends on the value of* $X$

]
.pull-right[

```{r fig.retina=3}

curve=function(x){10/x}

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=curve, geom="line", size=2, color="blue")+
  geom_segment(aes(x=1, xend=1, y=5, yend=9), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=1, xend=2, y=5, yend=5), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=5, xend=5, y=2, yend=1), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=5, xend=9, y=1, yend=1), linetype="dotted", color="black", size=1)+
  scale_x_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```
]

---

# Sources of Nonlinearities III

.pull-left[

- **Interaction Effect**:
  - slope $(\hat{\beta_1})$, $\frac{\Delta X_1}{\Delta Y}$ *depends on the value of* $X_2$

- Easy example: if $X_2$ is a dummy variable:
  - .red[`\\(X_2=0\\)` (control)] vs. .green[`\\(X_2=1\\)` (treatment)]

]
.pull-right[

```{r fig.retina=3}
treat=function(x){4+x}
control=function(x){2+0.5*x}

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=treat, geom="line", size=2, color="green")+
  geom_label(x=3, y=treat(3), color="green", label=expression(paste("When ",x[2]==1)))+
  geom_segment(aes(x=5, xend=6, y=treat(5), yend=treat(5)), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=6, xend=6, y=treat(5), yend=treat(6)), linetype="dotted", color="black", size=1)+
  stat_function(fun=control, geom="line", size=2, color="red")+
  geom_label(x=3, y=control(3), color="red", label=expression(paste("When ",x[2]==0)))+
  geom_segment(aes(x=5, xend=6, y=control(5), yend=control(5)), linetype="dotted", color="black", size=1)+
  geom_segment(aes(x=6, xend=6, y=control(5), yend=control(6)), linetype="dotted", color="black", size=1)+

  scale_x_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,10,1),
                     limits=c(0,10),
                     expand=expand_scale(mult=c(0,0.1)))+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```
]

---

# Polynomial Functions of $X$

$$\hat{Y_i} = \hat{\beta_0} + \hat{\beta_1} X_i + \hat{\beta_2} X_i^2 + \cdots + \hat{\beta_{\mathbf{r}}} X_i^{\mathbf{r}} + \epsilon_i$$

- Where $r$ is the highest power $X_i$ is raised to
  - e.g. quadratic $r=2$, cubic $r=3$, etc

- The graph of an $r$<sup>th</sup>-degree polynomial function has $(r-1)$ bends

- Just another multivariate OLS regression model

---

# Quadratic Model

$$\hat{Y_i} = \hat{\beta_0} + \hat{\beta_1} X_i + \hat{\beta_2} X_i^2$$

- .shout[Quadratic model] has $X$ and $X^2$ variables in it (yes, need both!)

--

- How to interpret coefficients (betas)?
  - $\beta_0$ as "intercept" and $\beta_1$ as "slope" makes no sense 
  - $\beta_1$ as effect $X_i \rightarrow Y_i$ holding $X_i^2$ constant makes no sense<sup>.red[1]</sup>

.footnote[<sup>.red[1]</sup> This is *not* a multicollinearity problem! Correlation only measures *linear* relationships!]

--

- **Calculate marginal effects** by calculating predicted $\hat{Y_i}$ for different levels of $X_i$

---

# Quadratic Model: Calculating Marginal Effects

$$\hat{Y_i} = \hat{\beta_0} + \hat{\beta_1} X_i + \hat{\beta_2} X_i^2$$

- What is the .shout[marginal effect] of $\Delta X_i \rightarrow \Delta Y_i$?

--

- Take the **derivative** of $Y_i$ with respect to $X_i$:
$$\frac{d Y_i}{d X_i} = \hat{\beta_1}+2\hat{\beta_2} X_i$$

--

- .onfire[Marginal effect] of a 1 unit change in $X_i$ is a $\left(\hat{\beta_1}+2\hat{\beta_2} X_i \right)$ unit change in $Y$

---

# Quadratic Model: Example I

.content-box-green[
.green[**Example**]: $$\widehat{\text{Life Expectancy}_i} = \hat{\beta_0}+\hat{\beta_1} \, \text{GDP per capita}_i+\hat{\beta_2}\, \text{GDP per capita}^2_i$$
]

- Use `gapminder` package and data

```{r, echo=T}
library(gapminder)
```

---

# Quadratic Model: Example II

- These coefficients will be very large, so let's transform `gdpPercap` to be in $1,000's

```{r, echo=T}
gapminder<-gapminder %>%
  mutate(GDP_t = gdpPercap/1000)
gapminder %>% head() # look at it
```

---

# Quadratic Model: Example III

- Let's also add the squared term, `gdp_sq`

```{r, echo=T}
gapminder<-gapminder %>%
  mutate(GDP_sq = GDP_t^2)
gapminder %>% head() # look at it
```

---

# Quadratic Model: Example IV

- Can "manually" run a multivariate regression with `GDP_t` and `GDP_sq`

.font80[
```{r, echo=T}
reg1<-lm(lifeExp ~ GDP_t + GDP_sq, data = gapminder)
summary(reg1)
```
]

---

# Quadratic Model: Example V

- OR use `gdp_t` and add the "transform" command in regression, `I(gdp_t^2)`

.font80[
```{r, echo=T}
reg1_alt<-lm(lifeExp ~ GDP_t + I(GDP_t^2), data = gapminder)
summary(reg1_alt)
```
]

---

# Quadratic Model: Example VI

.pull-left[
```{r quadreg, eval=F, echo=T}
library(broom)
tidy(reg1)
```
]

.pull-right[
```{r, ref.label="quadreg"}
```
]

--

$$\widehat{\text{Life Expectancy}_i} = 50.52+1.55 \, \text{GDP per capita}_i - 0.02\, \text{GDP per capita}^2_i$$

--

- Marginal effect of GDP per capita on Life Expectancy:

--

$$\begin{align*}
\frac{d \, Y}{d \; X} &= \hat{\beta_1}+2\hat{\beta_2} X_i\\
\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} &= 1.55+2(-0.02) \, \text{GDP}\\
 &= \mathbf{1.55-0.04 \, \text{GDP}}\\
\end{align*}$$

---

# Quadratic Model: Example VII

$$\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} = 1.55-0.04 \, \text{GDP}$$
- Positive, with diminishing returns

- Effect on Life Expectancy of increasing GDP depends on initial value of GDP!

---

# Quadratic Model: Example VIII

$$\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} = 1.55-0.04 \, \text{GDP}$$

Marginal effect of GDP if GDP $=\$5$ (thousand):

$$\begin{align*}
\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} &= 1.55-0.04\text{GDP}\\
&= 1.55-0.04(5)\\
&= 1.55-0.20\\
&=1.35\\
\end{align*}$$

--

- i.e. for every addition $1 (thousand) in GDP per capita, average life expectancy increases by 1.35 years

---

# Quadratic Model: Example IX

$$\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} = 1.55-0.04 \, \text{GDP}$$

Marginal effect of GDP if GDP $=\$25$ (thousand):

--

$$\begin{align*}
\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} &= 1.55-0.04\text{GDP}\\
&= 1.55-0.04(25)\\
&= 1.55-1.00\\
&=0.55\\
\end{align*}$$

--

- i.e. for every addition $1 (thousand) in GDP per capita, average life expectancy increases by 0.55 years

---

# Quadratic Model: Example X

$$\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} = 1.55-0.04 \, \text{GDP}$$

Marginal effect of GDP if GDP $=\$50$ (thousand):

--

$$\begin{align*}
\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} &= 1.55-0.04\text{GDP}\\
&= 1.55-0.04(50)\\
&= 1.55-2.00\\
&=-0.45\\
\end{align*}$$

--

- i.e. for every addition $1 (thousand) in GDP per capita, average life expectancy *decreases* by 0.45 years

---

# Quadratic Model: Example XI

$$\begin{align*}\widehat{\text{Life Expectancy}_i} &= 50.52+1.55 \, \text{GDP per capita}_i - 0.02\, \text{GDP per capita}^2_i \\
\frac{d \, \text{Life Expectancy}}{d \, \text{GDP}} &= 1.55-0.04\text{GDP} \\ \end{align*}$$

| *Initial* GDP per capita | Marginal Effect<sup>.red[1]<sup> |
|----------------|-------------------:|
| $5,000 | $1.35$ years |
| $25,000 | $0.55$ years |
| $50,000 | $-0.45$ years |

.footnote[<sup>.red[1]</sup> Of +$1,000 GDP/capita on Life Expectancy.]

---

# Quadratic Model: Example XII

.pull-left[

.font90[
```{r quadplot, echo=T, eval=F}

ggplot(data = gapminder)+
  aes(x = GDP_t,
      y = lifeExp)+
  geom_point(color="blue", alpha=0.5)+
  stat_smooth(method = "lm", #<<
              formula = y ~ x + I(x^2), #<<
              color="green")+ #<<
  geom_vline(xintercept=c(5,25,50),
             linetype="dashed",
             color="red", size = 1)+
  scale_x_continuous(labels=scales::dollar,
                     breaks=seq(0,120,10))+
  scale_y_continuous(breaks=seq(0,90,10),
                     limits=c(0,90))+
  labs(x = "GDP per Capita (in Thousands)",
       y = "Life Expectancy (Years)")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]
]

.pull-right[
```{r, ref.label="quadplot", fig.retina=3}
```
]

---

# Quadratic Model: Maxima and Minima I

- For a polynomial model, we can also find the predicted **maximum** or **minimum** of $\hat{Y_i}$

--

- A quadratic model has a single global maximum or minimum (1 bend)

--

- By calculus, a minimum or maximum occurs where:

$$\begin{align*}
		\frac{ d Y_i}{d X_i} &=0\\
		\beta_1 + 2\beta_2 X_i &= 0\\
		2\beta_2 X_i&= -\beta_1\\
		X_i^*&=-\frac{1}{2}\frac{\beta_1}{\beta_2}\\
		\end{align*}$$

---

# Quadratic Model: Maxima and Minima II

.pull-left[

```{}
tidy(reg1)
```
]
--

.pull-right[

$$\begin{align*}
GDP_i^*&=-\frac{1}{2}\frac{\beta_1}{\beta_2}\\
GDP_i^*&=-\frac{1}{2}\frac{(1.55)}{(-0.015)}\\
GDP_i^*& \approx -\frac{1}{2}(-103.333)\\
GDP_i^*& \approx 51.67\\ \end{align*}$$

]

---

# Quadratic Model: Maxima and Minima III

.pull-left[

```{r quadplotmax, echo=T, eval=F}
ggplot(data = gapminder)+
  aes(x = GDP_t,
      y = lifeExp)+
  geom_point(color="blue", alpha=0.5)+
  stat_smooth(method = "lm", #<<
              formula = y ~ x + I(x^2), #<<
              color="green")+ #<<
  geom_vline(xintercept=51.67, linetype="dashed", color="red", size = 1)+
  geom_label(x=51.67, y=90, label="$51.67", color="red")+
  scale_x_continuous(labels=scales::dollar,
                     breaks=seq(0,120,10))+
  scale_y_continuous(breaks=seq(0,90,10),
                     limits=c(0,90))+
  labs(x = "GDP per Capita (in Thousands)",
       y = "Life Expectancy (Years)")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[
```{r, ref.label="quadplotmax", fig.retina=3}
```
]

---

# Determining if Higher-Order Polynomials are Necessary I

```{r, echo=F}
tidy(reg1)
```

- Is the quadratic term necessary?

--

- Determine if $\hat{\beta_2}$ (on $X_i^2)$ is statistically significant:
  - $H_0: \hat{\beta_2}=0$
  - $H_a: \hat{\beta_2} \neq 0$

--

- Statistically significant $\implies$ we should keep the quadratic model
  - If we only ran a linear model, it would be incorrect!

---

# Determining if Higher-Order Polynomials are Necessary I

.pull-left[
- Should we keep going up in polynomials?

$$\widehat{\text{Life Expectancy}_i} = \hat{\beta_0}+\hat{\beta_1} GDP_i+\hat{\beta_2}GDP^2_i+\hat{\beta_3}GDP_i^3$$

]

.pull-right[
```{r, fig.retina=3}
p2+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color = "green")+
  stat_smooth(method="lm", formula=y~x+I(x^2)+I(x^3), color = "purple")
```

]

---

# Determining if Higher-Order Polynomials are Necessary II

.pull-left[

- In general, should have a compelling theoretical reason why data or relationships should "change direction" multiple times

- Or clear data patterns that have multiple "bends"

]

.pull-right[
```{r, fig.retina=3}
# make obviously cubic data
df_3<-tibble(x = seq(0,20,0.05),
           y = (500 + 0.4 * (x-10)^3)+rnorm(length(x), 10, 50)) # real function + random noise

ggplot(data = df_3)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  stat_smooth(method = "lm", formula = y~x+I(x^2)+I(x^3), color="red")+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

---

# A Second Polynomial Example I

.pull-left[

.content-box-green[
.green[**Example**]: How does a school district's average income affect Test scores?
]

$$\widehat{\text{Test Score}_i}=\hat{\beta_0}+\hat{\beta_1}\text{Income}_i$$

]

--

.pull-right[
```{r}
library(haven)
CASchool<-read_dta("../data/caschool.dta")
```

```{r, fig.retina=3}
income_plot<-ggplot(data = CASchool)+
  aes(x = avginc,
      y = testscr)+
  geom_point(color="blue")+
  scale_x_continuous(labels=scales::dollar)+
  scale_y_continuous(limits=c(600,710))+
  labs(x = "District Average Income (Thousands of $)",
       y = "Test Scores")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
income_plot
```
]

---

# A Second Polynomial Example I

.pull-left[

.content-box-green[
.green[**Example**]: How does a school district's average income affect Test scores?
]

$$\widehat{\text{Test Score}_i}=\hat{\beta_0}+\hat{\beta_1}\text{Income}_i$$

]

.pull-right[
```{r, fig.retina=3}
income_plot+geom_smooth(method="lm", color="red")
```
]

---

# A Second Polynomial Example I

.pull-left[

.content-box-green[
.green[**Example**]: How does a school district's average income affect Test scores?
]

$$\widehat{\text{Test Score}_i}=\hat{\beta_0}+\hat{\beta_1}\text{Income}_i$$

]

.pull-right[
```{r, fig.retina=3}
income_plot+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color="green")
```
]

---

# A Second Polynomial Example II

.font80[
```{r}
income_quad_reg<-lm(testscr~avginc+I(avginc^2), data = CASchool)
summary(income_quad_reg)
```
]

---

# A Second Polynomial Example III

.pull-left[
- Should we keep going?

.font60[
```{r}
income_quad_reg<-lm(testscr~avginc+I(avginc^2)+I(avginc^3), data = CASchool)
summary(income_quad_reg)
```
]
]

.pull-right[
```{r, fig.retina=3}
income_plot+geom_smooth(method="lm", color="red")+
  stat_smooth(method="lm", formula=y~x+I(x^2), color="green")+
  stat_smooth(method="lm", formula=y~x+I(x^2)+I(x^3), color="orange")
```

]

---

# Strategy for Polynomial Model Specification

1. Are there good theoretical reasons for relationships changing (e.g. increasing/decreasing returns)?

--

2. Plot your data: does a straight line fit well enough?

--

3. Specify a polynomial function of a higher power (start with 2) and estimate OLS regression

--

4. Use $t$-test to determine if higher-power term is significant

--

5. Interpret effect of change in $X$ on $Y$

--

6. Repeat steps 3-5 as necessary