### 2019-11-19 content/class/20-class.Rmd
Add class 20 content

---
title: "3.8: Polynomial Regression - Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, November 19, 2019*

# Overview

This week we will cover non-linear models, which we anticipated last week with an interaction effect between two continuous variables: $X_1 \times X_2$, where the marginal effect of $X_1 \rightarrow Y$ is dependent on the value of $X_2$.

We cover two types of nonlinear models. *Today*, we cover **polynomial models**, where we include higher-order $X$ variables to explain a non-linear effect. *Next* class we will cover **logarithmic models**, which are more frequently used - as well as a few final tests and transformations for multivariate regression before we turn to panel data.

Today, we will spend the first half of class finishing [last class' practice problems](/practice/19-r-practice) (on dummy variables and interaction effects), then I will cover polynomial models. After we finish next class' lecture, we will do another set of R practice problems, and you will be given **HW 5** to work on this material.

# Slides

- [<i class="fas fa-external-link-square-alt"></i> Lecture Slides](/slides/19-slides.html)

# R Practice Problems

We will begin by finishing your work on [Last Class' Practice Problems](/r/17-r-practice). Answers are posted there tonight.

# Problem Set 4 Due Thursday Nov 21

[Problm Set 4](/assignment/04-problem-set) (on classes 3.1-3.5) is due by Thursday November 21.