### 2019-12-05 static/slides/23-slides.Rmd
Final tweak for slides 23

---
title: "4.2: Difference-in-Difference Models"
subtitle: "ECON 480 · Econometrics · Fall 2019"
author: 'Ryan Safner<br> Assistant Professor of Economics <br> <a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i> safner@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsf19"><i class="fa fa-github fa-fw"></i> ryansafner/metricsf19</a><br> <a href="https://metricsF19.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i> metricsF19.classes.ryansafner.com</a><br>'
#date:
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML" # rescales math with css changes https://github.com/yihui/xaringan/issues/143
    lib_dir: libs
    df_print: paged
    #seal: false
    css: [custom.css, custom-fonts.css, "hygge"] #, metropolis, metropolis-fonts
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"] # first is for rescaling images , second is for embedding tweets, https://github.com/yihui/xaringan/issues/100
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
    includes:
      in_header: header.html # for font awesome, used in title  
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
library(tidyverse)
set.seed(256)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
```

# Clever Research Designs Identify Causality

Again, **this toolkit** of research designs to **identify causal effects** is the economist's **comparative advantage** that firms and governments want!

```{r, fig.retina=3, fig.align="center", fig.width=12}
ggplot(data = tibble(x=c(0,10),
                     y=c(0,10)))+
  aes(x = x,
      y = y)+
  geom_label(x=1,y=2, label="Correlation", size =10, color = "#fde0dd")+
  geom_label(x=5,y=2, label="Causation", size = 10, color = "#7a0177")+

  geom_label(x=1,y=4, label="Differences", size =5, color = "#fde0dd")+
  geom_label(x=1,y=3.5, label="Pre-Post", size =5, color = "#fde0dd")+
  geom_label(x=2,y=4.5, label="Multiple Regression", size =5, color = "#fbb4b9")+
  geom_label(x=2,y=3.5, label="Matching", size =5, color = "#fbb4b9")+
    geom_label(x=3,y=5, label="Fixed Effects", size =5, color = "#f768a1")+
    geom_label(x=3.5,y=4.5, label="Diff-in-Diff", size =5, color = "#c51b8a")+
  geom_label(x=4,y=3.5, label="Natural Experiments", size =5, color = "#c51b8a")+
  geom_label(x=3.5,y=4, label="Regression Discontinuity", size =5, color = "#c51b8a")+
  geom_label(x=5,y=4.5, label="RCTs", size =5, color = "#7a0177")+
  annotate("segment", x = 1, xend = 5, y = 3, yend = 3, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="last", type="closed"))+
  scale_x_continuous(breaks=seq(0,6,1),
                     limits=c(0,6))+
  scale_y_continuous(breaks=seq(1,5,1),
                     limits=c(1,5))+
  theme_void()

```

---

class: inverse, center, middle

# Difference-in-Difference Models

---

# Difference-in-Difference Models I

.pull-left[

- Often, we want to examine the consequences of a change, such as a law or policy

]

---

# Difference-in-Difference Models I

.pull-left[

- Often, we want to examine the consequences of a change, such as a law or policy

- .green[**Example**]: how do States that implement law $X$ see changes in $Y$
  - **Treatment**: States that implement law $X$
  - **Control**: States that did not implement law $X$

]

---

# Difference-in-Difference Models I

.pull-left[

- Often, we want to examine the consequences of a change, such as a law or policy

- .green[**Example**]: how do States that implement law $X$ see changes in $Y$
  - **Treatment**: States that implement law $X$
  - **Control**: States that did not implement law $X$

- If we have **panel data** with observations for all states before *and* after the change...

- Find the *difference* between treatment & control groups *in* their *differences* before and after the treatment period

]

--

.pull-right[
.center[
![](https://www.dropbox.com/s/vpk560fok07yo9t/dndhot.jpg?raw=1)
]
]

---

# Difference-in-Difference Models II

- The .shout[difference-in-difference model] (aka .shout["diff-in-diff"] or .shout[DND"]) identifies treatment effect by differencing the difference pre- and post-treatment values of $Y$ between treatment and control groups

--

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

--

- $Treated_i= \begin{cases}1 \text{ if } i \text{ is in treatment group}\\ 0 \text{ if } i \text{ is not in treatment group}\end{cases}$

- $After_t= \begin{cases}1 \text{ if } t \text{ is after treatment period}\\ 0 \text{ if } t \text{ is before treatment period}\end{cases}$

--

|    | Control | Treatment | **Group Diff** $(\Delta Y_i)$ |
|----|---------|-----------|---------------------------|
| Before | $\beta_0$ | $\beta_0+\beta_1$ | $\beta_1$ |
| After | $\beta_0+\beta_2$ | $\beta_0+\beta_1+\beta_2+\beta_3$ | $\beta_1+\beta_3$ |
| **Time Diff** $(\Delta Y_t)$ | $\beta_2$ | $\beta_2+\beta_3$ | **Diff-in-diff** $\Delta_i \Delta_t: \beta_3$ |

---

```{r}
control<-tribble(
  ~x, ~y,
  1, 1,
  2, 3
)
treatment<-tribble(
  ~x, ~y,
  1, 4,
  2, 8
)
```

# Visualizing Diff-in-Diff

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "red", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 2
  geom_label(x=2.1,y=2, color = "red", label=expression(hat(beta)[2]))+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- .red[Control group `\\((Treated = 0)\\)`]

- .red[`\\(\hat{\beta_0}\\)`]: value of $Y$ for **control** group **before** treatment

- .red[`\\(\hat{\beta_2}\\)`]: time *difference* (for **control** group)

]

---

# Visualizing Diff-in-Diff

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "red", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 2
  geom_label(x=2.1,y=2, color = "red", label=expression(hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "green", label=expression(hat(beta)[0]+hat(beta)[1]))+

    # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # beta 2
  geom_label(x=2.1,y=5, color = "red",label=expression(hat(beta)[2]))+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- .red[Control group `\\((Treated = 0)\\)`]

- .red[`\\(\hat{\beta_0}\\)`]: value of $Y$ for **control** group **before** treatment

- .red[`\\(\hat{\beta_2}\\)`]: time *difference* (for **control** group)

- .green[Treated group `\\((Treated = 1)\\)`]

]

---

# Visualizing Diff-in-Diff

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "red", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 2
  geom_label(x=2.1,y=2, color = "red", label=expression(hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "green", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # beta 1 dif
  geom_segment(x=1, y=1, xend=1, yend=4, color="black", linetype = "dashed", size = 1)+
  geom_label(x=0.9,y=2.5, color = "black", label=expression(hat(beta)[1]))+

    # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # beta 2
  geom_label(x=2.1,y=5, color = "red",label=expression(hat(beta)[2]))+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- .red[Control group `\\((Treated = 0)\\)`]

- .red[`\\(\hat{\beta_0}\\)`]: value of $Y$ for **control** group **before** treatment

- .red[`\\(\hat{\beta_2}\\)`]: time *difference* (for **control** group)

- .green[Treated group `\\((Treated = 1)\\)`]

- $\hat{\beta_1}$: *difference* between groups **before** treatment

]

---

# Visualizing Diff-in-Diff

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "red", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 2
  geom_label(x=2.1,y=2, color = "red", label=expression(hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "green", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # beta 1 dif
  geom_segment(x=1, y=1, xend=1, yend=4, color="black", linetype = "dashed", size = 1)+
  geom_label(x=0.9,y=2.5, color = "black", label=expression(hat(beta)[1]))+

    # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  # beta 2
  geom_label(x=2.1,y=5, color = "red",label=expression(hat(beta)[2]))+

  # beta 3
  geom_label(x=2.1,y=7, color = "blue", label=expression(hat(beta)[3]))+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- .red[Control group `\\((Treated = 0)\\)`]

- .red[`\\(\hat{\beta_0}\\)`]: value of $Y$ for **control** group **before** treatment

- .red[`\\(\hat{\beta_2}\\)`]: time *difference* (for **control** group)

- .green[Treated group `\\((Treated = 1)\\)`]

- $\hat{\beta_1}$: *difference* between groups **before** treatment

- .blue[`\\(\hat{\beta_3}\\)`]: difference-in-difference: .blue[treatment effect]
]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "black", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

- $Y_i$ for **Treatment** group **before**: $\hat{\beta_0}+\hat{\beta_1}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "black", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  # beta 3
  geom_label(x=2.25,y=8, color = "black",
             label=expression(hat(beta)[0]+hat(beta)[1]+hat(beta)[2]+hat(beta)[3]))+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

- $Y_i$ for **Treatment** group **before**: $\hat{\beta_0}+\hat{\beta_1}$

- $Y_i$ for **Treatment** group **after**: $\hat{\beta_0}+\hat{\beta_1}+\hat{\beta_2}+\hat{\beta_3}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "black", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  # beta 3
  geom_label(x=2.25,y=8, color = "black",
             label=expression(hat(beta)[0]+hat(beta)[1]+hat(beta)[2]+hat(beta)[3]))+

  # beta 1 dif
  annotate("segment", x = 1, xend = 1, y = 1, yend = 4, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=1,y=2.5, size = 5, color = "black", label=expression(hat(beta)[1]))+

    scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

- $Y_i$ for **Treatment** group **before**: $\hat{\beta_0}+\hat{\beta_1}$

- $Y_i$ for **Treatment** group **after**: $\hat{\beta_0}+\hat{\beta_1}+\hat{\beta_2}+\hat{\beta_3}$

- **Group Difference (before)**: $\hat{\beta_1}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "black", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  # beta 3
  geom_label(x=2.25,y=8, color = "black",
             label=expression(hat(beta)[0]+hat(beta)[1]+hat(beta)[2]+hat(beta)[3]))+

  # beta 1 dif
  annotate("segment", x = 1, xend = 1, y = 1, yend = 4, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=1,y=2.5, size = 5, color = "black", label=expression(hat(beta)[1]))+

  # beta 2 dif
  annotate("segment", x = 2, xend = 2, y = 1, yend = 3, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=2,y=2, size = 5, color = "black", label=expression(hat(beta)[2]))+

  # beta 2 dif
  annotate("segment", x = 2, xend = 2, y = 4, yend = 6, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=2,y=5, size = 5, color = "black", label=expression(hat(beta)[2]))+

    scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

- $Y_i$ for **Treatment** group **before**: $\hat{\beta_0}+\hat{\beta_1}$

- $Y_i$ for **Treatment** group **after**: $\hat{\beta_0}+\hat{\beta_1}+\hat{\beta_2}+\hat{\beta_3}$

- **Group Difference (before)**: $\hat{\beta_1}$

- **Time Difference**: $\hat{\beta_2}$

]

---

# Visualizing Diff-in-Diff II

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # beta 0 
  geom_label(x=0.9,y=1, color = "black", label=expression(hat(beta)[0]))+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+
  # beta 1+2
  geom_label(x=2.2,y=3, color = "black", label=expression(hat(beta)[0]+hat(beta)[2]))+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # beta 0 plus beta 1 
  geom_label(x=0.8,y=4, color = "black", label=expression(hat(beta)[0]+hat(beta)[1]))+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  # beta 3
  geom_label(x=2.25,y=8, color = "black",
             label=expression(hat(beta)[0]+hat(beta)[1]+hat(beta)[2]+hat(beta)[3]))+

  # beta 1 dif
  annotate("segment", x = 1, xend = 1, y = 1, yend = 4, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=1,y=2.5, size = 5, color = "black", label=expression(hat(beta)[1]))+

  # beta 2 dif
  annotate("segment", x = 2, xend = 2, y = 1, yend = 3, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=2,y=2, size = 5, color = "black", label=expression(hat(beta)[2]))+

  # beta 2 dif
  annotate("segment", x = 2, xend = 2, y = 4, yend = 6, colour = "black", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=2,y=5, size = 5, color = "black", label=expression(hat(beta)[2]))+

  # beta 2 dif
  annotate("segment", x = 2, xend = 2, y = 6, yend = 8, colour = "blue", size=2, alpha=1, arrow=arrow(length=unit(0.5,"cm"), ends="both", type="closed"))+
  geom_label(x=2,y=7, size = 5, color = "blue", label=expression(hat(beta)[3]))+

    scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- $Y_i$ for **Control** group **before**: $\hat{\beta_0}$

- $Y_i$ for **Control** group **after**: $\hat{\beta_0}+\hat{\beta_2}$

- $Y_i$ for **Treatment** group **before**: $\hat{\beta_0}+\hat{\beta_1}$

- $Y_i$ for **Treatment** group **after**: $\hat{\beta_0}+\hat{\beta_1}+\hat{\beta_2}+\hat{\beta_3}$

- **Group Difference (before)**: $\hat{\beta_1}$

- **Time Difference**: $\hat{\beta_2}$

- .blue[**Difference-in-difference**: `\\(\hat{\beta_3}\\)` (treatment effect)]
]

---

# Comparing Group Means (Again)

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

|    | Control | Treatment | **Group Diff** $(\Delta Y_i)$ |
|----|---------|-----------|---------------------------|
| Before | $\beta_0$ | $\beta_0+\beta_1$ | $\beta_1$ |
| After | $\beta_0+\beta_2$ | $\beta_0+\beta_1+\beta_2+\beta_3$ | $\beta_1+\beta_3$ |
| **Time Diff** $(\Delta Y_t)$ | $\beta_2$ | $\beta_2+\beta_3$ | **Diff-in-diff** $\Delta_i \Delta_t: \beta_3$ |

---

# Key Assumption: Counterfactual

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=6, color="red", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=6, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- Key assumption for DND: **time trends** (for treatment and control) are **parallel**

- Treatment and control groups assumed to be identical over time on average, **except for treatment**

- .shout[Counterfactual]: if the treatment group had not recieved treatment, it would have changed identically over time as the control group $(\hat{\beta_2})$

]

---

# Key Assumption: Counterfactual

$$\hat{Y_{it}}=\beta_0+\beta_1 \text{Treated}_i +\beta_2 \text{After}_{t}+\beta_3 (\text{Treated}_i \times \text{After}_{t})+u_{it}$$

.pull-left[

```{r, fig.height=6, fig.retina=3}
ggplot(data = tibble(x=1,3))+
  aes(x = x)+
  # control
  geom_point(data = control, aes(x=x,y=y), size = 3, color = "red")+
  geom_text(x=1,y=0.5, color = "red", label=expression(C[1]))+
  geom_text(x=2,y=3.5, color = "red", label=expression(C[2]))+

  # line
  geom_segment(x=1, y=1, xend=2, yend=3, color="red", size = 2)+

  # line slope
  geom_segment(x=1, y=1, xend=2, yend=1, color="red", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=1, xend=2, yend=3, color="red", linetype = "dotted", size = 1)+

  # treatment
  geom_point(data = treatment, aes(x=x,y=y), size = 3, color = "green")+
  geom_text(x=1,y=4.5, color = "green", label=expression(T[1]))+
  geom_text(x=2,y=8.5, color = "green", label=expression(T[2]))+

    # line
  geom_segment(x=1, y=4, xend=2, yend=8, color="green", size = 2)+

  # line slope
  geom_segment(x=1, y=4, xend=2, yend=4, color="green", linetype = "dotted", size = 1)+
  geom_segment(x=2, y=4, xend=2, yend=4.5, color="purple", linetype = "dotted", size = 1)+

  # diag
  geom_segment(x=1, y=4, xend=2, yend=4.5, color="purple", linetype = "dotted", size = 1)+

  geom_segment(x=2, y=4.5, xend=2, yend=8, color="blue", linetype = "dashed", size = 1)+

  scale_x_continuous(breaks=c(1,2),
                  labels=c(expression(t[before]),expression(t[after])),
                  limits=c(0.5,2.5))+
  scale_y_continuous(breaks=NULL,
                     limits=c(0,9))+
  labs(x = "Time",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```
]

.pull-right[

- If the time-trends would have been *different*, a **biased** measure of the treatment effect $(\hat{\beta_3})$!

]

---

# Diff-in-Diff Example I

.content-box-green[
.green[**Example**]: In 1993 Georgia initiated a HOPE scholarship program to let state residents with at least a B average in high school attend public college in Georgia for free. Did it increase college enrollment?
]

- Micro-level data on 4,291 young individuals

--

- $\text{InCollege}_{it}=\begin{cases}1 \text{ if } i \text{ is in college during year }t\\ 0 \text{ if } i \text{ is not in college during year }t\\ \end{cases}$

--

- $\text{Georgia}_i=\begin{cases}1 \text{ if } i \text{ is a Georgia resident}\\ 0 \text{ if } i \text{ is not a Georgia resident}\\ \end{cases}$

--

- $\text{After}_t=\begin{cases}1 \text{ if } t \text{ is after 1992}\\ 0 \text{ if } t \text{ is after 1992}\\ \end{cases}$

.source[Dynarski, Susan (2000), "Hope for Whom? Financial Aid for the Middle Class and Its Impact on College Attendance"]

.footnote[Note: With a dummy *dependent* (Y) variable, coefficients estimate the probability `\\(Y=1\\)`, i.e. the *probability* a person is enrolled in college]

---

# Diff-in-Diff Example II

- We can use a DND model to measure the effect of HOPE scholarship on enrollments

- Georgia and nearby States, if not for HOPE, changes should be the same over time

--

- Treatment period: after 1992

--

- Treatment: Georgia

--

- Differences-in-differences: $$\Delta_i \Delta_t Enrolled = (\text{GA}_{after}-\text{GA}_{before})-(\text{neighbors}_{after}-\text{neighbors}_{before})$$

--

- Regression equation:
$$\widehat{Enrolled_{it}} = \beta_0+\beta_1 \, Georgia_{i}+\beta_2 \,  After_{t}+\beta_3 \, (Georgia_{i} \times After_{t})$$

---

# Example: Regression

```{r}
hope<-read_csv("https://metricsf19.classes.ryansafner.com/data/HOPE.csv")
hope<-hope %>%
  mutate_at(c("StateCode", "Year"),factor)
```

.pull-left[

.font70[
```{r, echo=T}
DND_reg<-lm(InCollege ~ Georgia + After + Georgia:After, data = hope)
summary(DND_reg)
```
]
]

--

.pull-right[
$$\widehat{Enrolled_{it}}=0.406-0.105 \, Georgia_{i}-0.004 \, After_{t}+0.089 \, (Georgia_{i} \times After_{t})$$
]

---

# Example: Interpretting the Regression

$$\widehat{Enrolled_{it}}=0.406-0.105 \, Georgia_{i}-0.004 \, After_{t}+0.089 \, (Georgia_{i} \times After_{t})$$

--

- $\beta_0$: 
--
A **non-Georgian** **before** 1992 was 40.6% likely to be a college student

--

- $\beta_1$: 
--
**Georgians** **before** 1992 were 10.5% less likely to be college students than neighboring states

--

- $\beta_2$: 
--
**After** 1992, **non-Georgians** are 0.4% less likely to be college students

--

- $\beta_3$: 
--
**After** 1992, **Georgians** are 8.9% more likely to enroll in colleges than neighboring states

--

- .onfire[Treatment effect: HOPE increased enrollment likelihood by 8.9%]

---

# Example: Comparing Group Means

$$\widehat{Enrolled_{it}}=0.406-0.105 \, Georgia_{i}-0.004 \, After_{t}+0.089 \, (Georgia_{i} \times After_{t})$$

- A group mean for a dummy $Y$ is $E[Y=1]$, i.e. the probability a student is enrolled:

- **Non-Georgian enrollment probability pre-1992**: 
--
$\beta_0=0.406$

--
- **Georgian enrollment probability pre-1992**: 
--
$\beta_0+\beta_1=0.406-0.105=0.301$

--

- **Non-Georgian enrollment probability post-1992**: 
--
$\beta_0+\beta_2=0.406-0.004=0.402$

--

- **Georgian enrollment probability post-1992**: 
--
$\beta_0+\beta_1+\beta_2+\beta_3=0.406-0.105-0.004+0.089=0.386$

---

# Example: Comparing Group Means in R

.pull-left[
```{r, echo=T}
# group mean for non-Georgian before 1992
hope %>%
  filter(Georgia==0,
         After==0) %>%
  summarize(prob = mean(InCollege))

```
]

--

.pull-right[
```{r, echo=T}
# group mean for non-Georgian AFTER 1992
hope %>%
  filter(Georgia==0,
         After==1) %>%
  summarize(prob = mean(InCollege))

```
]

---

# Example: Comparing Group Means in R II

.pull-left[
```{r, echo=T}
# group mean for Georgian before 1992
hope %>%
  filter(Georgia==1,
         After==0) %>%
  summarize(prob = mean(InCollege))
```
]

--

.pull-right[
```{r, echo=T}
# group mean for Georgian AFTER 1992
hope %>%
  filter(Georgia==1,
         After==1) %>%
  summarize(prob = mean(InCollege))
```
]

---

# Example: Diff-in-Diff Summary

$$\widehat{Enrolled_{it}}=0.406-0.105 \, Georgia_{i}-0.004 \, After_{t}+0.089 \, (Georgia_{i} \times After_{t})$$

|    | Neighbors | Georgia | **Group Diff** $(\Delta Y_i)$ |
|----|--------:|----------:|--------------------------:|
| Before | $0.406$ | $0.301$ | $-0.105$ |
| After | $0.402$ | $0.386$ | $0.016$ |
| **Time Diff** $(\Delta Y_t)$ | $-0.004$ | $0.085$ | **Diff-in-diff**: $0.089$ |

--

$$\begin{align*}
	\Delta_i \Delta_t Enrolled &= (\text{GA}_{after}-\text{GA}_{before})-(\text{neighbors}_{after}-\text{neighbors}_{before})\\
	&=(0.386-0.301)-(0.402-0.406)\\
	&=(0.085)-(-0.004)\\
	&=0.089\\ \end{align*}$$

---

# Example: Diff-in-Diff Graph

```{r, fig.retina=3, fig.align="center", fig.width=16}
hope %>%
  ggplot(data=.)+
  aes(x = After,
      y = InCollege,
      color = factor(Georgia))+
  geom_smooth(method="lm", se=FALSE)+
  #scale_y_continuous(breaks=seq(0,1,0.125), limits=c(0,0.05))+
  scale_x_continuous(breaks=seq(0,1,1), labels=c("Before", "After"))+
scale_color_manual(name = "State",
                   labels = c("Neighbors", "Georgia"), values = c("blue", "red"))+
  labs(x = "Before or After HOPE",
       y = "Probability of Being Enrolled in College")+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=16)+
  theme(legend.position = "top")
```

---

class: inverse, center, middle

# Generalizing DND Models

---

# Generalizing DND Models

- DND can be **generalized** with a **two-way fixed effects** model:  	
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 (\text{Treated}_i * \text{After}_{t})+\nu_{it}$$
    - $\alpha_i$: **group fixed effects** (treatments/control groups)
    - $\theta_t$: **time fixed effects** (pre/post treatment)

--

- Allows *many* periods, and treatment(s) can occur at different times to different units (so long as some do not get treated)

--

- Can also add control variables that vary within units and over time 
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 \, (\text{Treated}_i \times \text{After}_{t})+\beta_4 X_{it}+\nu_{it}$$

---

# Our Example, Generalized I

$$\widehat{Enrolled_{it}} = \alpha_i+\theta_t+\beta_3 \, (Georgia_{it} \times After_{it})$$

- `StateCode` is a variable for the State $\implies$ create State fixed effect

- `Year` is a variable for the year $\implies$ create year fixed effect

- Using LSDV method...

```{r dnd-gen-reg, echo=T, eval=F}
DND_fe<-lm(InCollege ~ Georgia:After + factor(StateCode) + factor(Year),
           data = hope)
summary(DND_fe)
```

---

# Our Example, Generalized II

```{css, echo = FALSE}
.smallcode .remark-code { /*Change made here*/
  font-size: 55% !important;
}
```

.pull-left[
.smallcode[
```{r, ref.label="dnd-gen-reg"}

```
]
]

--

.pull-right[
$$\widehat{InCollege_{it}}=\alpha_i+\theta_t+0.091 \, (\text{Georgia}_i \times \text{After}_{it})$$

]

---

# Intuition behind DND

- Diff-in-diff models are the quintessential example of exploiting .shout[natural experiments]

- A major change at a point in time (change in law, a natural disaster, political crisis) separates groups where one is affected and another is not---identifies the effect of the change (treatment)

- One of the cleanest and clearest causal **identification strategies**

---

class: inverse, center, middle

# Example: "The" Card-Kreuger Minimum Wage Study

---

# Example: "The" Card-Kreuger Minimum Wage Study I

.content-box-green[
.green[**Example**]: *The* controversial minimum wage study, Card & Kreuger (1994) is a quintessential (and clever) diff-in-diff.
]

.source[Card, David, Krueger, Alan B, (1994), "Minimum Wages and Employment: A Case Study of the Fast-Food Industry in New Jersey and Pennsylvania," *American Economic Review* 84 (4): 772–793	]

---

# Card & Kreuger (1994): Background I

.pull-left[
- Card & Kreuger (1994)	compare employment in fast food restaurants on New Jersey and Pennsylvania sides of border between February and November 1992.

- Pennsylvania & New Jersey both had a minimum wage of $4.25 before February 1992

- In February 1992, New Jersey raised minimum wage from $4.25 to $5.05
]

.pull-right[
.center[
![](https://www.dropbox.com/s/gdoip37e26j253b/PANJ.png?raw=1)
]
]

---

# Card & Kreuger (1994): Background II

.pull-left[
- If we look only at New Jersey before & after change: 
  - **Omitted variable bias**: macroeconomic variables (there's a recession!), weather, etc.
	- Including PA as a control will control for these time-varying effects if they are national trends

- Surveyed 400 fast food restaurants on each side of the border, before & after min wage increase
  - .onfire[Key assumption]: Pennsylvania and New Jersey follow parallel trends, 
	- **Counterfactual**: if not for the minimum wage increase, NJ employment would have changed similar to PA employment
]

.pull-right[
.center[
![](https://www.dropbox.com/s/gdoip37e26j253b/PANJ.png?raw=1)
]
]

---

# Card & Kreuger (1994): Comparisons

.center[
![:scale 25%](https://www.dropbox.com/s/d0z7vb9so296eun/CardKreuger1.png?raw=1)
]

---

# Card & Kreuger (1994): Summary I

.center[
![:scale 65%](https://www.dropbox.com/s/9igiex5ly8u67vq/CardKreuger2.png?raw=1)
]

---

# Card & Kreuger (1994): Summary II

.center[
![:scale 65%](https://www.dropbox.com/s/gj4spkyz3jkop2z/CardKreuger3.png?raw=1)
]

---

# Card & Kreuger (1994): Identification Strategy and Model

$$\widehat{Employment_{i t}}=\beta_0+\beta_1 \, NJ_{i}+\beta_2 \, After_{t}+\beta_3 \, (NJ_i \times After_t)$$

- PA Before: $\beta_0$

- PA After: $\beta_0+\beta_2$

- NJ Before: $\beta_0+\beta_1$

- NJ After: $\beta_0+\beta_1+\beta_2+\beta_3$

- **Diff-in-diff**: $(NJ_{after}-NJ_{before})-(PA_{after}-PA_{before})$

--

|    | PA | NJ | **Group Diff** $(\Delta Y_i)$ |
|----|---------|-----------|---------------------------|
| Before | $\beta_0$ | $\beta_0+\beta_1$ | $\beta_1$ |
| After | $\beta_0+\beta_2$ | $\beta_0+\beta_1+\beta_2+\beta_3$ | $\beta_1+\beta_3$ |
| **Time Diff** $(\Delta Y_t)$ | $\beta_2$ | $\beta_2+\beta_3$ | **Diff-in-diff** $\Delta_i \Delta_t: \beta_3$ |

---

# Card & Kreuger (1994): Results

.center[
![:scale 65%](https://www.dropbox.com/s/ms5gpn8lfkws6a9/CardKreuger5.png?raw=1)
]