### 2019-12-06 content/assignment/02-exam.Rmd
Add HW 5 answers, HW 6 and answers, finalize schedule page (put extra content after final), add final exam prep

---
title: "Final Exam"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Thursday, December 12, 11:30 AM -2:00 PM*

# Concepts Study Guide

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/oldexams/finalconcepts.pdf)
- [<i class="fas fa-external-link-square-alt"></i> Review Questions Slides (TBD)](/slides/final-review.html)

# "Practice" Exam

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/oldexams/02-old-exam.pdf)

# Formula Sheet (On Exam)

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/oldexams/finalformulas.pdf)

# Exam Information

**The exact number and type of questions may change.** Note both the concepts study guide and the

**You may not have anything with you** for the exam (no notes, etc) **except a calculator.** I write questions such that you can get a perfect score without a calculator, and answers should typically be simple whole numbers. However, I understand that if nothing else, calculators are moral support and you can use them. I will provide simple calculators if you need to borrow one, as well as extra paper.

**There will be no use of R on the exam**. However, you *should* be familiar with certain outputs of R that I can ask you questions about (such as regression output from `lm()`).

**There will be few, if any, calculation questions**. Questions are more about *interpretation* of concepts and interpretting numbers in the context of various regressions and research questions.

Students that require accommodations and students that know they will be absent must confirm with me any arrangements at least a week *in advance* of the Exam.

# My Advice

Make sure you do all of the homework problems and learn from the answer keys to the homeworks, as well as the in-class practice problems. While some of the questions should be novel applications, conceptual questions on homeworks will get you in the right headspace to think about answering a question on an exam.