### 2019-12-06 static/homeworks/06-problem-set.Rmd
Add HW 6 empty markdown file, answers to #1-2 in HW 6 answers

---
title: "Problem Set 6 (Answers)"
author: "Ryan Safner" # put your name here! 
date: "ECON 480 - Fall 2019"
output:
  html_document:
    df_print: paged
    #theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

<!--CLICK "KNIT" ABOVE TO RENDER TO HTML, PDF, OR WORD OUTPUT
In fact, try knitting right away and see what this produces!
-->

*Due by Thursday, November 21, 2019*

# Instructions

Answers may be longer than I would deem sufficient on an exam. Some might vary slightly based on points of interest, examples, or personal experience. These suggested answers are designed to give you both the answer and a short explanation of *why* it is the answer.

# Theory and Concepts

## Question 1
**In your own words, describe what *fixed effects* are, when we can use them, and how they remove endogeneity.**

---

<!--WRITE YOUR ANSWERS BELOW -->

---

## Question 2
**In your own words, describe the logic of a *difference-in-difference* model: what is it comparing against what, and how does it estimate the effect of treatment? What assumption must be made about the treatment and control group for the model to be valid?**

---

<!--WRITE YOUR ANSWERS BELOW -->

---

# R Questions

Answer the following questions using `R`. When necessary, please write answers in the same document (knitted `Rmd` to `html` or `pdf`, typed `.doc(x)`, or handwritten) as your answers to the above questions. Be sure to include (email or print an `.R` file, or show in your knitted `markdown`) your code and the outputs of your code with the rest of your answers.

## Question 3

- [<i class="fas fa-table"></i> `PeaceCorps.csv`](http://metricsf19.classes.ryansafner.com/data/PeacceCorps.csv)

How do people respond to changes in economic conditions? Are they more likely to pursue public service when private sector jobs are scarce? This dataset contains variables at the U.S. State (& D.C.) level:

| Variable | Description |
|----------|-------------|
| `state` | U.S. State |
| `year` | Year |
| `appspc` | Applications to the Peace Corps (per capita) in State |
| `unemployrate` | State unemployment rate |

> Do more people apply to the Peace Corps when unemployment increases (and reduces other opportunities)?

### Part A

**Before looking at the data, what does your economic intuition tell you? Explain your hypothesis.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part B

**To get the hang of the data we're working with, `count` (separately) the number of `state`s, and the number of `year`s. Get the number of `n_distinct()` `state`s and `year`s^[Do this inside the `summarize()` command], as well as the `distinct()` values of each^[Don't use the `summarize()` command for this part].**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part C

**Continuing our pre-analysis inspection, (install, and) load the `plm` package, and check the dimensions of the data with `pdim`.**^[Set `index=c("state","year")` to indicate the group and time dimensions.]

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part D

**Create a scatterplot of `appspc` (Y) on `unemployrate` (X). Which State is an outlier? How would this affect the pooled regression estimates? Create a *second* scatterplot that does not include this State.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part E

**Run two *pooled* regressions, one with the outliers, and one without them. Write out the estimated regression equation for each. Interpret the coefficient, and comment on how it changes between the two regressions.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part F

**Now run a regression with State fixed effects using the dummy variable method.^[Ensure that `state` is a factor variable, and insert in the regression. You can either `mutate()` it into a `factor` beforehand, or just do `as.factor(state)` in the `lm` command.] Interpret the marginal effect of `unemployrate` on `appspc`. How did it change?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part G

**Find the coefficient for Maryland and interpret it. How many applications per capita does Maryland have?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part H

**Now try using the `plm()` command, which de-means the data, and make sure you get the same results as Part F.^[Inside `plm()`, set `index = "state"` to indicate variable, and `model = "within"` to indicate a fixed effects model.] Do you get the same marginal effect of `unemployrate` on `appspc`?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part I

**Now include *year* fixed effects in your regression, using the dummy variable method. Interpret the marginal effect of `unemployrate` on `appspc`. How did it change?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part I

**What would be the predicted number of applications in Maryland in 2011 at an unemployment rate of 5%?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part K

**Now try using the `plm()` command, which de-means the data, and make sure you get the same results as Part I.^[Inside `plm()`, set `index = c("state", "year")` to indicate both variables, and `effect = "twoways"` to indicate a 2-way fixed effects model.] Do you get the same marginal effect of `unemployrate` on `appspc`?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part L

**Can there still be endogeneity in this model? Give some examples.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part M
**Create a nice regression table (using `huxtable`) for comparison of the regressions in E, G, and I.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

## Question 4

- [<i class="fas fa-table"></i> `TexasSchools.csv`](http://metricsf19.classes.ryansafner.com/data/TexasSchools.csv)

Are teachers paid more when school board members are elected "off cycle" when there are not major national political elections (e.g. odd years) than "on cycle?" The argument is that during "off" years, without attention on state or national elections, voters will pay less attention to the election, and teachers can more effectively mobilize for higher pay, versus "on" years where voters are paying more attention. This data comes from Anzia, Sarah (2012), "The Election Timing Effect: Evidence from a Policy Intervention in Texas." *Quarterly Journal of Political Science* 7(3): 277-297, and follows 1,020 Texas school board districts from 2003–2009.

From 2003–2006, all districts elected their school board members off-cycle. A change in Texas policy in 2006 led some, but not all, districts to elect their school board members on-cycle from 2007 onwards.

| Variable | Description |
|----------|-------------|
| `LnAvgSalary` | logged average salary of teachers in district |
| `Year` | Year |
| `OnCycle` | `=1` if school boards elected on-cycle (e.g. same year as national and state elections), `=0` if elected off-cycle |
| `pol_freedom` | Political freedom index score (2018) from 1 (least) top 10 (most free) |
| `CycleSwitch` | `=1` if district switched from off- to on-cycle elections |
| `AfterSwitch` | `=1` if year is after 2006 |

### Part A

**Run a pooled regression model of `LnAvgSalary` on `OnCycle`. Write the estimated regression equation, and interpret the coefficient on `OnCycle.` Are there any sources of bias (consider in particular the argument in the question prompt)?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part B

**Some schools decided to switch to an on-cycle election after 2006. Consider this, `CycleSwitch` the "treatment." Create a variable to indicate post-treatment years (i.e. years after 2006). Call it `After`. Create a second, *interaction* variable to capture the interaction effect between those districts that *switched*, and *after* the treatment.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part C

**Now estimate a difference-in-difference model with your variables in Part B: `CycleSwitch` is the treatment variable, `After` is your post-treatment indicator, and add an _interaction_ variable to capture the interaction effect between those districts that _switched_, and _after_ the treatment. Write down the estimated regression equation (to four decimal places). **

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part D

**Interpret what each coefficient means from Part C.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part E

**Using your regression equation in Part C, calculate the expected logged average salary $(Y)$ for districts in Texas:**

- i. *Before* the switch that did *not* switch
- ii. *After* the switch that did *not* switch
- iii. *Before* the switch that *did* switch
- iv. *After* the switch that *did* switch

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part F

**Confirm your estimates in Part E by finding the mean logged average salary for each of those four groups in the data.^[Hint, `filter()` properly then `summarize()`.]**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part G

**Write out the difference-in-difference equation, and calculate the difference-in-difference. Make sure it matches your estimate from the regression.**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part H

**Can we say anything about the types of districts that switched? Can we say anything about all salaries in the districts in the years after the switch?**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part I

**Now let’s generalize the diff-in-diff model. Instead of the treatment and post-treatment dummies, use district-and year-fixed effects and the interaction term. Interpret the coefficient on the interaction term**^[This is doable with the dummy variable method, but there will be a *lot* of dummies! I suggest using `plm()`.]

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```

---

### Part J

**Create a nice regression table (using `huxtable`) for comparison of the regressions in (a), (c), and (i).**

---

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# Write your code here
```